### 2016-02-17 Manuscript.tex
Initial ShareLaTeX Import

\documentclass{sig-alternate}
\usepackage{tikz}
  \def\firstcircle{(90:1.75cm) circle (2.5cm)}
  \def\secondcircle{(210:1.75cm) circle (2.5cm)}
  \def\thirdcircle{(330:1.75cm) circle (2.5cm)} 
\usepackage{comment}
\usepackage{cite}
\usepackage{framed,graphicx,xcolor}

\usepackage{stfloats}

\usepackage[shortlabels]{enumitem} 
\usepackage{amsmath}
\usepackage{url}
\usepackage{balance}
\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\be}{\begin{enumerate}}
\newcommand{\ee}{\end{enumerate}}
\newcommand{\tion}[1]{\textsection\ref{sect:#1}}
\newcommand{\fig}[1]{Figure~\ref{fig:#1}}
\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
%\setlist{nolistsep,leftmargin=5mm}
%\usepackage[pdftex]{graphicx}
\usepackage{program}
\newcommand{\Sample}{{\bf SAMPLE}}
\newcommand{\PEEKING}{{\bf PEEKING2}}
%\usepackage[table]{xcolor}
\definecolor{darkgreen}{rgb}{0,0.3,0}
\definecolor{Gray}{rgb}{0.88,1,1}
\definecolor{Gray}{gray}{0.85}
\definecolor{Blue}{RGB}{0,29,193}
\usepackage{colortbl}
\usepackage{picture}
\usepackage{url}
\usepackage{hyperref}
%\usepackage{listings}
\DeclareMathOperator*{\argmin}{arg\,min} 
\DeclareMathOperator*{\argmax}{arg\,max}
\definecolor{lightgray}{gray}{0.8}
\definecolor{darkgray}{gray}{0.6}
\definecolor{Gray}{gray}{0.95}
\definecolor{LightGray}{gray}{0.975}

\definecolor{Code}{rgb}{0,0,0}
\definecolor{Decorators}{rgb}{0.5,0.5,0.5}
\definecolor{Numbers}{rgb}{0.5,0,0}
\definecolor{MatchingBrackets}{rgb}{0.25,0.5,0.5}
\definecolor{Keywords}{rgb}{0,0,1}
\definecolor{self}{rgb}{0,0,0}
\definecolor{Strings}{rgb}{0,0.63,0}
\definecolor{Comments}{rgb}{0,0.63,1}
\definecolor{Comments}{rgb}{0.5,0.5,0.5}
\definecolor{Backquotes}{rgb}{0,0,0}
\definecolor{Classname}{rgb}{0,0,0}
\definecolor{FunctionName}{rgb}{0,0,0}
\definecolor{Operators}{rgb}{0,0,0}
\definecolor{Background}{rgb}{1,1,1}
\bibliographystyle{unsrt}
 \definecolor{lavenderpink}{rgb}{0.98, 0.68, 0.82}
 \definecolor{celadon}{rgb}{0.67, 0.88, 0.69}
\newcommand{\G}{\cellcolor{green}}
\newcommand{\Y}{\cellcolor{yellow}}

\newcommand{\quart}[4]{\begin{picture}(80,4)%1
{\color{black}\put(#3,2){\circle*{4}}\put(#1,2){\line(1,0){#2}}}\end{picture}}

\definecolor{MyDarkBlue}{rgb}{0,0.08,0.45} 
\newenvironment{changed}{\par\color{MyDarkBlue}}{\par}
%\newenvironment{changed}{\par}{\par}

\newcommand{\ADD}[1]{\textcolor{MyDarkBlue}{{\bf #1}}}
\usepackage{times}
\pagenumbering{arabic} 
\begin{document}

\definecolor{shadecolor}{gray}{0.9}
\conferenceinfo{ISCE}{'16 Austin, Texas}

\title{How to Learn Useful Changes to Software Projects\\(to Assist Refactoring)}
\numberofauthors{4} 
\author{  
\alignauthor
Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
       \affaddr{CS, NC State, USA}\\
       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
\alignauthor
Andrian Marcus \\
       \affaddr{CS, UT Dallas  } \\ 
       \affaddr{Texas, USA}\\ 
       {amarcus@utdallas.edu}
       \alignauthor
Naveen  \\ Lekkalapudi\\
 \affaddr{Bellhops } \\ 
       \affaddr{Tennessee, USA}\\ 
       {navek91@gmail.com}
\alignauthor
Lucas Layman \\
       \affaddr{Fraunhofer CESE  } \\ 
       \affaddr{College Park, USA}\\ 
       {llayman@fc-md.umd.edu}
\setlength{\columnsep}{7mm}
}
\maketitle
\begin{abstract}

Business users now demand more insightful analytics; specifically, tools that generate ``plans''-- specific suggestions on  what to change  in order to improve  the predicted values.

This paper proposes XTREE, a planner for software projects. XTREE receives tables of data with independent features and a corresponding weighted class which indicates the quality (``bad'' or ``better'') of each row in the table. Plans are edits to  the rows which ensures the changed row is more likely to be of a ``better'' quality.  XTREE learns those plans by building a decision tree across the data, then reporting the differences in the branches from some current branch to another desired branch.
Using data from 11 software projects,  XTREE can find better plans compared to three
alternate methods. Those plans have lead to improvements with a median size of (56\%, 28\%) and   largest size of (60\%, 77\%) in (defect counts, runtimes), respectively.
\end{abstract}

\keywords{Defect prediction; Configuration; Planning; Decision trees.}
\section{Introduction}

Due to the continuous change undergone by modern software systems, developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}). These anti-patterns are generally viewed as poor solutions to recurring design problems. In addition to there being anecdotal evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{}\cite{}. Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{}\cite{}. This makes early detection of these code smells a necessity to maintain sufficient quality in software systems. This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed by (i) identifying potential anti-patterns; (ii) generating reliable refactoring options, and (iii) critiquing bad refactoring options.

This is quite similar to the concerns voiced by several business users, who are demanding tools that support business-level interpretations of their data. At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}. Panelists commented  ``prediction is all well and good, but what about decision making?''. Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself. So:

\bi
\item
Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to generate \textit{plans} to remove the bad smells
\ei
In response to this , we propose a novel {\em planning} method called XTREE for learning changes to a software system such that its performance ``improves'', according to some quantitative measure. This paper uses XTREE to reduce  the expected value of the defects in Jureczko et al.'s JAVA systems~\cite{jureczko10};

The contributions of this paper are (1)  the new XTREE algorithm and (2) an evaluation strategy that shows the algorithm performing significantly ``better'' than  the  planner  proposed in our prior   work~\cite{me12c}. In this paper, we focus on three research questions:

\bi
\item \textbf{RQ1: Effective: Can a planner suggest plans to improve the quality of a defective class?}
\item \textbf{RQ2: Succinct: Are the plans suggested by a planner succinct?}
\item \textbf{RQ3: Insightful: Do the plans tell us thing we do not expect?}
\ei

The rest of this paper  describes our data, our planners, and the experiment that
ranks XTREE against alternate approaches.  This is followed by notes on related work
and validity.
To enable reproducibility, all scripts and data used in this 
study are available online at \url{http://git.io/vG3DG}.

\section{Preliminaries}\label{sect:prelim}

\subsection{Code Smells?}

Anti-patterns are poor solutions and recurring design problems that are unwillingly introduced by developers while designing and implementing large object-oriented systems. The impact of these code smells have been subjected to several empirical studies. As a result of the, the identification of these anti patterns has been receiving more attention from both researchers and practitioners.

During lifecycle, a software system undergoes continuous changes in order to keep up with the 
business level demands \cite{36}. Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{24}.

Fowler \cite{24} and Brown et al. \cite{13} defined a catalogue of more than 30 anti-patterns. For each anti-pattern, they reported the definition of the anti-pattern as well as specific refactoring operations aimed at removing it in order to improve the quality of source code.

To illustrate this, consider for example developers adding several responsibilities to a class rather that including them in separate classes. This results in a class that grows rapidly, and when there are further added responsibilities, the classes grow and breed. Ultimately, the class becomes too complex and its quality deteriorates. Such classes are know as Large Class \cite{24} and could be the cause of a design problem in source code.

For a while now, there have been anecdotal evidence that design problems (such as anti-patterns) negatively impacts software comprehension and maintenance. However, in the past few years, anti-patterns have been the subject of several empirical studies aiming at empirically analyzing their impact on software maintainability. From these studies, we now have empirical evidence that code containing code smells or participating in anti-patterns is significantly more change-prone than ``clean'' code \cite{31}. Also, code participating in anti-patterns has a higher fault-proneness than the rest of the system code \cite{31}\cite{37}.

Other studies aimed at understanding the impact of anti-patterns on program comprehension \cite{1} showed that the presence of an anti-pattern in the source code does not decrease the developers’ performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{1}\cite{61}. While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern. In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{3}\cite{17}.

All these findings suggest that code smells and anti-pattern need to be carefully detected and
monitored and, whenever necessary, refactoring operations should be planned and performed to deal with them. Unfortunately, the identification and the correction of design flaws in large and non-trivial software systems can be very challenging. These observations call for \textit{planners} capable of supporting the software engineer by (i) identifying anti-patterns, and (ii) designing and applying a refactoring solution to remove them. Such planners, are the focus of the current paper.

\subsection{What is a ``Plan''?}

Our planners   use tables of data with independent features and a dependent class
feature. Classes have weights  that  indicate  what rows are ``bad'' or ``better''. Plans  change   a row such that it is more likely to be ``better''. 
Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
  adjust   feature $Z_j$:

{\small\[
\forall \delta_j \in \Delta :  Z_j =  
\begin{cases}
     Z_j + \delta_j& \text{if $Z_j$ is numeric}\\
    \delta_j              & \text{otherwise}
\end{cases}
\]}

For example, to simplify a  large bug-prone  method, our planners might suggest
to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
two simpler functions).

Additionally, if a developer were to suggest a specific refactoring option, our planners can query the option and report on its possible usability. Note that we make no assumption that a plan mentions every feature
(so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than
plan2).

\subsection{From Detection to Planning}\label{sect:p2p}

This paper is about the next step {\em after} detection of bad smells. Suppose developers are presented with  a list of anti-patterns and they do not like what they see. This user may then ask a {\em planning}  question; i.e. ``what can we change to do better than that?''.

Before exploring automatic methods to answer the planning question, we first comment
on two manual methods.

One way to propose changes to a project would be to   ask some smart experienced
person for their opinion on the impact of these anti-patterns and the possible refactoring options. Most advices offered by these developers are based on experiences, sometimes
such advice is an effective strategy and sometimes it is not.

According to Passos et al.~\cite{passos11},  developers
may  assume that the lessons they learn from a few past
projects are general to 
all their future projects. They comment ``past experiences were taken into account without 
much consideration for their context''~\cite{passos11}.  
J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that the 
supposed software engineering  ``Gurus'' rarely use lessons
from past projects to improve their future reasoning and that such poor
past advice can be detrimental to new projects.~\cite{jorgensen09}.
Accordingly, we   propose a ``trust, but verify'' approach.
After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
if there are any better options
(just as a sanity check).
The rest of this paper discusses some methods to build automatic oracles 
to implement that   sanity check.

Another way  to find  changes to a project
might be to rely
on the peer review processes used by the 
SE research community. This approach would propose changes to software
projects that concur with internationally accept best practices. 
There are two
problems with that approach. Firstly, given the rapid pace of change in software
engineering, we may be asking questions for which there is no current widely
accepted concept of
``best practice''.

Secondly, given the diversity of SE products and practices
and personnel, it may well be that the current project being discussed is 
substantively different to prior work. 
Numerous recent {\em local learning} results compare (1) single models
learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
A repeated result in those studies is that the local models generated better effort
and defect predictions (better median results,
lower variance in the predictions). \tion{surprise} of this  paper offers yet another locality result:
\bi
\item
One standard rule in the literature
is that it is useful to implement modules such that they are internally cohesive (use
much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
\item
While that may be true in general, for particular classes other changes may be more important
(later in this paper, we show one set of results where that is indeed the case).
\ei
In summary, 
it is useful to have automatic methods to recommend changes. Such
methods can fill in for human expertise (if such experts are absent) or 
to offer a second opinion.
Also, prior to making automatic recommendations, it is wise to first stratify the data
(clump it into related examples) then generate advice specific to each clump.
Accordingly, the rest of this paper defines and evaluates
automatic methods to find plans from
  $N$ examples divided  into  many clumps.

\input{tex/jur_char.tex}

%XXX end

%  \begin{figure}[!t]
%  \small
%  \begin{center}
%  \begin{tabular}{r|rr}
%  data set & cases & \% defective\\\hline
%   ant &947& 22\\
%   camel& 1819& 19\\
%  jedit& 1257& 2\\
%  ivy &352& 11\\
%  log4j& 244 &92\\
%  lucebe &442 &59\\
%  poi& 936 &64\\
%  synapse &379 &34\\
%  velocity& 410& 34\\
%  xalan& 2411& 99\\
%  xerces &1055& 74
%  \end{tabular}
%  \end{center}
%  \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
%  \end{figure}

%  \begin{figure}[!t]
% \scriptsize
% \begin{tabular}{llllll}
%   \hline
%   \rowcolor{lightgray}
% Project & Domain & Lang. & LOC & Features & Config\\\hline
% BDBC: Berkeley DB   & Database & C & 219,811 & 18 & 2560\\
% BDBJ: Berkeley DB   & Database & Java & 42,596 & 32  & 400\\
% Apache & Web Server & C & 230,277 & 9 & 192\\
% SQLite & Database & C & 312,625 & 39 & 3,932,160\\
% LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
% x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
% \end{tabular}

% \caption{Siegmund data.}\label{fig:cpm}
% \end{figure}

% \begin{figure}[!t]
% \includegraphics[width=0.9\linewidth]{figs/BDBC.eps}
% \caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
% \end{figure}

\section{Test Data}\label{sect:tesd}

To assess our planning methods, our data (see git.io/vGYxc) comes from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}.

The Jureczko data records the number of known defects for each class using a post-release bug tracking system. The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc. For details on the Jureczko data, see  \fig{ck} and \fig{jd}. All the planning methods of this paper reflect on the numeric value of the raw defect counts. The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise. The choice of this data set was is based on the taxonomy of the smells, in which we focus on the measurable properties needed to detect a given smell. This set of metrics is not restricted and can be easily extended with other metrics.

% The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To obtain the  data, Siegmund et al. perturbed
% the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
% Berkeley database (one written in ``C'' and one in Java). Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}. It's worth noting that in each of the above data sets, several features are interdependent, this is expressed using a \textit{feature model}.
% \fig{bdbc} shows an example of such a feature model defining valid combinations of settings in the Berkley database (``C'' version). These feature models were used by Siegmund et al. to ensure all their perturbations are valid
% (we will use the same models to cull invalid plans).

\section{Planning Methods}\label{sect:planners}

This section describes XTREE (which we call Method4) and  three  
alternate methods for learning plans.
  XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
  The other methods use the   top-down
	bi-clustering method described in \fig{where}.C  which recursively divides the
	data in two  using a dimension that captures the greatest variability in the data. 
	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}. 
All  methods have the  properties proposed in \tion{prelim}: (1)~they are {\em local learners}
that find   plans   particular to each test case; and (2)~they  are {\em trust-increasing}; 
  i.e. the changed examples are moved {\em closer} to the training data.

 \subsection{  Methods}

Our  description of the methods adopts the following convention. All variables
  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$.
  In this paper, we show our current settings to these variables produces useful
  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
  find better settings but  we have nothing definitive yet to report
  on auto-tuning planners.

\subsubsection{Centroid Deltas (CD)}

 {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be  (which we will call $C_j$).

{\em Assumption1: } Large data sets can be adequately represented by a few dozen (or so) 
centroids.

 {\em Details1:} Method1 clusters project data by reflecting on the independent variables, then
  reports the delta between the cluster centroids. 
  After clustering training data using the WHERE algorithm of \fig{where}.C, Method1
replaces all clusters with a  centroid $C_i$ computed from the median value of each
continuous/discrete feature. Then, it
finds the closest centroid $C_j$ that has a better
performance score. For Jureczko data, ``better'' means fewer defective examples while for the Seigmund data,
``better'' means lower median runtimes for the examples in that cluster.
Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
in $C_j$. 
Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
the nearest centroid $C_i$.  It then proposes a plan for improving that test case
that is the conjunction of all the deltas between $C_i$ and $C_j$.

%\input{tex/pseudo-code-1.tex}

\input{tex/dtree.tex}

 \begin{figure}[!t]
{\small 
\[
\begin{array}{r} 
\mathrm{project}\\
\mathrm{data}
\end{array} 
\left\{\begin{array}{l}\mathit{train}
        \left\{\begin{array}{l}
                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
              \end{array}\right.
       \\
      ~\\
\mathit{test}  
    \left\{\begin{array}{l@{~}l}
           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
           \mathrm{\bf then} &
           \left\{
            \begin{array}{l}
                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
            \end{array}
          \right.
   \end{array}\right.
\end{array} \right. 
\]}
 \caption{Experimental design .}\label{fig:design}
 \end{figure}

\subsubsection{XTREE}

XTREE builds a decision tree,  then generates
plans from the difference between two branches:
(1)~the branch to where you are and (2)~the branch to where you want to be.

One potential problem with Method 1 is the {\em unsupervised} nature of
the clustering algorithm (WHERE) that  
executes without knowledge of the target class.  {\em Supervised} methods assume that it is useful to also reflect on the target class.

XTREE uses a supervised   decision tree algorithm of \fig{where}. D to divide the data.
Next, XTREE builds plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
That code asks three questions, the last of which returns the plan:
\be
\item
What {\em current} branch does a test case fall in?
\item What {\em desired} branch would the test case want to move to?
\item What are the {\em deltas} between current and desired? 
\ee

\section{Evaluation Strategy}
\input{tex/prediction_stats.tex}

Our evaluation strategy (discussed below) divides this data into a training and a testing set.
From the training set we apply a data miner (to learn a quality predictor) and
various planning methods (to learn different plans). Next, we try applying
each of those plans to the  test set and ask the quality predictor to assess the changed examples.
Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
in the changed examples.

As mentioned in the last section,  this approach depends on having effective predictors for assessing the results. It proved to be  more complicated to commission the Jureczko data sets for this study. For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could
be salvaged using the techniques discussed in this section.

%   For the Siegmund data, this criteria was   relatively easy to achieve.
%   The data in those data sets have a continuous class (runtime of the compiled system)
%   so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
% This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
%   x264. 
%   As a preliminary study, we split that data into equal sized train:test groups
%   and trained a Random Forest
%   Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other. After repeating this process 40 times, we achieved nearly perfect accuracy of $s=$\{98.45\%, 99.62\%, 94.86\%, 98.89\%, 98.60\%, 99.87\%\};
%   i.e.  we are confident that the predictors from the Siegmund data can assess
% our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)

 \fig{j} shows our preliminary studies with the Jureczko   data.
Given access to $V$ released
versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
shown in \fig{j}, this means that we are training on hundreds to thousands
of classes and testing on smaller test suites).
Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).  In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).

In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
For such data, quality of the predictor can be measured using
(1) the  probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
the test data detected
by the {\em predictor}; and (2) the 
probability of false alarm (a.k.a. ``pf''): the percent of non-fault
classes that are {\em predicted} to be defective.

As a preliminary study, we split the Jureczko  data into train and test groups.
Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was
built from the training data, then applied to the test data.
The ``untuned'' columns of \fig{j} shows those results. We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly. For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets. With this threshold, however, none of our data sets were be suitable for this study.

Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets. Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e. an over-sampling of minority-class examples and an under-sampling of majority-class examples. Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
for tree generation, etc. The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques. For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%. However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.

In summary, while we cannot trust predictors from some of our Jureczko data sets,
we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
Accordingly, when this study explores the Jureczko data, we will use these five data sets.

(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
parameter tunings, those techniques were applied to the training data and {\em not}
the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)

\section{Experiments}

This section describes an experimental design (and results) for evaluating the above four methods. 
\subsection{Experimental Design}

\subsubsection{A Strategy for Evaluating Planners}

Our experimental design is shown in \fig{design}. We divide the
project data  into two disjoint sets {\em train} and {\em test}
(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
Next, from the train set, we build both a {\em planner} and
 a {\em  predictor}.

Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
\bi
\item Our {\em planners} will be XTREE and CD;
\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.   We use these
data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
\ei
As for the {\em test} data, this is passed to the { predictor}
to measure performance statistics related to effectiveness.

If our { predictors} fail to perform effectively on the test data,
then we cannot trust them to comment on our plans. Accordingly,
if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
we should not use some of the  Jureczko data.

Else, we (1)~apply the { planner} to alter the {\em test} data;
then (2)~apply the { predictor} to the altered data $test'$;
then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
\bi
\item If $R  = 0\%$, this means  ``no change from baseline''; 
\item If $R \gt 0\%$, this indicates ``

improvement over the baseline'';
\item If $R \lt 0\%$, this indicates ``optimization failure''.
\ei
 \subsubsection{Statistical Methods}
 Our methods use some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs,
  with different random number seeds
(we use 40 since that is  more than the 30 samples  needed
to satisfy the central limit theorem).

  To rank our methods using the results from these 40
  repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}.

In accordance to that test, using the median values of each method, we
sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
Then, we
split $l$ into sub-lists $m,n$ in order to maximize the expected value of
 differences  in the observed performances
before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
 \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]

We then apply a apply a statistical hypothesis test $H$ to check
if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping). 
If so, Scott-Knott  recurses on the splits.
In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).

For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}. For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}. These researchers warn that even if a hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small. Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.

\subsubsection{Report Format}

Our results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
The black dot shows the median $R$ value and the horizontal likes stretch 
from the 25th percentile to the 75th percentile (a region called the inter-quartile
range, or IQR).

\begin{center}

{\small Example \begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
\arrayrulecolor{lightgray}
\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
\hline 2 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
\hline \end{tabular}}
\end{center}

In this example table, the rows are  sorted on the median values of each method. Note that all the methods
have   $R\gt0\%$; i.e. all these methods reduced the expected value of the performance score in that experiment
while XTREE achieved the greatest reduction (of 62\% from the original value).

The above example table has a  left-hand-side  {\bf Rank} column, computed using the
Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and CD+FS together are ranked the worst.

\subsubsection{Other Details}

\fig{jur} shows the effectiveness of our methods seen in 40 repeats with each data set.
In these experiments,   the dependent variables of Jureczko data set is discrete. Hence, while choosing the predictor, we used Random Forest as a classifier for Jureczko data.

% For Siegmund data, we performed a k-fold cross validation. Given the relatively small sample sizes in Apache and BDBJ, we chose $k=2$. This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably. However, in 2-fold cross validation, the division of data tends to affect the outcome significantly. To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the other, repeating the process 40 times as mentioned above.

The Jureczko data, being temporal in nature, allows us
to implement a validation procedure that ensures only past data is ever used to
predict future values. Hence, in that data, we used the train/test sets shown in \fig{j}. 
(Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).
\input{tex/results_jur.tex}

\subsection{Experimental Results}

Recall from our introduction that we are assessing planners on three criteria:
{\em effectiveness}, which is how much they reduce the expected value of the changed examples;
{\em succinctness}, which is how many things we need to change to achieve a plan;
and {\em insightfulness}, which is how different are the plans from standard truisms.

\begin{figure*}[!t]
\centering
\includegraphics[width=\linewidth]{figs/Deltas-both.eps}
\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
\end{figure*}

% \includegraphics[width=\linewidth]{figs/ant.png}

\begin{figure*}[htbp!]
\centering
\includegraphics[width=\linewidth]{figs/changes.png}
\caption{Suggested Changes}
\label{figs:changes}
\end{figure*}

\subsubsection{RQ1: Effectiveness: Can plans to improve the quality of a defective class?}

Measured in terms of effectiveness, some data sets were harder to optimize than others. However, in most data sets, large reductions were observed. We noticed an improvement of 85.54\%, as compared to the original baseline in Ant and 36.36\% in Lucene, \fig{jur}.

Overall, XTREE was most the effective. It was the top-ranked method in 3 cases and performed just as well in the other two. Where it ranked better, it had significant improvements in the median performance values. In the Jureczko data sets, there was an improvement of at least 25\% compared to our previous method.

\subsubsection{RQ2: Succinctness: Are the plans suggested by a planner succinct?}

\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
The left-hand-side plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the right-hand-side
  shows a Siegmund data set ({\em BDBJ}).

  In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e. the vertical bars in that plot
  are {\em lower}). For example,
  XTREE's plans were usually succinct --  in all  data sets,
 XTREE   changes around a fifth of the features (see \fig{types}).
On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features 
 (observe the change frequencies as 
high as 100\% for all features). 
Method1's policy of ``change everything''  might be acceptable
if this approach lead 
to the most effective changes. However, in
\fig{jur} and  \fig{conf1}, there is no evidence for this.

An  interesting feature of \fig{changed} was that fewer things 
were changed in the Seigmund data sets like {\em BDBJ} than in the Jureczko data set like
{\em lucene}. 
In turns out that this holds true across nearly all our data sets.
\fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
there are fewer changed features in the Seigmund data than in the defect prediction data. 
One explanation for that is the nature of the features: the
Jureczko data sets have continuous
features while the 
Seigmund data has binary independent features (where settings can be turned ``ON'' or ``OFF''). While the features of the Jureczko data sets are not constrained, those of Seigmund data sets are subject to several constrains (dictated by their feature models, see \fig{bdbc}). It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores, thereby making insignificant changes. As a result all methods, except XTREE, fail to optimize the runtime, as evidenced by the results in \fig{conf1}.

\input{tex/deltas_avg.tex}

\subsubsection{RQ3: Do the plans tell us thing we do not expect?}\label{sect:surprise}

If a planner only ever reported conclusions that were already known, then that planner offers
little value over ``just use established wisdom''. Accordingly, we studied our results
for plans that were somewhat counter-intuitive.

Such a surprising plan can be  seen in {\em lucene}. Recall the standard advice for
OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
to specific projects by checking how often a planner changes the
 coupling-related features:
\bi
\item {\em ca}:   afferent couplings  =  \# classes using this
				class;
\item {\em  	ce}:  efferent couplings =  \# classes  used by this
				  class. 
\item {\em cbm}: coupling between methods =  \# new/redefined methods
				to which all the inherited methods are coupled
\item	{\em cbo}:  coupling between objects = a value that increases when the methods of one
				class access services of another.

\item {\em ic}:   inheritance coupling =  \# parent classes  which a given
				class is coupled (including methods and variables inherited)
\ei
In many of our results with the Jureczko data, it was indeed true that the changes
proposed by XTREE lead to lower coupling. However, the {\em lucene} results were quite
different and rather surprising.

In the {\em lucene} XTREE results  of  \fig{changed},
the most frequent change was to alter the lines of code in a class (see the tallest
red histogram in that figure on the {\em loc}, or lines of code). 
Looking at the logs of our planner, we can see
that XTREE's proposed change is to {\em reduce} the size of a class.
The only way to do that,  while keeping the  same functionality, is to create a network
of smaller classes that interact to produce that functionality.
That is, we would need to {\em increase} the coupling of those classes to achieve XTREE's plan.

In theory, increasing coupling between classes complicates and confuses a class design.
But the {\em lucene} XTREE results  of  \fig{changed} rarely proposes changing   the
coupling features {\em ca, ce, cbm, ic} (in fact, XTREE never proposes any change to {\em ic}).

The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
XTREE is warning against, say, some factory class generating a large
community of agents, all of the same class, who
co-ordinate on some task. This is a different issue to the class redesign issue that would
be triggered by altering {\em loc}.

In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
At least for the {\em lucene} data set, we can see advice that recommends {\em increasing coupling}
to reduce defects.

\section{Threats to Validity}\label{sect:valid}

As with any empirical study, biases can affect the final results. Therefore, any
conclusions made from this work must be considered with the following issues in
mind.

\subsection{Trusting the Changes}\label{sect:trust}
   XTREE is evaluated by  comparing
predicted performance scores before and after a planner makes changes to the feature values of an example:
After making those
changes, we may have a new example that has never been seen before. Therefore, it must be asked
{\em ``can we trust the predictions made on such new examples?''}

To answer this question, we note
that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples
(for a visualization of these clouds, see \fig{howxy}).
\begin{figure}[!t]
  \includegraphics[width=\linewidth]{figs/2d.eps} 
 % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
\caption{Gray, red, green show (1) training examples, (2) test examples and 
  (3) tests that have been altered by planners.
  This figure uses axes generated from the first two components of a PCA analysis of all points. 
}\label{fig:howxy}
\end{figure}
We should mistrust the predictions made by a  model   if it is being applied to examples  that are
too far away from the
training cloud.
To test for ``too far'', we can run a data mining experiment that tests how well
a model learned from the training data applies to the test data. Such experiments return some performance value.

Note that predictions  about changes that  fall within the space of the training+test data, will be at least
as accurate as the predictions of the original test data. With this, we assert that the predictions for changes that move examples towards/away from the training data can be trusted more/less (respectively).

Accordingly, we need {\em trust-increasing} planners to generate new examples {\em closer} to the
training examples.  To see how this works, 
 \fig{howxy} is from the {\em ivy} data
set (one of the Jureczko data sets used in this paper). It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
changed  examples displaced after applying a plan (in green).
 Note that the  the   changed examples
cases  (shown in green)  fall closer to the training cases (shown in gray) than
the test cases (shown in red).

In that green region of changed examples, our belief in the value of predictions
will be just as much as, if not more than, our belief in the value of the predictions in the red region (that
contains the original test data).
This pattern of \fig{howxy} (where the new examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
paper. Hence,  we can assert that
predictors learned from these training examples have some authority in the regions
containing the changes examples.

That said, the above comes with some important caveats:
\bi
\item 
The quality of the prediction depends on the nature of the training data. Thus, we strongly recommend that both the data set and the predictor be assessed prior to planning. This ensures that the predictor's performance is adequate for a data set. We tackle this issue in detail in \tion{tesd}.
\item
Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
\item
Where possible, planners should be assessed via some external
oracle that can accurately assess new examples. For an example of that kind of analysis,
see  \tion{coc}.
\ei

\subsection{ Learner Bias}
For building the defect predictors in this study, we elected
to use  Random Forest  and Random Forest Regressors .
We chose this approach,  based on its reputation for having the better  performance of 
21 other learners for defect prediction~\cite{lessmann}.
Data mining is a
large and active field and any single study can only use a small
subset of the known classification algorithms.

That said, we have taken care to document in this paper the decisions made by engineering
judgement that could effect our conclusions. The above code used a set of variables which future
work should vary in order to test the internal validity of our conclusions:
\bi
\item All our planners divide data into groups of size $\alpha=\sqrt{N}$;
\item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);
\item Method3 uses  $\omega=0.5$ to find its deltas;
\item XTREE considered a sibling useful if it's score was less than $\gamma=0.5$ times the mean score of the current leaf.
 \ei

\subsection{  Sampling bias} 
Sampling bias threatens any data mining experiment; i.e., what matters
there may not be true here. For example, the data sets used here comes from two sources
(Seigmund et al. and Jureczko et al.) and any biases in their selection procedures
threaten the validity of these results. 
That said,
the best we can do is define our methods and publicize our data and code so that other researchers can
try to repeat our results and, perhaps, point out a previously unknown bias
in our analysis. Hopefully, other researchers will emulate our methods in
order to repeat, refute, or improve our results.

\subsection{  Evaluation Bias}\label{sect:coc}
Another threat to validity of this work is our use
of predictors learned on the training data to assess the impact of our planners.
This issue was discussed in detail in \tion{trust}.

To those notes, we add a few more details. If possible, planners should be assessed via some external oracle that can accurately assess new examples. For example, in search-based software engineering,
examples can be assigned objective scores via  some model. In this approach, a changed example can be assessed by
generating actual objective scores from the model.

\input{tex/results_models.tex}

The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring management challenges. POM3 implements the Boehm and Turner model of agile programming~\cite{boehm2003balancing} where teams select tasks as they appear in the scrum backlog.  POM3 can study the implications of different ways to adjust task lists in the face of shifting priorities. The model outputs estimated task completion rates; programmer idle rates; and total overall cost. POM3 models requirements as a dependency tree. A single requirement in the tree of a prioritization value has a cost, along with a list of child-requirements and dependencies. Before any requirement can be satisfied, its children and dependencies must first be satisfied. POM3 simulates changing priorities by making teams aware of random items in the requirements tree at random intervals, thus forcing teams to constantly readjust their "to do" lists. For further details on this model see~\cite{boehm2003using},~\cite{port2008}, and~\cite{boehm2003balancing}.

\fig{coc} shows results from using POM3 as an oracle to assess our planning methods and their ability to reduce the project cost. In this experiment, we generated a training and testing set with 1000 randomly generated instances, which we passed to our four methods.
40 times, we let those methods propose changes to those projects. 
For assessment purposes, the changed projects were then fed back to the POM3
oracle.

Using this approach, it is possible to assess the value of a plan by measuring its
effectiveness with respect to some ground truth (in this case, the POM3 model).
As shown in \fig{coc}, XTREE reduces the cost by 59\% thereby passing this assessment. These results from the POM3 model further endorse our previous conclusions; i.e. compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.

\section{Related work}

\subsection{Planning in AI}

The XTREE planner is somewhat different to the logic-based planners explored by 
classical AI. 
Those kinds of planners employ a logical procedure~\cite{Fikes1971}
that seeks an ordering on {\em operators} to take some domain
{\em state} from a {\em start} state to a {\em  goal} state.
This classical logical approach is known to suffer from
computational bottlenecks~\cite{Bylander1994}. On the other hand, tools like XTREE will scale to any domain
that can generate decision trees.

\subsection{Evaluating Changes}

Some organizations have the resources to 
run repeated trials to assess  project changes.
For example, in one   study, Bente et al. reported results
where the same  specification was developed by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
then check in 12 months time
 which teams have fewer defects than the others.

Seldom do industrial or research groups have access
to the kinds of resources needed for this kind study  (evidence: in the six years since the
publication of that work, we know of only one   similar study to Bente et al.). Also, given the
diversity of modern software projects, it might be unreasonable to demand that all
proposed changes for all projects are always evaluated by something like the Bente et al. study.
Hence, this paper has used data miners to build an oracle that can assess changed examples. The advantage
of our approach is that it required far less resources to assess the effectiveness of proposed
changes to a project.

\subsection{Search-based SE}

Another way way to propose changes to software artifacts
is   via some search-based method~\cite{Harman2009,Harman2011}. Such SBSE methods are   evolutionary programs that 
make
 extensive changes to  some initial sample of project data
 (perhaps 
100s to 100,000s of mutations). Each of these mutations
is reassessed using some domain model.
Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}.

One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
to prevent new examples from mutating away from the kinds of data used to commission the model (in which case, we would
start doubting the model's output).

Another issue with standard search-based SE methods is that they require ready access to 
trustworthy domain model that can offer an assessment
of newly generated examples. While some domains have such models (e.g. see the COCOMO effort estimation model
used in the last section), our experience is that many others do not.  For example, 
consider software defect prediction and all the intricate issues that may lead to defects in a product. A model that includes {\em all} those
potential issues would be very large and complex. Further,
the empirical data required to validate any/all parts
of that model can be hard to find.

What we would recommend is a two-pronged policy.
In domains with ready access to trusted models, we recommend
the kinds of tools that are widely used in the search-based
software engineering community such as GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like XTREE.

\section{Conclusions}

The planner proposed in this paper proposes changes to software project details in order to improve the expected
value of the performance scores of that part of the project.
To evaluate these planners,
data miners can be used to build oracles to assess planners.
Such planners should be {\em trust-increasing}; i.e. they should propose changes that generate
changed examples that are closer to the training data of the data miner.
One caveat here is that the evaluations we can make on the planner are only as good as the predictive
performance of the data miner. Hence, if domain data does not support satisfactory predictors, then
planning in that domain cannot be evaluated.

Four planners were assessed here for the tasks of reducing defects and runtimes. 
Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
paper is that a novel fourth method clearly out-performs the other three
(measured in terms of {\em effectiveness, succinctness, and surprise}).
We conjecture that XTREE worked better than the rest since:
\bi
\item
It uses a supervised method to divide the data and;
\item
When planning how to move examples to better classes, it is best to reflect over 
differences between those classes.
\ei
\section{Future Work}
Future work in this area could explore numerous questions. For example, 
XTREE, as used here, sought improvements in a one goal (the class variable). Does XTREE work for multi-goal reasoning?

Also, the XTREE algorithm seems quite general to any data of table with rows containing
weighted classes (so we can distinguish ``bad'' rows from ``better'' ones). 
Does   XTREE works  on   domains (other than the defect/runtime data explored here)?

As an example of a domain that might benefit from XTREE, recent results raise doubts about
the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``bad smell'' detector to select the subset of possible refactorings that have the most potential benefit?

As to scalability, XTREE is a post-processor to a decision tree algorithm. Hence, in theory, XTREE   works on domain where data miners can generate decision trees. Given the current state of the art in Big Data, can XTREE  be applied to  very large data sets?

We discussed above in \tion{p2p} the general  conclusions of Passos, J{\o}rgensen  et.al~\cite{passos11,jorgensen09};
i.e. software developers are reluctant to surrender their old biases
in the face of new data. 
Accordingly, it must be asked if  the {\em mental resistance} of developers
will prevent them applying XTREE's automatically generated recommendations of tools?
Note this this issue is not just a concern for XTREE, but also for any 
automatic tool proposing refactorings.

There are many more methods for generating plans and
no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta$, and $\gamma$
parameters that controlled XTREE. Would we get better results if we varied those parameters?

That said, the goal of this paper was not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
was to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
The authors of this paper would actively support other teams exploring this
method (with or without using our current code base). So our last questions is this:
will other  researchers try to repeat and/or improve
(or even refute) our results?  Perhaps.

\section*{Acknowledgements}
The work has partially funded by NSF  award \#1506586.
\bibliographystyle{plain}

\balance
\bibliography{References}
\end{document}

  Next, build one centroid for each cluster (using the median and mode value for continuous
and discrete values, respectively).
After that, write a spreadsheet with one column per centroid;
\bi
\item
Sort the columns such that any current projects client fall into the left-most
columns. That is, make the left-hand-side of the sheet  ``current usual practice''
and the right-hand side  ``alternatives to current usual practice''.
\item
Sort the rows of that spreadsheet such that the rows with
maximum  variability appear at the top 
\ei
Now show the sheet to the business user, encouraging them to make recommendations by
\bi
\item
Comparing the left and right-sided columns;
\item
Using the features mentioned in the top-most rows (where changes selects for the most different centroids).
\ei

In theory, One  advantage of this method is that it focuses the attention of the business users
away from rows that do not select for different centroids and towards differences between
current usual practice and everything else.  
In the summer of 2011 and 2012, one of us (Menzies) spent two months
	working on-site at Microsoft Redmond,
	observing data mining analysts.  In that study, he took special
	note about how Microsoft's data scientists
	discussed the results of their data mining sessions with  business users.

	One surprising observation was how  
	little time was spent by business users 
	inspecting  of the output of standard data miners. Prior to that visit,
	we had the mistaken impression that   decision trees,
	clustering algorithms, etc were useful ``off-the-shelf''; i.e    business
	users would inspect and understand the output of those tools.

Standard data mining tools are not necessarily the best tool for supporting that dialogue.
Menzies found that he had to do considerable work pre-processing data mining output
prior to the weekly briefing meetings for the business users. Initially,
that pre-processing was just clustering. This evolved into feature selection and finally
the a case-based reasoning tool called HOW.

	as compared to another process, which we call {\em peeking}.
	In {\em peeking}, analysts and users spend much time
	inspecting and discussing small samples of either raw or exemplary or synthesized project data.  Further, very little of those discussions were  focused on classification
	(the addition of a labels to some unlabelled data). Rather, much time
	was spent in those meetings discussing {\em what to do next}; i.e. trying
	to determine what could be altered to better improve some business outcome.

	That   Microsoft  study found two common ``peeking'' methods.
	In {\em data engagement meetings},
	users debated the implications of data
	displayed on a screen. In this way, users
	engaged with the data and with each other by
	monitoring each others' queries and check each others'
	conclusions.

	Another data analysis pattern observed
	at Microsoft was  {\em cluster + contrast} in which
	data is  reduced to a few
	clusters. Users are then just shown the delta between those
	clusters. While contrasting, if feature values are
	the same in both clusters, then these were pruned from
	the reports to the user. In this way, very large
	data sets can be shown on one PowerPoint
	slide. Note that {\em cluster+contrast} is a tool that can be usefully employed within
	{\em data engagement meetings}.

	Cluster+contrast and engagement
	meetings are common practices at Microsoft. Yet  these methods had never been rigorously studied or certified.
	For both those reasons,
	we reflected over those tools to discover and analyze their
	underlying process. The result was HOW~\cite{howase}: a tool
	that combines (a)~feature selection; (b)~centroid generation from   clusters;
	(c)~contrast methods between centroids.
	While method (a) is widely used (e.g.~\cite{Menzies2010}),
	to the best of our knowledge, this combination of (abc) has not been thoroughly explored before.

This paper assesses two methods  for ``peeking'': a model-based method called DTREE and an instance-based method   called HOW. 
DTREE were first developed by Lekkalapudi and Menzies to  explaining results from multi-objective optimizers~\cite{nva14}. This paper is the first
to apply DTREE to defect prediction and contrast set learning. Also, that prior work evaluated
DTREE against multi-objective optimizers and not   instance-based methods.

This paper uses three criteria to assess the value of HOW and DTREE for
learning actionable analytics:
\be
\item
A planning system needs to be {\em effective}; i.e. if its recommendations
are applied then some statistically significant change should be observed.
To predict the number of defects in a data set before and after applying our changes,
we build a    prediction system (built by data mining; specifically: Random Forest). Note that this
predictor was built from some hold-out data (i.e. from  different data than that used
to build the predictor).
\item
Any  conclusion made to a business user must be understandable;
it must generate {\em succinct} changes.  
\item
Recommendations should be {\em stable}; i.e. they shouldn't widely vary due to minor changes in the data. Hence, in our experiments, we will add a little randomness to our analysis then report results across 20 repeated runs. 
\ee
We will find that
 {\em effectiveness} of DTREE and HOW are similar, but   DTREE wins on {\em succinctness} and {\em stability}.

\section{Cluster and Contrast}
\label{clust_contrast}
A recurring data analysis pattern in this paper is $cluster+contrast$. The data is distilled into a few clusters using a clustering scheme. Then lessons are inferred by studying the differences between these clusters. These lessons are used to generate rules that can be applied in any context.

\subsection{Clustering}
There are a wide variety of clustering methods to choose from. A study by Ganesan~\cite{div14} explored different clustering methods for software engineering data using the effort and Jureczko data from the PROMISE repository~\cite{promise}. In that study methods such as WHERE, K-Means, mini-batch-K-Means, DBScan, EM, and Ward were investigated. The results of the study showed that the size and number of clusters is more important that the specifics of the techniques used.

For this purposes of this work, we have chosen WHERE, a clustering scheme which is capable of generating at least $\sqrt{N}$ clusters given $N$ instances. In addition to this, WHERE has been shown to run fast while ignoring spurious dimensions~\cite{menzies2013}. This is particularly useful, for much of the SE data is noisy, they contain a information not associated with the target variable.

\subsection{Finding Contrasts}
All the following methods use clustering in the form of WHERE, a top-down clustering method which recursively splits the data in two along a dimension that represents the highest variability. It works as follows:
\bi
\item Find   two   distance samples from the data, say  $X,Y$. This can be done by  picking any case $W$ at random, then setting $X$ to its most distant case, then setting $Y$ to the case most distant from $X$ (this requires only $O(2N)$ comparisons of $N$ cases).
\item Project each case $Z$ onto a {\tt Slope} that  runs between $X,Y$ using the cosine rule. 
\item Split the data at the median $X$ value of all cases and recurse on each half  (stopping when one half has less  than $\sqrt{N}$ of the original population).
\ei

Clustering is followed by generating \textit{contrast sets}. These contrast sets represent recommendations on what could be altered to better improve an outcome. In this work we have explored several algorithms as possible tools to identify contrast between clusters. These fall into two broad categories:
\begin{enumerate}
\item Case Based Reasoning techniques (Nearest Neighbors and a gradient base planner called HOW)
\item Decision Trees.
\end{enumerate}

These techniques are discussed below. It is worth noting that the following techniques are organized in such a way that each one seeks to address certain shortcomings in the ones that precede it.

\subsection{Case Based Reasoning}

Case-based reasoning seeks to find solutions to problems by emulating human recollection and adaptation from past experiences. It has found extensive usage in Artificial Intelligence because it offers several advantages. However, one of the most important benefits that CBR has to offer is that it works on a ``case-by-case'' basis. Therefore it's advise is tailored to be specific to the particular case being considered. Several paper in SE have applied this technique, most of all for effort estimation ~\cite{keung2008analogy, 6600685, walkerden1999empirical, shepperd1997estimating, kocaguneli2010use}.

A classic example of a CBR is K Nearest Neighbor. Our version of nearest neighbor for planning is rather straight forward. It has been developed as a "straw-man"; i.e., a simple baseline tool to act as a benchmark used to evaluate other methods.

\subsection{HOW}

HOW is very different compared to conventional CBR planners in that it explores the gradient between pairs of nearby clusters instead of studying the clusters themselves. HOW works by clustering the data during training using WHERE and then drawing \texttt{slopes} between the centroids of pairs of nearby clusters. Assuming the cluster pairs are labeled X and Y, with X having slightly better performance score than Y, the \texttt{slope} between X and Y acts as an indicator pointing to a direction in which to displace the data; i.e. away from Y and towards X.

While testing, HOW finds the nearest slope to every test case. The slope provides the exact magnitude and direction of displacements. Contrast sets are derived from these displacements. HOW offers a distinct advantage over the other CBR planners by limiting the displacements to very small regions (the displacements are never more than the separation between two clusters).

Although HOW manages to generate plans by localizing displacements to regions small enough to produce a statistically significant improvement, it needs to be noted that it fails to provide succinct summaries. Lack of succinctness makes it difficult draw generalizable conclusions about the test data. This is a trend commonly observed in most CBR systems. They tend to reason directly from a loaded training data instead of first summarizing the data into a model.

% However, not all domains come with reliable models. For instance, a model that encompasses all the intricate issues that may lead to defects in software would be very large and indeed rather complex. In addition to this, finding empirical data to validate such models can be hard to come by and also time consuming ~\cite{me09i,me09j}.

\subsection{DECISION TREES}

Trustworthy domain models are a popular alternatives to CBR planners. In our previous work, we have used executing source code as a "model" to check if our mutations to test suites minimize the test suite size while maximizing the number of statements covered~\cite{me09m,andrews07,andrews10}. However, we do not always have access to ready-to-use models. Further, finding empirical data to validate existing models can be hard.

Taking into account the wealth of data that is available to us, this issue can fortunately be circumvented by constructing a categorical model based on decision trees. Unlike the previous planners which use where, here we construct a decision tree using information gain as a metric to find ideal splits in the data. This allows us to identify the most informative attributes to select. A decision tree (hereafter referred to as DTREE) built in this fashion emulates a model built from the training data.

Using DTREE, the test cases can be categorized into one of the branches of the tree. Now, to generate the contrast sets, we determine (1) What \textit{current} branch does a test case fall in?; (2) What \textit{desired} branch would the test case have to move to?; (3) What are the deltas between \textit{current} and \textit{desired}? The last question can be answered by finding the deltas in branches of the decision tree that lead to \textit{desired} from \textit{current}. For full algorithmic details, refer to \fig{xtrees_bare}.

A tree structure such as DTREE is characterized by attributes such as size and depth. It is worth noting that these attributes have a profound impact on its performance. Our initial motivation for using DTREE was that they could serve as a medium for experts to reason about a data. The size of the tree, if too large, jeopardizes the readability of the solutions by increasing the complexity. We have, therefore, endeavored to reduce the size of the tree by pruning away irrelevant solutions that do not contribute to better solutions (refer to step-2 of \fig{xtrees_bare}).

DTREE offers a great number of benefits compared to the other methods that were discussed above. Firstly, it offers a visual medium for experts to identify and explore solutions spaces that are local to the problem. Secondly, its solutions are much more stable than other cluster specific and instance specific learners. This stability can be attributed to the consistency and general reproducibility of the tree structure. Thirdly, and perhaps most importantly, the tree summarizes the training data succinctly, making it easier for a business user to examine the solution sets.

\lstnewenvironment{code}[1]{
\lstset{
mathescape,
numbers=left,
numberstyle=\scriptsize,
stepnumber=1,
numbersep=0.5em,
xleftmargin=1em,
framextopmargin=2em,
framexbottommargin=2em,
showspaces=false,
showtabs=false,
showstringspaces=false,
tabsize=2,
% Basic
basicstyle=\ttfamily\scriptsize,
backgroundcolor=\color{Background},
language=Python,
% Comments
commentstyle=\color{Comments}\slshape,
% Strings
stringstyle=\color{Strings},
morecomment=[s][\color{Strings}]{"""}{"""},
morecomment=[s][\color{Strings}]{'''}{'''},
% keywords
morekeywords={[1]import,from,class,def,for,while,if,is,in,elif,else,not,and,or,print,break,continue,return,True,False,None,access,as,,del,except,exec,finally,global,import,lambda,pass,print,raise,try,assert, dot, norm, zip, sorted},
keywordstyle={[1]\color{Code}\bfseries},
% additional keywords
morekeywords={[3]fastmap,Slope,bPruning,clister,train,leafs,weightedFeatures,HOW,envy, score,kNN,contrastSet,exemplar,Prune, FSel, exemplar,nearestSlope,dist,displace,geometry,splitAcross2Points,leaves,How,nearest,bPruning,Stats,divide,recurse,weight1,project, furthest, split,WHERE,clusterer, getContours,envied, fWeight, nearestContour, projection, mutate, HERE, knn},
keywordstyle={[3]\color{Keywords}\bfseries},
morekeywords={[2]@invari},
keywordstyle={[2]\color{Decorators}\slshape},
emph={self},
emphstyle={\color{self}\slshape},
firstnumber=last
%
}}{}

### 2016-02-17 tex/algos.tex
Initial ShareLaTeX Import

	\begin{figure} 
	\begin{shaded}
  ~\hrule~

 {\bf Figure 7.A: Measuring Variability}

	For  continuous and discrete values,
the {\em variability} can be measured using standard deviation $\sigma$ or entropy $e$.
Note that \mbox{$\sigma = \sqrt{\sum^n_{i=1} \left( \frac{(x_i - \bar{x})^2}{n-1}\right)}$}
where $\bar{x}$ is the mean of  numeric features $x_1,x_2,..x_n$.
Also \mbox{$e = - \sum^n_{i=1} p_i \mathit{log}_2(p_i)$} 
for  $n$ discrete values  at frequency
$f_1,f_2,.. f_n$ for \mbox{$N = \sum^n_{i=1} f_i$} and $p_i = f_i/N$.

  ~\hrule~

	{\bf Figure 7.B: Measuring distance}

	We use
		 Aha et al.'s standard Euclidean distance measure~\cite{aha91}.  For $F$ independent features, the measure returns   $d(X,Y)=\sqrt{\sum_i^F  w_i\Delta(X_i,Y_i)^2}$. Here, $w_i$ is a weight term for each feature
		 (usually set to 1).
			Within $\Delta$, if   $X_i,Y_i$ are both missing  values, then  return 1.
			Otherwise, replace any  missing items with values that maximizes the following.
			For numerics, $\Delta$ normalizes $X_i,Y_i$ (to the range 0,1 for min,max) then
			returns 
			$X_i - Y_i$. For discrete variables, $\Delta$ returns 0,1 if $X_i,Y_i$ are the
			same,different (respectively).

 ~\hrule~

	{\bf Figure 7.C: Top-down Clustering with WHERE}

				WHERE  divides data into  groups of size $\alpha=\sqrt{N}$. 
		Using this measure, WHERE runs as follows:
		\begin{enumerate}[leftmargin=3mm]
			\item Find   two   distance cases,  $X,Y$
			by picking any case $W$ at random, then setting $X$ to its most
			distant case, then setting $Y$ to the case most distant from
			$X$
			(which requires only $O(2N)$ comparisons).
			\item Project each case $Z$
			to position $x$ on a    lines running from $X$ to $Y$: if $a,b$  are distances  $Z$ to $X,Y$  then  $x = (a^2+c^2 - b^2)/(2ac)$.
			\item Split the data at the median $x$ value of all cases.
			\item For   splits larger than  $\alpha=\sqrt{N}$, recurse from step1.
		\end{enumerate}
 In terms of related work,
	  the above is similar in approach to Boley's PDDP algorithm~\cite{boley98}, but PDDP requires an $O(N^2)$ calculation
	  at each recursive level to find the PCA principle component. Our method, on the other hand,
	  performs the same task with only $O(2N)$ distance calculations 	using the 
	  FASTMAP heuristic~\cite{Faloutsos1995} shown in step1. Platt~\cite{platt05} notes that FASTMAP is a  Nystr\"om approximation to the first component of PCA.

	   ~\hrule~

	{\bf Figure 7.D: Top-down division with Decision Trees}

	Find a split in the values of  independent features that most reduces the variability
  of the dependent feature (measured using \fig{where}.A). Construct a standard decision tree using these splits.

	 ~\hrule~

	 	{\bf Figure 7.E: Finding the most informative rows}

Discretize all numeric features using the Fayyad-Iranni discretizer~\cite{FayIra93Multi}
(divide numeric columns into bins $B_i$, each of which  select for the fewest cluster ids).
Let feature $F$ have bins $B_i$, each of which contains $n_i$ rows
and 
let each bin $B_i$ have entropy $e_i$ computed from the frequency of clusters seen in that bin (computed from \fig{where}.A).
Cull the the features as per~\cite{papa13}; i.e. just use the $\beta=33\%$ most informative features
where  the   value of  feature $F$ is $\sum_i e_i\frac{n_i}{N}$ ($N$ is the number of rows).

~\hrule~

	  \end{shaded}
		\caption{Some algorithms used in this paper.}\label{fig:where}
		\end{figure}

### 2016-02-19 Manuscript.tex
Updates from ShareLaTeX

@@
        {llayman@fc-md.umd.edu}
-\setlength{\columnsep}{7mm}
-}
+\setlength{\columnsep}{7mm}}
 \maketitle
@@
 
-Business users now demand more insightful analytics; specifically, tools that generate ``plans''-- specific suggestions on  what to change  in order to improve  the predicted values.
-
-This paper proposes XTREE, a planner for software projects.
-XTREE receives tables of data with independent features and a corresponding weighted class which indicates the quality (``bad'' or ``better'') of each row in the table.
-Plans are edits to  the rows which ensures the changed row is more likely to be of a ``better'' quality.
-XTREE learns those plans by building a decision tree across the data, then reporting the differences in the branches from some current branch to another desired branch.
-Using data from 11 software projects,  XTREE can find better plans compared to three
-alternate methods. Those plans have lead to improvements with a median size of (56\%, 28\%) and   largest size of (60\%, 77\%) in (defect counts, runtimes), respectively.
 \end{abstract}
@@
 
-Due to the continuous change undergone by modern software systems, developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
-These anti-patterns are generally viewed as poor solutions to recurring design problems.
-In addition to there being anecdotal evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{}\cite{}.
+When modern software systems are developed, their design changes continuously.
+Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
+These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system.
+In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{}\cite{}\cite{}.
 Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{}\cite{}.
@@
 
-This is quite similar to the concerns voiced by several business users, who are demanding tools that support business-level interpretations of their data.
+Tools for detecting code smells have been the focus of several researchers lately, \cite{}.
+One popular analysis technique is the use of static code analysis and code quality metrics\cite{}.
+While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers in planning actions to improve the quality of source code.
+Similar to the concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
+
 At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
 Panelists commented  ``prediction is all well and good, but what about decision making?''.
-Note that these panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
-So:
+The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
+We call such tools \textit{planners}.
+In the context of detecting code smells, the planners are designed to address the following issues.
 
@@
 \item
-Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to generate \textit{plans} to remove the bad smells
+Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to develop tools that can be used to reason about the code smells. 
+\item
+In addition to detecting anti-patterns and code smells, tools must be capable of having a contextual understanding of the data to trigger only the most useful refactoring options.
+\item Tools must provide a visual medium for developer to gauge the nature of refactoring options, thereby discarding bad solutions.
 \ei
-In response to this , we propose a novel {\em planning} method called XTREE for learning changes to a software system such that its performance ``improves'', according to some quantitative measure.
-This paper uses XTREE to reduce  the expected value of the defects in Jureczko et al.'s JAVA systems~\cite{jureczko10};
-
-The contributions of this paper are (1)  the new XTREE algorithm and (2) an evaluation strategy that shows the algorithm performing significantly ``better'' than  the  planner  proposed in our prior   work~\cite{me12c}.
-In this paper, we focus on three research questions:
+
+In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively.
+As a measure of source code quality we take into account the number of classes with defects in them.
+To this end, we use defect count data set by Jureczko et al.
+for JAVA systems~\cite{jureczko10}.
+
+The contributions of this paper are (1) the new XTREE planning algorithm and (2) an evaluation strategy that shows the tool performing significantly ``better'' than  the  planner  proposed in our prior work~\cite{me12c}.
+
+The work on XTREE is motivated by the following three research questions:
 
 \bi
-\item \textbf{RQ1: Effective: Can a planner suggest plans to improve the quality of a defective class?}
-\item \textbf{RQ2: Succinct: Are the plans suggested by a planner succinct?}
-\item \textbf{RQ3: Insightful: Do the plans tell us thing we do not expect?}
+\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reduce the number of defects.
+Specifically, we use XTREE to suggest valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
+\item \textbf{RQ2:} \textit{Are the plans suggested by a planner succinct?} The second question aims to measure the number of refactorings required 
+\item \textbf{RQ3:} \textit{Insightful: Do the plans tell us thing we do not expect?}
 \ei
 
-The rest of this paper  describes our data, our planners, and the experiment that
-ranks XTREE against alternate approaches.  This is followed by notes on related work
-and validity.
-To enable reproducibility, all scripts and data used in this 
-study are available online at \url{http://git.io/vG3DG}.
+The rest of this paper  describes our data, our planners, and the experiment that ranks XTREE against alternate approaches.
+This is followed by notes on related work and validity.
+To enable reproducibility, all scripts and data used in this study are available online at \url{http://git.io/vG3DG}.
 
@@
 
-Other studies aimed at understanding the impact of anti-patterns on program comprehension \cite{1} showed that the presence of an anti-pattern in the source code does not decrease the developers’ performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{1}\cite{61}.
+Other studies aimed at understanding the impact of anti-patterns on program comprehension \cite{1} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{1}\cite{61}.
 While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern.
@@
 \section{Evaluation Strategy}
+\label{sect:eval}
 \input{tex/prediction_stats.tex}
@@
 \centering
-\includegraphics[width=\linewidth]{figs/Deltas-both.eps}
+\includegraphics[width=\linewidth]{figs/PercentChanges.png}
 \caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
 \end{figure*}
+
+% \begin{figure*}
+% \centering
+% \caption{My caption}
+% \label{my-label}
+% \begin{tabular}{lllllllllll}
+%  & CBO             & LCOM           & LOC & AMC & WMC & NPM    & CAM & RFC         & CA & AVG\_CC \\
+%  & 4               & 3              & 3   & 3   & 2   & 2      & 2   & 2           & 1  & 1       \\
+%  & Max: 9 \cite\{} & Min: 0, Max: 2 & *   & *   & 20  & Min: 1 & *   & 40 \cite\{} & *  & *      
+% \end{tabular}
+% \end{figure*}
 

### 2016-02-28 Manuscript.tex
Updates from ShareLaTeX

@@
 \newcommand{\fig}[1]{Figure~\ref{fig:#1}}
-\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
+\newcommand{\eq}[1]{Equation~\ref{ 0eq:#1}}
 %\setlist{nolistsep,leftmargin=5mm}
@@
 
-\title{How to Learn Useful Changes to Software Projects\\(to Assist Refactoring)}
+\title{Learning Which Bad Smells are 
+Real Stinkers?}
 \numberofauthors{4} 
@@
 
-When modern software systems are developed, their design changes continuously.
+When modern software systems are developed, their design 
+changes continuously.
 Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
 These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system.
-In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{}\cite{}\cite{}.
-Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{}\cite{}.
+In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{khomh12}\cite{li07}\cite{fab15_1}.
+Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{fab15_1}\cite{chatzigeorgiou10}\cite{arcoverde11}.
 This makes early detection of these code smells a necessity to maintain sufficient quality in software systems.
-This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed by (i) identifying potential anti-patterns; (ii) generating reliable refactoring options, and (iii) critiquing bad refactoring options.
-
-Tools for detecting code smells have been the focus of several researchers lately, \cite{}.
-One popular analysis technique is the use of static code analysis and code quality metrics\cite{}.
-While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers in planning actions to improve the quality of source code.
-Similar to the concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
+This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed not only by identifying potential anti-patterns, but also by (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
+
+Tools for detecting code smells have been the focus of several researchers lately, \cite{Tufano2015}\cite{boussaa13}.
+There are two  popular techniques for detecting smells: (1) by the making use of static code analysis and code quality metrics\cite{tsantalis09}; (2) Analysing software change history\cite{moha2010}.
+While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers with planning actions to improve the quality of source code.
+Similar concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
 
@@
 We call such tools \textit{planners}.
-In the context of detecting code smells, the planners are designed to address the following issues.
+In the context of detecting code smells, the planners are required to address the following issues.
 
@@
 for JAVA systems~\cite{jureczko10}.
+In addition to defects, the aforementioned data set measures a list of static code attributes associated with each class.
+These code attributes allow us to reason about the measurable properties of anti-patterns.
 
@@
 \bi
-\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reduce the number of defects.
-Specifically, we use XTREE to suggest valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
-\item \textbf{RQ2:} \textit{Are the plans suggested by a planner succinct?} The second question aims to measure the number of refactorings required 
-\item \textbf{RQ3:} \textit{Insightful: Do the plans tell us thing we do not expect?}
+\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reducing the number of defects.
+Specifically, we use XTREE to suggest a number of valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
+In order to ensure the changes are feasible, XTREE generate succinct plans.
+
+\item \textbf{RQ2:} \textit{Can a planner be used to critique bad refactoring options?} In addition to being able to suggesting changes, it's beneficial for planners to be able to quantitatively comment on the quality of certain refactoring options, thereby preventing unnecessary refactoring.
+
 \ei
@@
 
-\section{Preliminaries}\label{sect:prelim}
-
-\subsection{Code Smells?}
-
-Anti-patterns are poor solutions and recurring design problems that are unwillingly introduced by developers while designing and implementing large object-oriented systems.
-The impact of these code smells have been subjected to several empirical studies.
-As a result of the, the identification of these anti patterns has been receiving more attention from both researchers and practitioners.
-
-During lifecycle, a software system undergoes continuous changes in order to keep up with the 
-business level demands \cite{36}.
-Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{24}.
-
-Fowler \cite{24} and Brown et al.
-\cite{13} defined a catalogue of more than 30 anti-patterns.
-For each anti-pattern, they reported the definition of the anti-pattern as well as specific refactoring operations aimed at removing it in order to improve the quality of source code.
-
-To illustrate this, consider for example developers adding several responsibilities to a class rather that including them in separate classes.
-This results in a class that grows rapidly, and when there are further added responsibilities, the classes grow and breed.
-Ultimately, the class becomes too complex and its quality deteriorates.
-Such classes are know as Large Class \cite{24} and could be the cause of a design problem in source code.
-
-For a while now, there have been anecdotal evidence that design problems (such as anti-patterns) negatively impacts software comprehension and maintenance.
-However, in the past few years, anti-patterns have been the subject of several empirical studies aiming at empirically analyzing their impact on software maintainability.
-From these studies, we now have empirical evidence that code containing code smells or participating in anti-patterns is significantly more change-prone than ``clean'' code \cite{31}.
-Also, code participating in anti-patterns has a higher fault-proneness than the rest of the system code \cite{31}\cite{37}.
-
-Other studies aimed at understanding the impact of anti-patterns on program comprehension \cite{1} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{1}\cite{61}.
+\section{Motivation}\label{sect:prelim}
+
+\input{tex/jur_char.tex}
+
+The thesis of this paper is that the XTREE tool can stop developers
+wasting time on useless software refactorings. Before presenting experiments
+that supports that claim, this section motivates the need for 
+such tools.
+
+In many fields,  practitioners hang on to persistent beliefs, even
+if the evidence for those beliefs is not strong.
+For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10).
+Even when the evidence for or against a treatment or intervention is clear, medical providers and patients may not accept it~\cite{aschwanden10}.
+Aschwanden warns that {\em cognitive biases} such as confirmation bias (the tendency to look for evidence that supports what you already know and to ignore the rest) influence how we process information~\cite{aschwanden15}.
+
+As in medicine, so too in software engineering.
+There is growing evidence that amongst programmers,
+many of their  widely-held beliefs are inaccurate and misleading. 
+According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects.
+They comment ``past experiences were taken into account without much consideration for their context''~\cite{passos11}.
+J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer a similar warning.
+They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.
+
+To support the last paragraph, we offer two
+specific examples
+were  widely-held views in SE that are   now questionable, given recent results. 
+Firstly, Nagappan et al. analyzed~\cite{mei15} 
+how the {\tt goto} statement was used  in 11,000 Github repositories, 
+they found that developers
+rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
+memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
+
+Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
+the ``phase delay'' truism that {\em fixing a requirements error  during
+coding can be thousands of times more expensive than recognizing and fixing that error
+as soon as it arises}. Boehm and Basili comment that this phase delay effect
+``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
+After reading the literature, and surveying academics
+and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
+this effect seems now heavily mitigated by $21^{st}$ software development languages,
+tools and development practices.  When we mined the 171 software projects shepherded
+by the SEI in the period 2006-2014, we found that phase delay results
+in only minor changes in the 
+time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.
+
+These two studies are symptomatic
+of a widespread malaise in software engineering.
+Devanbu et al.  examined responses from 564 Microsoft software developers from around
+the world, They found that  `` (a)~programmers do indeed have very
+strong beliefs on certain topics; (b)~their beliefs are primarily formed
+based on personal experience, rather than on findings in empirical
+research; (c)~beliefs can vary with each project, but do not necessarily
+correspond with actual evidence in that project''~\cite{prem16}.
+Devanbu et al. further  comment that ``programmers give personal experience
+as the strongest influence in forming their opinions''. This is a troubling
+result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
+careers.
+
+One area where poorly-justifed beliefs can waste valuable project resources is
+refactoring in order to remove code bad smells.
+
+\subsection{Need for Planning}
+During lifecycle, a software system undergoes continuous changes in order to keep up with the business level demands \cite{lehman79}.
+Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{fowler09}.
+Code smells are symptomatic of design issues in codes like code understandability and code changeability.
+If left unaddressed, these can lead to a variety of maintenance problems and faults in software systems \cite{fowler09}\cite{marinescu06}.
+Fowler \cite{fowler09} and Brown et al.
+\cite{brown98} defined a catalogue of more than 30 anti-patterns.
+It is a common notion that the critiques motivated by anti-patterns and code smells would be more comprehensible to developers because many code smells are associated with conventional refactoring operations.
+
+The choice of refactoring operations to address the code smells arise from strongly held beliefs of developers on these topics, and their beliefs in turn are motivated by personal experiences.
+
+The distinction between developers' beliefs and empirical findings is particularly profound in the case of code smells.
+For instance, \cite{abbes11} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{abbes11}\cite{yama13}.
 While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern.
-In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{3}\cite{17}.
-
-All these findings suggest that code smells and anti-pattern need to be carefully detected and
-monitored and, whenever necessary, refactoring operations should be planned and performed to deal with them.
-Unfortunately, the identification and the correction of design flaws in large and non-trivial software systems can be very challenging.
-These observations call for \textit{planners} capable of supporting the software engineer by (i) identifying anti-patterns, and (ii) designing and applying a refactoring solution to remove them.
-Such planners, are the focus of the current paper.
+In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{arcoverde11}\cite{chatzigeorgiou10}.
+Additionally, Yamashita and Moonen\cite{} observed that several developers did not take conscious actions to alleviate bad smells that were present in the code.
+
+Closer inspection revealed that this is not due to the lack of relevance of code smells, but a number of other factors, like: (1) lack of awareness; (2) More importantly, lack of developer friendly tools \cite{}.
+The study goes on to show that these developers are in need of better support tools that assist them during the software evolution cycle.
+Specifically, developers needed tools to (1) be developer friendly; (2) be operable in real-time; and (3) be capable of identifying problematic areas using error prediction.
+
+Therefore to assist developers in their refactoring efforts by providing them with reliable analytics it is useful to have automatic methods that can augment human expertise (or replace them if such experts are absent) in addition to offering a second opinion.
+Accordingly, the rest of this paper defines and evaluates automatic methods to address the above issues.
 
 \subsection{What is a ``Plan''?}
-
-Our planners   use tables of data with independent features and a dependent class
-feature. Classes have weights  that  indicate  what rows are ``bad'' or ``better''. Plans  change   a row such that it is more likely to be ``better''. 
-Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to
-  adjust   feature $Z_j$:
+To address the issues discussed in the previous section, we conjecture that the planners must generate valid and meaningful refactoring operations, we refer to these as \textit{plans}.
+These plans are synthesized by the planner and are thus unique in that they are agnostic to any prior notions of the developers in addition to them being local to the software project being refactored.
+
+To do this, our planners use tables of data with independent features and a dependent class feature.
+Classes have weights  that  indicate  what rows are ``bad'' or ``better''.
+Plans  change   a row such that it is more likely to be ``better''.
+Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to adjust   feature $Z_j$:
 
@@
 
-For example, to simplify a  large bug-prone  method, our planners might suggest
-to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across
-two simpler functions).
+For example, to simplify a  large bug-prone  method, our planners might suggest to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across two simpler functions).
 
 Additionally, if a developer were to suggest a specific refactoring option, our planners can query the option and report on its possible usability.
-Note that we make no assumption that a plan mentions every feature
-(so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than
-plan2).
-
-\subsection{From Detection to Planning}\label{sect:p2p}
-
-This paper is about the next step {\em after} detection of bad smells.
-Suppose developers are presented with  a list of anti-patterns and they do not like what they see.
-This user may then ask a {\em planning}  question; i.e.
-``what can we change to do better than that?''.
-
-Before exploring automatic methods to answer the planning question, we first comment
-on two manual methods.
-
-One way to propose changes to a project would be to   ask some smart experienced
-person for their opinion on the impact of these anti-patterns and the possible refactoring options. Most advices offered by these developers are based on experiences, sometimes
-such advice is an effective strategy and sometimes it is not.
-
-According to Passos et al.~\cite{passos11},  developers
-may  assume that the lessons they learn from a few past
-projects are general to 
-all their future projects. They comment ``past experiences were taken into account without 
-much consideration for their context''~\cite{passos11}.  
-J{\o}rgensen \& Gruschke~\cite{jorgensen09} offer  a similar warning. They report that the 
-supposed software engineering  ``Gurus'' rarely use lessons
-from past projects to improve their future reasoning and that such poor
-past advice can be detrimental to new projects.~\cite{jorgensen09}.
-Accordingly, we   propose a ``trust, but verify'' approach.
-After a software guru offers some sage wisdom,  it is wise to ask some other oracle 
-if there are any better options
-(just as a sanity check).
-The rest of this paper discusses some methods to build automatic oracles 
-to implement that   sanity check.
-
-Another way  to find  changes to a project
-might be to rely
-on the peer review processes used by the 
-SE research community. This approach would propose changes to software
-projects that concur with internationally accept best practices. 
-There are two
-problems with that approach. Firstly, given the rapid pace of change in software
-engineering, we may be asking questions for which there is no current widely
-accepted concept of
-``best practice''.
-
-Secondly, given the diversity of SE products and practices
-and personnel, it may well be that the current project being discussed is 
-substantively different to prior work. 
-Numerous recent {\em local learning} results compare (1) single models
-learned from all available data to (2) multiple models learned from clusters within the data~\cite{betten14,yang11,yang13,minku13,me12d,me11m,posnett11}.
-A repeated result in those studies is that the local models generated better effort
-and defect predictions (better median results,
-lower variance in the predictions). \tion{surprise} of this  paper offers yet another locality result:
-\bi
-\item
-One standard rule in the literature
-is that it is useful to implement modules such that they are internally cohesive (use
-much of their own local methods) while being loosely coupled with other classes~\cite{Dhama199565}.
-\item
-While that may be true in general, for particular classes other changes may be more important
-(later in this paper, we show one set of results where that is indeed the case).
-\ei
-In summary, 
-it is useful to have automatic methods to recommend changes. Such
-methods can fill in for human expertise (if such experts are absent) or 
-to offer a second opinion.
-Also, prior to making automatic recommendations, it is wise to first stratify the data
-(clump it into related examples) then generate advice specific to each clump.
-Accordingly, the rest of this paper defines and evaluates
-automatic methods to find plans from
-  $N$ examples divided  into  many clumps.
-
-\input{tex/jur_char.tex}
-
-%XXX end
-
-%  \begin{figure}[!t]
-%  \small
-%  \begin{center}
-%  \begin{tabular}{r|rr}
-%  data set & cases & \% defective\\\hline
-%   ant &947& 22\\
-%   camel& 1819& 19\\
-%  jedit& 1257& 2\\
-%  ivy &352& 11\\
-%  log4j& 244 &92\\
-%  lucebe &442 &59\\
-%  poi& 936 &64\\
-%  synapse &379 &34\\
-%  velocity& 410& 34\\
-%  xalan& 2411& 99\\
-%  xerces &1055& 74
-%  \end{tabular}
-%  \end{center}
-%  \caption{ Jureczko data: columns in the format of \fig{ck}.}\label{fig:jd}
-%  \end{figure}
-
-%  \begin{figure}[!t]
-% \scriptsize
-% \begin{tabular}{llllll}
-%   \hline
-%   \rowcolor{lightgray}
-% Project & Domain & Lang. & LOC & Features & Config\\\hline
-% BDBC: Berkeley DB   & Database & C & 219,811 & 18 & 2560\\
-% BDBJ: Berkeley DB   & Database & Java & 42,596 & 32  & 400\\
-% Apache & Web Server & C & 230,277 & 9 & 192\\
-% SQLite & Database & C & 312,625 & 39 & 3,932,160\\
-% LLVM & Compiler & C++ & 47,549 & 11 & 1024\\
-% x264 & Video Enc. & C& 45,743 & 16 & 1152\\\hline
-% \end{tabular}
-
-% \caption{Siegmund data.}\label{fig:cpm}
-% \end{figure}
-
-% \begin{figure}[!t]
-% \includegraphics[width=0.9\linewidth]{figs/BDBC.eps}
-% \caption{ Berkeley database feature model   (``C'' version). }\label{fig:bdbc}
-% \end{figure}
+Note that we make no assumption that a plan mentions every feature (so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than plan2).
 
@@
 The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc.
-For details on the Jureczko data, see  \fig{ck} and \fig{jd}.
-All the planning methods of this paper reflect on the numeric value of the raw defect counts.
+For details on the Jureczko data, see  \fig{ck}.
+All the planning methods of this paper reflect on the numeric value of the raw defect counts to estimate the quality of the class.
 The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.
@@
 
-% The Siegmund data, described in \fig{cpm},  records  the runtimes of compiled systems. To obtain the  data, Siegmund et al. perturbed
-% the configuration parameters in the Makefiles of six systems: Apache, SQLite, LLVM, x264 and two versions of the
-% Berkeley database (one written in ``C'' and one in Java).
-Then, the performance was measured using standard benchmarking tools (delivered by \textsc{Oracle} for Berkley data sets and other popular tools such as \textsc{autobench} and \textsc{httpref} for the rest)~\cite{sven12}.
-It's worth noting that in each of the above data sets, several features are interdependent, this is expressed using a \textit{feature model}.
-% \fig{bdbc} shows an example of such a feature model defining valid combinations of settings in the Berkley database (``C'' version).
-These feature models were used by Siegmund et al.
-to ensure all their perturbations are valid
-% (we will use the same models to cull invalid plans).
+\section{Oracles}
+\label{sect:eval}
+\input{tex/prediction_stats.tex}
+
+As highlighted previously, there are several detection techniques for anti-patterns, yet these fail to address the issue of checking the proposed changes for their efficacy.
+Much of the refactoring needs to evaluated for their usefulness.
+To this end, there needs to be an oracle, a procedure that can be employed to distinguish between good and bad options.
+The dearth oracles is not unique to this domain, for instance a recent survey by Barr et al \cite{barr15}
+
+Our strategy (discussed in greater detail below) divides this data into a training and a testing set.
+The purpose of training set is twofold: (1) We apply a data miner to construct an oracle that acts as a quality predictor; (2) We build the planners to reason across the data.
+
+Next, we employ the planners to generate plans to improve the quality of defective classes in the test set.
+In addition to the plans obtained from our planners we may have some common refactoring options that the developers tend to use.
+The oracle built previously can be used to assess the the changed examples and this can be used to compare refactoring operations.
+
+It's worth noting that this approach depends on having effective oracles for assessing the results.
+It proved to be more complicated to commission the Jureczko data sets for this study.
+For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could be salvaged using the techniques discussed in this section.
+
+\fig{j} shows our preliminary studies with the Jureczko data.
+Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}, this means that we are training on hundreds to thousands of classes and testing on smaller test suites).
+Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
+In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
+
+In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
+For such data, quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
+``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
+``pf''): the percent of non-fault classes that are {\em predicted} to be defective.
+
+As a preliminary study, we split the Jureczko  data into train and test groups.
+Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was built from the training data, then applied to the test data.
+The ``untuned'' columns of \fig{j} shows those results.
+We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
+For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
+With this threshold, however, none of our data sets were be suitable for this study.
+
+Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples and an under-sampling of majority-class examples.
+Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
+
+In summary, while we cannot trust predictors from some of our Jureczko data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+Accordingly, when this study explores the Jureczko data, we will use these five data sets.
+
+(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+
+\input{tex/algos}
 
@@
 
-This section describes XTREE (which we call Method4) and  three  
-alternate methods for learning plans.
+In this we describe the planners used in our study.
+First, we reintroduce a planner called CD which was an outcome of our previous work~\cite{me12c} which generates plans using a top-down bi-clustering method described in \fig{where}.C to recursively partition the data along the dimension that captures the greatest variability in the data.
+Then, we present XTREE, a novel planner introduced in this work, which uses a decision tree learner \fig{where}.D for planning.
+
+% This section describes the new XTREE planner and a planner called CD which was an outcome of our previous work~\cite{me12c}.
 XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
-  The other methods use the   top-down
-	bi-clustering method described in \fig{where}.C  which recursively divides the
-	data in two  using a dimension that captures the greatest variability in the data. 
-	We proposed Method 1 and 2   in 2012~\cite{me12c} while Methods 3 comes from research conducted earlier this year~\cite{krishna15}. 
-All  methods have the  properties proposed in \tion{prelim}: (1)~they are {\em local learners}
-that find   plans   particular to each test case; and (2)~they  are {\em trust-increasing}; 
-  i.e. the changed examples are moved {\em closer} to the training data.
+
+Both methods aim to address the issues highlighted in \tion{prelim}.
+To achieve this: (1)~they find plans, which can be use to generate reliable refactoring options, particular to each test case by reflecting on local data; and (2)~they offer a mechanism for critiquing bad refactoring options.
+
+% (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
 
@@
 
-Our  description of the methods adopts the following convention. All variables
-  set via  our engineering judgement  with Greek letters; e.g. $\alpha,\beta,\gamma,\omega$.
-  In this paper, we show our current settings to these variables produces useful
-  results. Elsewhere\cite{krall14,fu:ase15}, we are exploring tuning methods to 
-  find better settings but  we have nothing definitive yet to report
-  on auto-tuning planners.
+Our  description of the methods adopts the following convention.
+All variables set via  our engineering judgement  with Greek letters; e.g.
+$\alpha,\beta,\gamma,\omega$.
+In this paper, we show our current settings to these variables produces useful results.
+Elsewhere \cite{krall14,fu:ase15}, we are exploring tuning methods to find better settings but  we have nothing definitive yet to report on auto-tuning planners.
 
@@
 
- {\em Summary1:} Method1  computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be  (which we will call $C_j$).
-
-{\em Assumption1: } Large data sets can be adequately represented by a few dozen (or so) 
-centroids.
-
- {\em Details1:} Method1 clusters project data by reflecting on the independent variables, then
-  reports the delta between the cluster centroids. 
-  After clustering training data using the WHERE algorithm of \fig{where}.C, Method1
-replaces all clusters with a  centroid $C_i$ computed from the median value of each
-continuous/discrete feature. Then, it
-finds the closest centroid $C_j$ that has a better
-performance score. For Jureczko data, ``better'' means fewer defective examples while for the Seigmund data,
-``better'' means lower median runtimes for the examples in that cluster.
-Method1 then caches the  delta between the independent features between $C_i$ and $C_j$. For continuous
-features, this delta is $C_j - C_i$. For discrete values, this delta is the value of that feature
-in $C_j$. 
-Finally, for every test case, Method1 use the distance measure $d$ shown in \fig{where}.B to find
-the nearest centroid $C_i$.  It then proposes a plan for improving that test case
-that is the conjunction of all the deltas between $C_i$ and $C_j$.
-
-%\input{tex/pseudo-code-1.tex}
+% Method1 computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be (which we will call $C_j$).
+
+Large data sets can be adequately represented by a few dozen (or so) clusters, and meaningful inferences can be drawn from the centroids of these clusters.
+This method generates plans for improving faulty classes by looking at the difference between the cluster centroid that most closely resembles the defective class from the training data (denoted by $C_i$) and the cluster centroid that represents a better, non-defective, class (denoted by $C_j$).
+
+To do this, CD clusters training data based on the independent variables using the WHERE algorithm of \fig{where}.C and replaces all clusters with their centroids computed from the median value of each continuous/discrete feature.
+Then, for every cluster $C_i$ it finds the closest centroid $C_j$ that has a better performance score.
+For Jureczko data, ``better'' means fewer defective examples.
+CD then caches the delta between the independent features of $C_i$ and $C_j$.
+For continuous features, this delta is $C_j - C_i$.
+Finally, for every test case, CD uses the distance measure $d$ shown in \fig{where}.B to find the nearest centroid $C_i$.
+It then proposes a plan for improving that test case: the conjunction of all the deltas between $C_i$ and $C_j$.
 
 \input{tex/dtree.tex}
+
+\subsubsection{XTREE}
+
+A potential problem with CD is the {\em unsupervised} nature of the clustering algorithm (WHERE) that executes without knowledge of the target class.
+It would be beneficial for us to use {\em Supervised} methods that also reflect on the target class.
+
+XTREE was developed to address this problem.
+It uses a supervised decision tree algorithm of \fig{where}.D to categorize the data Next, XTREE devises plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
+That code asks three questions, the last of which returns the plan:
+
+\be
+\item What {\em current} branch does a test case fall in?
+\item What {\em desired} branch would the test case want to move to?
+\item What are the {\em deltas} between current and desired? 
+\ee
+
+In finding the answers to the above questions, XTREE can (a) Generate an wealth of plans (that can be potential refactoring operations); (b) Critique plans based on their location in the decision tree constructed.
+In addition to the above, XTREE also offers developers a visual medium to validate their proposals for refactoring, see ~\fig{xtree_samp}.
+
+\begin{figure}[btp!]
+\centering
+\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
+\caption{Sample plans from XTREE}\label{fig:xtree_samp}
+\end{figure}
+
+\section{Experiments}
+
+This section describes an experimental design (and results) for evaluating the above four methods. 
+\subsection{Experimental Design}
+
+\subsubsection{A Strategy for Evaluating Planners}
+
+Our experimental design is shown in \fig{design}. We divide the
+project data  into two disjoint sets {\em train} and {\em test}
+(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
+Next, from the train set, we build both a {\em planner} and
+ a {\em  predictor}.
+
+Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
+\bi
+\item Our {\em planners} will be XTREE and CD;
+\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.
+We use these
+data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
+\ei
+As for the {\em test} data, this is passed to the { predictor}
+to measure performance statistics related to effectiveness.
+
+If our { predictors} fail to perform effectively on the test data,
+then we cannot trust them to comment on our plans. Accordingly,
+if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
+we should not use some of the  Jureczko data.
+
+Else, we (1)~apply the { planner} to alter the {\em test} data;
+then (2)~apply the { predictor} to the altered data $test'$;
+then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
+\bi
+\item If $R  = 0\%$, this means  ``no change from baseline''; 
+\item If $R \gt 0\%$, this indicates ``
+
+improvement over the baseline'';
+\item If $R \lt 0\%$, this indicates ``optimization failure''.
+\ei
 
@@
 
-\subsubsection{XTREE}
-
-XTREE builds a decision tree,  then generates
-plans from the difference between two branches:
-(1)~the branch to where you are and (2)~the branch to where you want to be.
-
-One potential problem with Method 1 is the {\em unsupervised} nature of
-the clustering algorithm (WHERE) that  
-executes without knowledge of the target class.  {\em Supervised} methods assume that it is useful to also reflect on the target class.
-
-XTREE uses a supervised   decision tree algorithm of \fig{where}. D to divide the data.
-Next, XTREE builds plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
-That code asks three questions, the last of which returns the plan:
-\be
-\item
-What {\em current} branch does a test case fall in?
-\item What {\em desired} branch would the test case want to move to?
-\item What are the {\em deltas} between current and desired? 
-\ee
-
-\section{Evaluation Strategy}
-\label{sect:eval}
-\input{tex/prediction_stats.tex}
-
-Our evaluation strategy (discussed below) divides this data into a training and a testing set.
-From the training set we apply a data miner (to learn a quality predictor) and
-various planning methods (to learn different plans). Next, we try applying
-each of those plans to the  test set and ask the quality predictor to assess the changed examples.
-Finally, we say that the  ``best'' planner is the one that most reduces the predicted values
-in the changed examples.
-
-As mentioned in the last section,  this approach depends on having effective predictors for assessing the results.
-It proved to be  more complicated to commission the Jureczko data sets for this study.
-For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could
-be salvaged using the techniques discussed in this section.
-
-%   For the Siegmund data, this criteria was   relatively easy to achieve.
-%   The data in those data sets have a continuous class (runtime of the compiled system)
-%   so the performance of a quality predictor can  be measured in terms of  difference between the predicted runtime $p$ of test case items and their actual runtimes $a$ using  $s= (1 - \frac{abs(a - p)}{a})\times100\%$ ({\em higher} values are {\em better}).
-% This paper  explores six Siegmund configuration data sets:  Berkeley DB (Java and C versions), Apache, SQLite, LLVM, and
-%   x264. 
-%   As a preliminary study, we split that data into equal sized train:test groups
-%   and trained a Random Forest
-%   Regressor (from the SciKit learn kit~\cite{Pedregosa2012})   on one half, then applied to the other.
-After repeating this process 40 times, we achieved nearly perfect accuracy of $s=$\{98.45\%, 99.62\%, 94.86\%, 98.89\%, 98.60\%, 99.87\%\};
-%   i.e.  we are confident that the predictors from the Siegmund data can assess
-% our plans. (Aside: if the reader doubts that such high scores are achievable, we note that these scores are consistent with those achieved by predictors built by Siegmund et al.~\cite{sven12}.)
-
- \fig{j} shows our preliminary studies with the Jureczko   data.
-Given access to $V$ released
-versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as
-shown in \fig{j}, this means that we are training on hundreds to thousands
-of classes and testing on smaller test suites).
-Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
-In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
-
-In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
-For such data, quality of the predictor can be measured using
-(1) the  probability of detection (a.k.a. ``pd'' or recall):  the percent of faulty classes in
-the test data detected
-by the {\em predictor}; and (2) the 
-probability of false alarm (a.k.a. ``pf''): the percent of non-fault
-classes that are {\em predicted} to be defective.
-
-As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was
-built from the training data, then applied to the test data.
-The ``untuned'' columns of \fig{j} shows those results.
-We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
-For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
-With this threshold, however, none of our data sets were be suitable for this study.
-
-Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
-Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples and an under-sampling of majority-class examples.
-Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
-The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
-For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
-However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
-
-In summary, while we cannot trust predictors from some of our Jureczko data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-Accordingly, when this study explores the Jureczko data, we will use these five data sets.
-
-(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
-
-\section{Experiments}
-
-This section describes an experimental design (and results) for evaluating the above four methods. 
-\subsection{Experimental Design}
-
-\subsubsection{A Strategy for Evaluating Planners}
-
-Our experimental design is shown in \fig{design}. We divide the
-project data  into two disjoint sets {\em train} and {\em test}
-(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
-Next, from the train set, we build both a {\em planner} and
- a {\em  predictor}.
-
-Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
-\bi
-\item Our {\em planners} will be XTREE and CD;
-\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.
-We use these
-data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
-\ei
-As for the {\em test} data, this is passed to the { predictor}
-to measure performance statistics related to effectiveness.
-
-If our { predictors} fail to perform effectively on the test data,
-then we cannot trust them to comment on our plans. Accordingly,
-if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
-we should not use some of the  Jureczko data.
-
-Else, we (1)~apply the { planner} to alter the {\em test} data;
-then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
-\bi
-\item If $R  = 0\%$, this means  ``no change from baseline''; 
-\item If $R \gt 0\%$, this indicates ``
-
-improvement over the baseline'';
-\item If $R \lt 0\%$, this indicates ``optimization failure''.
-\ei
 \subsubsection{Statistical Methods}
- Our methods use some stochastic algorithms; e.g. WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and
-  XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan
-  (see \fig{xtrees_bare}). Hence, we report the $R$ values seen in 40 repeated runs,
-  with different random number seeds
-(we use 40 since that is  more than the 30 samples  needed
-to satisfy the central limit theorem).
-
-  To rank our methods using the results from these 40
-  repeats, we use the Scott-Knott test recommended by Mittas and  Angelis~\cite{mittas13}.
+Our methods use some stochastic algorithms; e.g.
+WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan (see \fig{xtrees_bare}).
+Hence, we report the $R$ values seen in 40 repeated runs, with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
+
+To rank our methods using the results from these 40 repeats, we use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}.
 
@@
 sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
-Then, we
-split $l$ into sub-lists $m,n$ in order to maximize the expected value of
- differences  in the observed performances
-before and after divisions. E.g. for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$:
- \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
+Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.
+E.g.
+for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$: \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
 
@@
 
-In this example table, the rows are  sorted on the median values of each method. Note that all the methods
-have   $R\gt0\%$; i.e. all these methods reduced the expected value of the performance score in that experiment
-while XTREE achieved the greatest reduction (of 62\% from the original value).
+In this example table, the rows are  sorted on the median values of each method.
+Note that all the methods have $R\gt0\%$; i.e.
+all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 62\% from the original value).
 
@@
 
-% For Siegmund data, we performed a k-fold cross validation.
-Given the relatively small sample sizes in Apache and BDBJ, we chose $k=2$.
-This, we reasoned, would maintain a sufficiently large test data size in order for R to be measured reliably.
-However, in 2-fold cross validation, the division of data tends to affect the outcome significantly.
-To counter this, we randomized the order of the data, training on one half while identifying treatment plans on the other, repeating the process 40 times as mentioned above.
-
-The Jureczko data, being temporal in nature, allows us
-to implement a validation procedure that ensures only past data is ever used to
-predict future values. Hence, in that data, we used the train/test sets shown in \fig{j}. 
+The Jureczko data, being temporal in nature, allows us to implement a validation procedure that ensures only past data is ever used to predict future values.
+Hence, in that data, we used the train/test sets shown in \fig{j}.
 (Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).
+
 \input{tex/results_jur.tex}
@@
 \subsection{Experimental Results}
+
+\begin{figure}[tb!]
+\centering
+\includegraphics[width=\linewidth]{figs/thresh.png}
+\caption{Thresholds used by DECOR to detect Spaghetti Code~\cite{moha10}.}\label{fig:thresh}
+\end{figure}
 
@@
 
-\begin{figure*}[!t]
+\begin{figure}[tbp!]
 \centering
@@
 \caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
-\end{figure*}
-
-% \begin{figure*}
-% \centering
-% \caption{My caption}
-% \label{my-label}
-% \begin{tabular}{lllllllllll}
-%  & CBO             & LCOM           & LOC & AMC & WMC & NPM    & CAM & RFC         & CA & AVG\_CC \\
-%  & 4               & 3              & 3   & 3   & 2   & 2      & 2   & 2           & 1  & 1       \\
-%  & Max: 9 \cite\{} & Min: 0, Max: 2 & *   & *   & 20  & Min: 1 & *   & 40 \cite\{} & *  & *      
-% \end{tabular}
-% \end{figure*}
-
-% \includegraphics[width=\linewidth]{figs/ant.png}
-
-\begin{figure*}[htbp!]
+\end{figure}
+
+\begin{figure*}[tbp!]
 \centering
-\includegraphics[width=\linewidth]{figs/changes.png}
+\includegraphics[width=0.8\linewidth]{figs/changes0.png}
 \caption{Suggested Changes}
@@
 
-\subsubsection{RQ1: Effectiveness: Can plans to improve the quality of a defective class?}
+\subsubsection{RQ1: Can a planner suggest plans to improve the quality ofa defective class?}
 
@@
 
-\subsubsection{RQ2: Succinctness: Are the plans suggested by a planner succinct?}
-
 \fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
-The left-hand-side plot of that figure reports results from one of the  Jureczko data sets ({\em lucene}) and the right-hand-side
-  shows a Siegmund data set ({\em BDBJ}).
-
-  In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e. the vertical bars in that plot
-  are {\em lower}). For example,
-  XTREE's plans were usually succinct --  in all  data sets,
- XTREE   changes around a fifth of the features (see \fig{types}).
-On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features 
- (observe the change frequencies as 
-high as 100\% for all features). 
-Method1's policy of ``change everything''  might be acceptable
-if this approach lead 
-to the most effective changes. However, in
-\fig{jur} and  \fig{conf1}, there is no evidence for this.
-
-An  interesting feature of \fig{changed} was that fewer things 
-were changed in the Seigmund data sets like {\em BDBJ} than in the Jureczko data set like
-{\em lucene}. 
-In turns out that this holds true across nearly all our data sets.
-\fig{types} summarizes all the change frequencies for all data sets. As with  \fig{changed},
-there are fewer changed features in the Seigmund data than in the defect prediction data. 
-One explanation for that is the nature of the features: the
-Jureczko data sets have continuous
-features while the 
-Seigmund data has binary independent features (where settings can be turned ``ON'' or ``OFF'').
-While the features of the Jureczko data sets are not constrained, those of Seigmund data sets are subject to several constrains (dictated by their feature models, see \fig{bdbc}).
-It is possible that planners in Method1, Method2, and Method3 find very few valid settings that can improve the performance scores, thereby making insignificant changes.
-As a result all methods, except XTREE, fail to optimize the runtime, as evidenced by the results in \fig{conf1}.
+The left-hand-side plot of that figure reports results from one of the Jureczko data sets ({\em lucene}) and the right-hand-side shows a Siegmund data set ({\em BDBJ}).
+
+In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e.
+the vertical bars in that plot are {\em lower}).
+For example, XTREE's plans were usually succinct --- in all  data sets, XTREE   changes around a fifth of the features (see \fig{types}).
+On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features (observe the change frequencies as high as 100\% for all features).
+Method1's policy of ``change everything''  might be acceptable if this approach lead to the most effective changes.
+However, in \fig{jur} there is no evidence for this.
 
@@
 
-\subsubsection{RQ3: Do the plans tell us thing we do not expect?}\label{sect:surprise}
+\subsubsection{RQ2: Can a planner be used to critique bad refactoring options?}\label{sect:surprise}
 
@@
 
-\subsection{Trusting the Changes}\label{sect:trust}
-   XTREE is evaluated by  comparing
-predicted performance scores before and after a planner makes changes to the feature values of an example:
-After making those
-changes, we may have a new example that has never been seen before. Therefore, it must be asked
-{\em ``can we trust the predictions made on such new examples?''}
-
-To answer this question, we note
-that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples
-(for a visualization of these clouds, see \fig{howxy}).
-\begin{figure}[!t]
-  \includegraphics[width=\linewidth]{figs/2d.eps} 
- % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
-\caption{Gray, red, green show (1) training examples, (2) test examples and 
-  (3) tests that have been altered by planners.
-  This figure uses axes generated from the first two components of a PCA analysis of all points. 
-}\label{fig:howxy}
-\end{figure}
-We should mistrust the predictions made by a  model   if it is being applied to examples  that are
-too far away from the
-training cloud.
-To test for ``too far'', we can run a data mining experiment that tests how well
-a model learned from the training data applies to the test data. Such experiments return some performance value.
-
-Note that predictions  about changes that  fall within the space of the training+test data, will be at least
-as accurate as the predictions of the original test data.
+% \subsection{Trusting the Changes}\label{sect:trust}
+%   XTREE is evaluated by  comparing
+% predicted performance scores before and after a planner makes changes to the feature values of an example:
+% After making those
+% changes, we may have a new example that has never been seen before. Therefore, it must be asked
+% {\em ``can we trust the predictions made on such new examples?''}
+
+% To answer this question, we note
+% that data miners explore two  ``clouds'' of data: (1) the cloud of training examples and (2) the  cloud   of test examples
+% (for a visualization of these clouds, see \fig{howxy}).
+% \begin{figure}[!t]
+%   \includegraphics[width=\linewidth]{figs/2d.eps} 
+%  % \includegraphics[width=1\linewidth]{figs/BDBC.eps}
+% \caption{Gray, red, green show (1) training examples, (2) test examples and 
+%   (3) tests that have been altered by planners.
+%   This figure uses axes generated from the first two components of a PCA analysis of all points. 
+% }\label{fig:howxy}
+% \end{figure}
+% We should mistrust the predictions made by a  model   if it is being applied to examples  that are
+% too far away from the
+% training cloud.
+% To test for ``too far'', we can run a data mining experiment that tests how well
+% a model learned from the training data applies to the test data. Such experiments return some performance value.
+
+% Note that predictions  about changes that  fall within the space of the training+test data, will be at least
+% as accurate as the predictions of the original test data.
 With this, we assert that the predictions for changes that move examples towards/away from the training data can be trusted more/less (respectively).
 
-Accordingly, we need {\em trust-increasing} planners to generate new examples {\em closer} to the
-training examples.  To see how this works, 
- \fig{howxy} is from the {\em ivy} data
-set (one of the Jureczko data sets used in this paper). It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
-changed  examples displaced after applying a plan (in green).
- Note that the  the   changed examples
-cases  (shown in green)  fall closer to the training cases (shown in gray) than
-the test cases (shown in red).
-
-In that green region of changed examples, our belief in the value of predictions
-will be just as much as, if not more than, our belief in the value of the predictions in the red region (that
-contains the original test data).
-This pattern of \fig{howxy} (where the new examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
-paper. Hence,  we can assert that
-predictors learned from these training examples have some authority in the regions
-containing the changes examples.
-
-That said, the above comes with some important caveats:
-\bi
-\item 
-The quality of the prediction depends on the nature of the training data.
+% Accordingly, we need {\em trust-increasing} planners to generate new examples {\em closer} to the
+% training examples.  To see how this works, 
+%  \fig{howxy} is from the {\em ivy} data
+% set (one of the Jureczko data sets used in this paper). It shows: (1)~the training examples in gray, (2)~the test examples in red, and (3)~the
+% changed  examples displaced after applying a plan (in green).
+%  Note that the  the   changed examples
+% cases  (shown in green)  fall closer to the training cases (shown in gray) than
+% the test cases (shown in red).
+
+% In that green region of changed examples, our belief in the value of predictions
+% will be just as much as, if not more than, our belief in the value of the predictions in the red region (that
+% contains the original test data).
+% This pattern of \fig{howxy} (where the new examples are found closer to  the training cases than the test cases) has been observed in all the other data sets studied in this
+% paper. Hence,  we can assert that
+% predictors learned from these training examples have some authority in the regions
+% containing the changes examples.
+
+% That said, the above comes with some important caveats:
+% \bi
+% \item 
+% The quality of the prediction depends on the nature of the training data.
 Thus, we strongly recommend that both the data set and the predictor be assessed prior to planning.
@@
 We tackle this issue in detail in \tion{tesd}.
-\item
-Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
-\item
-Where possible, planners should be assessed via some external
-oracle that can accurately assess new examples. For an example of that kind of analysis,
-see  \tion{coc}.
-\ei
+% \item
+% Planners should be designed to be {\em trust increasing}. We list four such planning methods in \tion{planners}.
+% \item
+% Where possible, planners should be assessed via some external
+% oracle that can accurately assess new examples. For an example of that kind of analysis,
+% see  \tion{coc}.
+% \ei
 
@@
 For building the defect predictors in this study, we elected
-to use  Random Forest  and Random Forest Regressors .
-We chose this approach,  based on its reputation for having the better  performance of 
-21 other learners for defect prediction~\cite{lessmann}.
-Data mining is a
-large and active field and any single study can only use a small
-subset of the known classification algorithms.
+to use  Random Forests.
+We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.
+Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.
 
@@
 work should vary in order to test the internal validity of our conclusions:
+
 \bi
-\item All our planners divide data into groups of size $\alpha=\sqrt{N}$;
-\item Method2 used the top $\beta=33\%$ most informative features (ranked using INFOGAIN);
-\item Method3 uses  $\omega=0.5$ to find its deltas;
+\item The clustering scheme, WHERE, divides data into groups of size $\alpha=\sqrt{N}$;
 \item XTREE considered a sibling useful if it's score was less than $\gamma=0.5$ times the mean score of the current leaf.

### 2016-03-03 Manuscript.tex
Updates from ShareLaTeX

@@
 
-\title{Learning Which Bad Smells are 
-Real Stinkers?}
-\numberofauthors{4} 
+\title{ How to Recognize Useless ``Bad Smells''}
+\numberofauthors{2} 
 \author{  
 \alignauthor
-Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
+Rahul~Krishna, Tim Menzies, Xipeng~Shen \\
        \affaddr{CS, NC State, USA}\\
-       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
-\alignauthor
-Andrian Marcus \\
-       \affaddr{CS, UT Dallas  } \\ 
-       \affaddr{Texas, USA}\\ 
-       {amarcus@utdallas.edu}
-       \alignauthor
-Naveen  \\ Lekkalapudi\\
- \affaddr{Bellhops } \\ 
-       \affaddr{Tennessee, USA}\\ 
-       {navek91@gmail.com}
+       {\{i.m.ralk,~tim.menzies,~xipengshen\}@gmail.com}
 \alignauthor
@@
 \begin{abstract}
-
+Developers, text books, and tools
+disagree on what bad smells should be refactored.
+While it useful to fix  some bad smells, others are
+ can be ignored.
+How to distinguish useful from useless bad smells?
+
+We propose RANK, a tool that judges how badly
+a project needs a particular   refactoring.
+RANK is tuned using the historical data of the current project.
+Using RANK, we   show that the range of truly bad
+``bad smells'' changes from project to project. Hence,
+before refactoring, it is important to run a critique tool
+like RANK to stop work on useless refactorings.
+Further, using RANK, it is possible to show that many
+of the bad smells endorsed by standard texts or tools
+are not useful for particular projects. 
+Based on these results, we endorse  
+guiding refactoring via bad smells. However,  
+it is vital to RANK the bad smells defined for a 
+particular project.
+
+\vspace{3mm}
+\noindent
+{\bf Categories/Subject Descriptors:} D.2 [Software Engineering]; I.2 [Artificial Intelligence];
+
+\noindent
+{\bf Keywords:} Bad smells, refactoring,
+performance prediction,  decision trees 
 \end{abstract}
 
-\keywords{Defect prediction; Configuration; Planning; Decision trees.}
 \section{Introduction}
-
-When modern software systems are developed, their design 
-changes continuously.
-Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
-These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system.
-In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{khomh12}\cite{li07}\cite{fab15_1}.
-Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{fab15_1}\cite{chatzigeorgiou10}\cite{arcoverde11}.
-This makes early detection of these code smells a necessity to maintain sufficient quality in software systems.
-This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed not only by identifying potential anti-patterns, but also by (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
-
-Tools for detecting code smells have been the focus of several researchers lately, \cite{Tufano2015}\cite{boussaa13}.
-There are two  popular techniques for detecting smells: (1) by the making use of static code analysis and code quality metrics\cite{tsantalis09}; (2) Analysing software change history\cite{moha2010}.
-While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers with planning actions to improve the quality of source code.
-Similar concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
-
-At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
-Panelists commented  ``prediction is all well and good, but what about decision making?''.
-The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
-We call such tools \textit{planners}.
-In the context of detecting code smells, the planners are required to address the following issues.
-
-\bi
-\item
-Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to develop tools that can be used to reason about the code smells. 
-\item
-In addition to detecting anti-patterns and code smells, tools must be capable of having a contextual understanding of the data to trigger only the most useful refactoring options.
-\item Tools must provide a visual medium for developer to gauge the nature of refactoring options, thereby discarding bad solutions.
-\ei
-
-In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively.
-As a measure of source code quality we take into account the number of classes with defects in them.
-To this end, we use defect count data set by Jureczko et al.
-for JAVA systems~\cite{jureczko10}.
-In addition to defects, the aforementioned data set measures a list of static code attributes associated with each class.
-These code attributes allow us to reason about the measurable properties of anti-patterns.
-
-The contributions of this paper are (1) the new XTREE planning algorithm and (2) an evaluation strategy that shows the tool performing significantly ``better'' than  the  planner  proposed in our prior work~\cite{me12c}.
-
-The work on XTREE is motivated by the following three research questions:
-
-\bi
-\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reducing the number of defects.
-Specifically, we use XTREE to suggest a number of valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
-In order to ensure the changes are feasible, XTREE generate succinct plans.
-
-\item \textbf{RQ2:} \textit{Can a planner be used to critique bad refactoring options?} In addition to being able to suggesting changes, it's beneficial for planners to be able to quantitatively comment on the quality of certain refactoring options, thereby preventing unnecessary refactoring.
-
-\ei
-
-The rest of this paper  describes our data, our planners, and the experiment that ranks XTREE against alternate approaches.
-This is followed by notes on related work and validity.
-To enable reproducibility, all scripts and data used in this study are available online at \url{http://git.io/vG3DG}.
+\begin{raggedleft}
+{\em My offense is rank, it smells to heaven.}\\
+-- Hamlet, Act 3, Scene 3
+
+\end{raggedleft}
+
+~
+
+% Developers, tools, and research papers are either vague or disagree
+% about what code refactorings are most important to perform, or ignore.
+% As noted by Shantnawi~\cite{Shatnawi10} there is a dearth of
+% studies that formulate 
+% precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
+% claim that OO design metrics (e.g. depth of inheritance,
+% classes per method, etc) can be used to predict
+% code quality. Yet they their paper never comments at what point
+% a particular code measurement turns from ``ok'' to ``bad smell''.
+% Similary, code smell detection tools work for a certain
+% range of
+
+% When there are many possible code refactorings, it becomes difficult
+% to choose which ones to execute. 
+% Precious project resources can
+% be wasted if developers spend too much time on useless refactorings.
+
+According to   Fowler~\cite{fowler99}, code smells (a.k.a. bad smells)
+are ``a surface indication that usually corresponds to a deeper problem''.
+Much research enthusiastically endorses ``bad smell''-based refactoring, as witnessed
+by the  
+recent literature of Tufano et al. at ASE'15~\cite{Tufano2015} which
+lists dozens of  
+papers proposing bad smell detection and repair tools.
+
+Should we remove bad smells whenever we find them? Perhaps not.
+
+Firstly, as noted by Shantnawi~\cite{Shatnawi10} there are  there are few studies that formulate 
+precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
+discuss what OO design metrics might defect bad smells (e.g. a excessive
+depth of inheritance tree) but they never comment at what point
+a particular code measurement turns from ``ok'' to ``bad smell''.
+
+Secondly, it is not clear that all code smells are relevant to
+all projects.
+Developers often disagree on what bad smells, if any at all,
+should be refactored.
+For example, 
+Mantyla et al.~\cite{Mantyla2004} found that within an organization,
+ developers had different perceptions on the importance of different
+ bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
+ surveyed 85 professional developers to conclude that it remains
+ open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
+ Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
+ performing maintenance tasks that took three to four weeks to complete. 
+ After modifying nearly 300 Java files, it was observed
+ that none of the dozen investigated code smells explored in 
+ that study were mostly irrelevant to to that maintenance work.
+
+ How do we reconcile the enthusiasm of Fower,  Tufano, Darcy, Kemerer et al.
+ with the more pessimism of Mantyla, Yamashita, 
+ and Sj\"oberg et al.? The starting
+ point of this paper is a comment
+ from Mantyla et al.~\cite{Mantyla2004}. They say that
+ when making  refactoring decisions, it is not enough to
+ just apply some bad smell rule generated from another project.
+ Rather, they say, decisions about refactoring
+ should be based based on a combination of (1)~subjective evaluations and 
+ (2)~code metrics.
+
+ Given the nature of human
+ expertise, developers will offer many  strong opinions about what requires changing.  That is, we   anticipate no shortage of (1)~subjective
+ evaluations from developers.
+
+ As to lessons learned from (2)~code metrics, this paper introduces RANK,
+ a tool for sorting proposed changes to static code measures. Unlike
+ normal classification algorithms, RANK is a {\em contrast set learner}.
+  Contrast set learners do not report {\em what is}. Rather
+     they recommend {\em what to change} in order to move from
+     one classification to another (e.g. from defective=yes to
+     defective=no).
+     When applied to defect predictors from static code measures,
+that delta is a change to static code measures;
+i.e. a code refactoring.
+
+RANK's concept of operation is that developers {\em propose}
+numerous refactorings and RANK {\em disposes} of all those with
+no support in history data.
+Experiments with RANK explain the pessimism of Mantyla, Yamashita, 
+ and Sj\"oberg et al. As they noted, sometimes, pre-defined bad
+ smells are irrelevant to reducing the risk of defects in a project.
+ However, when bad smells are learned from the current project,
+ it is possible to recognize what are useful, and useless, local
+ changes (where ``useful'' is defined in terms of reducing the
+ predicted number of defects).
+  Hence,   we  can endorse
+ the use of bad smells to guide refactorings, but
+ only when tools like RANK can  tune the definitions of
+ ``good'' or ``bad'' refactorings using current project data.
+
+  The contributions of this paper are:
+  \begin{itemize}
+      \item A
+  \end{itemize}
+
+The rest of this paper is structured as follows. XXX
+
+Before beginning, we make the obvious aside
+that defect predictors are not the {\em only} way to assess
+bad smells. Some researchers report that code reviews and debating potential refactorings are useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
+debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
+about interactions within the code. In that view, 
+the concept of refactoring can be useful
+even if it does not address issues of defects. That said,
+reducing the risk of errors was
+Fowler's original motivation for introducing refactoring. Also, using
+defect predictors,
+it is possible to operationalize this test for $R_i \succ R_j$
+while it might be harder to otherwise assess refactorings.
+Hence, 
+this paper assumes that {\em better} refactorings and
+bettering at {\em reducing} the chances
+that code will contain defects.
 
@@
 
-\input{tex/jur_char.tex}
-
-The thesis of this paper is that the XTREE tool can stop developers
-wasting time on useless software refactorings. Before presenting experiments
-that supports that claim, this section motivates the need for 
-such tools.
-
-In many fields,  practitioners hang on to persistent beliefs, even
-if the evidence for those beliefs is not strong.
+A frequently asked question about this work is 
+why build tools like RANK to second-guess and critique developer intuitions?
+Our reply is that  recent studies point to a  widespread malaise in software engineering:
+beliefs about software are rarely revised and hence may be
+  inaccurate and 
+misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
+strong views on many issues, including bad smells, but those views may be wrong for the current
+project.
+
+To be fair, software engineering is not the only field with this problem.
   For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10).
@@
 As in medicine, so too in software engineering.
-There is growing evidence that amongst programmers,
-many of their  widely-held beliefs are inaccurate and misleading. 
 According to Passos et al.~\cite{passos11},  developers often  assume that the lessons they learn from a few past projects are general to all their future projects.
@@
 
-To support the last paragraph, we offer two
-specific examples
-were  widely-held views in SE that are   now questionable, given recent results. 
-Firstly, Nagappan et al. analyzed~\cite{mei15} 
-how the {\tt goto} statement was used  in 11,000 Github repositories, 
-they found that developers
+To support the last paragraph, we offer several
+specific studies
+showing that   widely-held views in SE that are   now questionable, given recent results. 
+The first three come from the general SE literature while the rest
+relate specifically to bad smells:
+\begin{enumerate} 
+\item
+Nagappan et al. analyzed~\cite{mei15} 
+how   {\tt goto}  was used  in 11,000 Github repositories. 
+They found that developers
 rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
 memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
-
-Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
+\item
+In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
 the ``phase delay'' truism that {\em fixing a requirements error  during
 coding can be thousands of times more expensive than recognizing and fixing that error
-as soon as it arises}. Boehm and Basili comment that this phase delay effect
-``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
-After reading the literature, and surveying academics
-and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
-this effect seems now heavily mitigated by $21^{st}$ software development languages,
-tools and development practices.  When we mined the 171 software projects shepherded
-by the SEI in the period 2006-2014, we found that phase delay results
-in only minor changes in the 
-time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.
-
-These two studies are symptomatic
-of a widespread malaise in software engineering.
+as soon as it arises}. 
+We found no evidence for such large   phase delay effect in modern
+software\footnote{171 software projects shepherded
+by the SEI, 2006-2014} possibly this has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices). 
+Whatever the reason, the main point here is that phase delay remains a strongly
+held belief, despite little recent evidence to support it.
+\item
 Devanbu et al.  examined responses from 564 Microsoft software developers from around
-the world, They found that  `` (a)~programmers do indeed have very
+the world, They found that  ``(a)~programmers do indeed have very
 strong beliefs on certain topics; (b)~their beliefs are primarily formed
@@
 correspond with actual evidence in that project''~\cite{prem16}.
+\end{enumerate}
 Devanbu et al. further  comment that ``programmers give personal experience
@@
 
-One area where poorly-justifed beliefs can waste valuable project resources is
-refactoring in order to remove code bad smells.
+\input{tex/refactor}
+
+ If widely-held beliefs in SE can be inaccurate and misleading,
+ it is timely to check the utility of bad smells.
+ This check is important since
+ the concept of bad smells is highly pervasive in the SE literature\footnote{.
+As of  Feburary 2016, ACM Digital Portal reports 
+that Fowler's 1999 text has 6434 references;
+while a search  on "bad smell"+refactoring yields 824 citations in
+the last decade at learning research
+venues (ICSE, FSE, etc).}. Accordingly,
+\fig{smells} 
+lists the commonly mentioned bad smells, in the first column.
+The other
+columns show other work that comments on which bad smells matter most.
+A  blank cell indicates where the other work has chosen to ignore
+of the Fowler bad smells as unimportant. 
+Note that most of the cells are blank.
+For example, the Atwood'06~\cite{Atwood06} study takes the time to comment on   
+all the Fowler bad smells.
+However, all the  other  studies ignore the majority of the Fowler bad smells.
+
+Two of the studies in \fig{smells} offers some comments on the relative importance
+of the different bad smells:
+\begin{itemize}
+\item The Yamashita’13~\cite{Yamashita2013} survey comments on the relative importance
+    of 14 code smells-- only eight of which are mentioned in
+    the left-hand-column.
+  \item  
+   The {\em developer survey}   (in the right-hand-side column}
+   shows the opinions we collected from a dozen developers from a Washington
+    D.C. web-tools development organization
+     Amongst those group, there was there was  some
+    consensus on  the priority of fixing which  bad smells   
+    (see the annotations VH=very high,
+    H=high, L=low, VL=very low, and ``?''= no consensus).  
+\end{itemize} 
+    One interesting point of agreement of the  Yamashita’13 study and the 
+    developer survey is the agreement on which  bad smells
+    were high priority:
+    \begin{itemize}
+        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey.
+        \item Those three were {\em duplicated code, large class}, 
+        and {\em long method}.
+        \ei
+ Note that this agreement also means that the
+  Yamashita’13 study and the {\em developer survey}   concluded
+        that most of the  listed code smells are {\em not} high priority issues
+        requiring refactoring.
+
+In summary, just because a developer strongly believes something
+about bad smells, does
+not necessary mean that that belief is useful for the current project.
+Human developers can be clever, but their thinking can also be flawed.
+Hence it is important to double check their beliefs in, for example,
+bad smells. Tools like HISTORY perform that check, using models learned
+from the historical record.
+
+\input{tex/jur_char}
+
+\subsection{Correlation, Causation, and Critique}
+
+We propose HISTORY as way to critique (and possibly dismiss) proposed
+refactorings. This section discusses the value of that approach.
+
+HISTORY is a statistical method for finding ways to change code
+such that the new code has a lower chance of having defects. 
+To make these recommendations,
+HISTORY  inputs frequency counts of how often a static
+code attribute is seen in defective or non-defective modules.
+Hence, its recommendations about what to change may be wrong
+due to the famous correlation vs causation problem.
+Statistical methods that use frequency counts can only see how often
+certain events occur together, but not why that happens. Hence,
+they may conclude incorrectly that with two co-located situations,
+one might cause the other. This is not necessarily true. For
+example, the number of storks and the number of births
+in Denmark is not causally
+connected. Nevertheless, both may be correlated if certain
+migratory patterns of storks coincide with certain sociological factors.
+
+Another reason that HISTORY's recommendations might be incorrect
+is the presence of hidden constraints that are not apparent in
+its input data.
 
@@
 }}{}
+
+It turns out that ranking codiiiiie smells, and ignoring the low priority
+ones, is a pressing problem. As shown in our {\em Related Work} section,
+different developers and  books and tools stress the importance of different code smells. How to assess these different claims? How to avoid
+needless refactoring? How to check if a code smell found at another site
+is relevant to this site?
+
+To answer all these questions, we propose NOSE, a tool that checks
+if cpde smells really smell 
+generates candidate refactorings to reduce the chances of software
+defects. XTREE is a back-end to a decision tree learner that
+reflects on how different branches in that tree predict for
+defective or non-defective code modules. Since this tree
+is built from static code attributes, the differences between
+those brnachesy reflecting on the
+{\em delta} between those branches, it is
+
+When modern software systems are developed, their design 
+changes continuously.
+Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
+These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system.
+In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{khomh12}\cite{li07}\cite{fab15_1}.
+Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{fab15_1}\cite{chatzigeorgiou10}\cite{arcoverde11}.
+This makes early detection of these code smells a necessity to maintain sufficient quality in software systems.
+This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed not only by identifying potential anti-patterns, but also by (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
+
+Tools for detecting code smells have been the focus of several researchers lately, \cite{Tufano2015}\cite{boussaa13}.
+There are two  popular techniques for detecting smells: (1) by the making use of static code analysis and code quality metrics\cite{tsantalis09}; (2) Analysing software change history\cite{moha2010}.
+While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers with planning actions to improve the quality of source code.
+Similar concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
+
+At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
+Panelists commented  ``prediction is all well and good, but what about decision making?''.
+The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
+We call such tools \textit{planners}.
+In the context of detecting code smells, the planners are required to address the following issues.
+
+\bi
+\item
+Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to develop tools that can be used to reason about the code smells. 
+\item
+In addition to detecting anti-patterns and code smells, tools must be capable of having a contextual understanding of the data to trigger only the most useful refactoring options.
+\item Tools must provide a visual medium for developer to gauge the nature of refactoring options, thereby discarding bad solutions.
+\ei
+
+In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively.
+As a measure of source code quality we take into account the number of classes with defects in them.
+To this end, we use defect count data set by Jureczko et al.
+for JAVA systems~\cite{jureczko10}.
+In addition to defects, the aforementioned data set measures a list of static code attributes associated with each class.
+These code attributes allow us to reason about the measurable properties of anti-patterns.
+
+The contributions of this paper are (1) the new XTREE planning algorithm and (2) an evaluation strategy that shows the tool performing significantly ``better'' than  the  planner  proposed in our prior work~\cite{me12c}.
+
+The work on XTREE is motivated by the following three research questions:
+
+\bi
+\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reducing the number of defects.
+Specifically, we use XTREE to suggest a number of valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
+In order to ensure the changes are feasible, XTREE generate succinct plans.
+
+\item \textbf{RQ2:} \textit{Can a planner be used to critique bad refactoring options?} In addition to being able to suggesting changes, it's beneficial for planners to be able to quantitatively comment on the quality of certain refactoring options, thereby preventing unnecessary refactoring.
+
+\ei
+
+The rest of this paper  describes our data, our planners, and the experiment that ranks XTREE against alternate approaches.
+This is followed by notes on related work and validity.
+To enable reproducibility, all scripts and data used in this study are available online at \url{http://git.io/vG3DG}.
+
+Fowler claims   removing   code smells   by
+\begin{quote}
+...applying a series of small behavior-preserving transformations, each 
+of which seem ``too small to be worth doing''. 
+The  effect of   these transformations is quite significant. By doing them in small steps you {\em reduce the risk of introducing errors}'' (our italics).
+\end{quote}
+The italics in the last paragraph offer one method for indicating if bad smell-based
+refactoring was successful. There are many automatic
+methods for learning   defect predictors from static
+code features~\cite{peters15,he13}.  We can use those
+predictors to rank $N$ proposed refactorings $R_1,R_2,...R_i$
+refactorings. Specifically, we  say that refactoring $i$ is preferred to $j$
+(denoted $R_i \succ R_j$) if the defect predictors predict fewer errors
+after applying $R_i$ than $R_j$.
+
+The ability to rank, and perhaps   reject, a proposed refactoring is
+an important contribution, given the current level of interest
+in bad smell-based refactoring.
+
+Nevertheless,
+certain results raise doubts about the value of   code smells for reducing the risk of defects in software.

### 2016-03-03 tex/refactor.tex
Updates from ShareLaTeX

\begin{figure*}[!t] 
\scriptsize
\centering
\begin{tabular}{r|l|c|c|c|c}
Mostly from Fowler'99& & & & & Developer\\
 (with a few additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey\\\hline
  Alt. Classes with Diff. Interfaces & & & & + & \\
  Combinatorial Explosion~\cite{Kerievsky2004} & & & & + & \\
  Comments & & & 11 & + & VL\\
  Conditional Complexity~\cite{Kerievsky2004} & & & 14 & + & ?\\
  Data Class & Data Class & & & + &\\
  Data Clumps &  &  & & + &\\
  Divergent Change & & & & + & \\
  Duplicated Code & Significant Duplication\textsuperscript{*} & + & 1 & + & VH\\
  Feature Envy & Feature Envy & & 8 & + &\\

  Inappropriate Intimacy & & - & & + & L\\
  Indecent Exposure~\cite{Kerievsky2004} & & & & + & ?\\
  Incomplete Library Class & & & & + &\\
  Large Class & God Class/Brain Class & ++ & 4 & + & VH\\
  Lazy Class/Freeloader & & - & 7 & + &\\
  Long Method & Brain Method & ++ & 2 & + & VH\\
  Long Parameter List &  & + & 9 & + & L \\

  Message Chains & & & & + & H\\
  Middle Man & &  & & + &\\
  Oddball Solution~\cite{Kerievsky2004} & & & & + & \\
  Parallel Inheritance Hierarchies & & & & + &\\
  Primitive Obsession &  & & & + &\\

  Refused Bequest & Refused Parent Bequest & - & & + & \\ 
  Shotgun Surgery & Shotgun Surgery & & & + & \\
  Solution Sprawl~\cite{Kerievsky2004} & & & & + &\\
  Speculative Generality & & & & + & L\\
  Switch Statements &  & - & & & L\\
  Temporary Field & & - & & + & ?\\
  \end{tabular}
%   \multicolumn{5}{l}{} \\
%   \multicolumn{5}{l}{\textsuperscript{\textdagger} + Exact rule, - Possibly related rule} \\
%     \multicolumn{5}{l}{\textsuperscript{\textdaggerdbl} + is included in list} \\
%     \multicolumn{5}{l}{\textsuperscript{\textexclamdown} VH: Very High, H: High, L: Low, VL: Very Low, ?: Still investigating} \\
%   \multicolumn{5}{l}{\textsuperscript{*} Computation is based on more than simple metrics.} \\

\caption{List of bad code smells from different sources. Empty cells
denote some bad smell proposed by Fowler'99 that was not found relevant
in subsequent studies. Note that there are many blank cells.}
\label{fig:smells}
\end{figure*}

### 2016-03-08 Manuscript.tex
Updates from ShareLaTeX

--- 2016-02-28 Manuscript.tex
@@
 
-\usepackage{stfloats}
+%\usepackage{stfloats}
 
@@
 \usepackage{amsmath}
+\usepackage{ amssymb }
 \usepackage{url}
@@
 
-\title{Learning Which Bad Smells are 
-Real Stinkers?}
-\numberofauthors{4} 
+\title{  How to Recognize  a Bad  ``Bad Smell'' (using Contrast Sets)}
+\numberofauthors{2} 
 \author{  
 \alignauthor
-Rahul~Krishna\\Tim Menzies, Xipeng~Shen \\
-       \affaddr{CS, NC State, USA}\\
-       {\{i.m.ralk,~tim.menzies,\\xipengshen\}@gmail.com}
-\alignauthor
-Andrian Marcus \\
-       \affaddr{CS, UT Dallas  } \\ 
-       \affaddr{Texas, USA}\\ 
-       {amarcus@utdallas.edu}
-       \alignauthor
-Naveen  \\ Lekkalapudi\\
- \affaddr{Bellhops } \\ 
-       \affaddr{Tennessee, USA}\\ 
-       {navek91@gmail.com}
+Rahul~Krishna, Tim Menzies  \\
+       \affaddr{Computer Science, NC State, USA}\\
+       {\{i.m.ralk,~tim.menzies\}@gmail.com}
 \alignauthor
 Lucas Layman \\
-       \affaddr{Fraunhofer CESE  } \\ 
-       \affaddr{College Park, USA}\\ 
+       \affaddr{Fraunhofer CESE, College Park, USA}\\ 
        {llayman@fc-md.umd.edu}
@@
 \begin{abstract} 
-
+It is hard to assess
+if a bad smell is ``bad''; i.e. if it is irrelevant or misleading
+for a specific project.
+Developers, text books, and tools
+disagree on which bad smells are important.
+Specifically, it is unclear at what threshold values
+for which metrics should trigger code reorganization since
+most studies do not offer
+metrics {\em and} a threshold for their proposed bad smells.
+This paper proposes the XTREE contrast set
+learner to assess bad smells
+and reject any that are not supported by
+evidence in the historical record.  
+Using XTREE, we   show that the range of useful
+``bad smells'' changes from project to project. Hence,
+before refactoring, it is important to run a critique tool
+like XTREE to stop work on useless code reorganizations.
+Further, using XTREE, it is possible to show that many
+of the bad smells endorsed by standard texts or tools
+are not useful for particular projects. 
+Based on these results, we endorse  
+guiding code reorganization via bad smells. However,  
+it is important to XTREE the bad smells defined for a 
+particular project.
+
+{\bf Categories/Subject Descriptors:} D.2 [Software Engineering]; I.2 [Artificial Intelligence];
+
+{\bf Keywords:} Bad smells,
+performance prediction,  decision trees 
 \end{abstract}
 
-\keywords{Defect prediction; Configuration; Planning; Decision trees.}
 \section{Introduction}
 
-When modern software systems are developed, their design 
-changes continuously.
-Due to market constraints, the developers tend to ignore the quality of source code, thereby introducing \textit{code smells} (or \textit{anti-pattern}).
-These anti-patterns are generally viewed as poor solutions to recurring design problems, possibly deterring the maintainability of the system.
-In the past decade there has been growing evidence of the impact of anti-patterns, there now is also some empirical evidence to suggest that code containing anti-patterns are more fault prone than the rest of the code~\cite{khomh12}\cite{li07}\cite{fab15_1}.
-Studies show that the number of code smells increase over time and in only a very few cases are these smells addressed by refactoring \cite{fab15_1}\cite{chatzigeorgiou10}\cite{arcoverde11}.
-This makes early detection of these code smells a necessity to maintain sufficient quality in software systems.
-This being said, detection only addresses a part of the problem, what we need are recommendation systems that can assist a software developed not only by identifying potential anti-patterns, but also by (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
-
-Tools for detecting code smells have been the focus of several researchers lately, \cite{Tufano2015}\cite{boussaa13}.
-There are two  popular techniques for detecting smells: (1) by the making use of static code analysis and code quality metrics\cite{tsantalis09}; (2) Analysing software change history\cite{moha2010}.
-While these tools are capable of detecting these smells with a reasonable accuracy, they fail to provide developers with planning actions to improve the quality of source code.
-Similar concerns were voiced by several business users, who are demanding tools that support business-level interpretations of their data.
-
-At a panel on software analytics at ICSE`12, industrial practitioners lamented the state of the art in software analytics~\cite{menzies12a}.
-Panelists commented  ``prediction is all well and good, but what about decision making?''.
-The panelists were more interested in the interpretations and follow-up that occurs after the mining, rather than just the mining itself.
-We call such tools \textit{planners}.
-In the context of detecting code smells, the planners are required to address the following issues.
-
-\bi
-\item
-Instead of just \textit{detecting} anti-patterns and code smells, it may be beneficial to develop tools that can be used to reason about the code smells. 
-\item
-In addition to detecting anti-patterns and code smells, tools must be capable of having a contextual understanding of the data to trigger only the most useful refactoring options.
-\item Tools must provide a visual medium for developer to gauge the nature of refactoring options, thereby discarding bad solutions.
-\ei
-
-In response to this, we propose a novel {\em planning} method called XTREE which reflects on static code quality metrics to learn refactoring options to a software system such that its performance ``improves'' quantitatively.
-As a measure of source code quality we take into account the number of classes with defects in them.
-To this end, we use defect count data set by Jureczko et al.
-for JAVA systems~\cite{jureczko10}.
-In addition to defects, the aforementioned data set measures a list of static code attributes associated with each class.
-These code attributes allow us to reason about the measurable properties of anti-patterns.
-
-The contributions of this paper are (1) the new XTREE planning algorithm and (2) an evaluation strategy that shows the tool performing significantly ``better'' than  the  planner  proposed in our prior work~\cite{me12c}.
-
-The work on XTREE is motivated by the following three research questions:
-
-\bi
-\item \textbf{RQ1:} \textit{Can a planner suggest plans to improve the quality of a defective class?} This research question investigates the effectiveness of a planner in suggesting useful refactoring options to improve the quality of class by reducing the number of defects.
-Specifically, we use XTREE to suggest a number of valid changes to source code and we use an oracle to measure the effectiveness of the changes, see \tion{eval}.
-In order to ensure the changes are feasible, XTREE generate succinct plans.
-
-\item \textbf{RQ2:} \textit{Can a planner be used to critique bad refactoring options?} In addition to being able to suggesting changes, it's beneficial for planners to be able to quantitatively comment on the quality of certain refactoring options, thereby preventing unnecessary refactoring.
-
-\ei
-
-The rest of this paper  describes our data, our planners, and the experiment that ranks XTREE against alternate approaches.
-This is followed by notes on related work and validity.
-To enable reproducibility, all scripts and data used in this study are available online at \url{http://git.io/vG3DG}.
+% Developers, tools, and research papers are either vague or disagree
+% about what code refactorings are most important to perform, or ignore.
+% As noted by Shantnawi~\cite{Shatnawi10} there is a dearth of
+% studies that formulate 
+% precise guidelines,   represented as threshold values, to interpret the complexity of the software design using metrics. For example,  Darcy and Kemerer~\cite{darcy05}
+% claim that OO design metrics (e.g. depth of inheritance,
+% classes per method, etc) can be used to predict
+% code quality. Yet they their paper never comments at what point
+% a particular code measurement turns from ``ok'' to ``bad smell''.
+% Similary, code smell detection tools work for a certain
+% range of
+
+% When there are many possible code refactorings, it becomes difficult
+% to choose which ones to execute. 
+% Precious project resources can
+% be wasted if developers spend too much time on useless refactorings.
+
+\noindent
+According to   Fowler~\cite{fowler99}, bad smells (a.k.a. code smells)
+are ``a surface indication that usually corresponds to a deeper problem''.
+Fowler strongly recommends   removing   code smells   by
+\begin{quote}
+``...applying a series of small behavior-preserving transformations, each 
+of which seem ``too small to be worth doing''. 
+The  effect of   these transformations is quite significant. By doing them in small steps you {\em reduce the risk of introducing errors}'' (our italics).
+\end{quote}
+The literature is divided on the merits of bad smells.
+Much research  endorses using them as a guide
+code change. For example,
+the recent literature review of Tufano et al. at ASE'15~\cite{Tufano2015}  
+lists dozens of  
+papers proposing bad smell detection and repair tools.
+
+On the other hand, other papers cast doubts on the value of use bad smalls
+as triggers for code change.
+In our own work, documented later in this paper, we  show
+that tools and text books and developers can disagree on what bad smells
+are important enough to hunt down and eliminate. Our findings
+are consistent with several other papers.
+Mantyla et al.~\cite{Mantyla2004} found that within an organization,
+ developers had different perceptions on the importance of different
+ bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
+ surveyed 85 professional developers to conclude that it remains
+ open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
+ Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
+ performing maintenance tasks that took three to four weeks to complete. 
+ After modifying nearly 300 Java files, it was observed
+ that none of the dozen investigated code smells explored in 
+ that study were mostly irrelevant to to that maintenance work.
+
+Debates on when/what code should be restructed are interesting, especially
+in the current context of  widely-used SE practices such as
+test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
+In that approach, developers are always asking themselves ``should I reorganize this code''?
+Bad smells are a widely used  technique for deciding when to reorganize\footnote{Fowler textbook is widely read-- as of  early 2016, ACM Digital Portal reports 
+it has 6434 references. Also, bad smell-based research
+is widely reported-- a search  on {\em "bad smell"+programming} 
+where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
+Hence it is important to understand the value of
+ bad smells.  
+Valuable development resources can be wasted if developers
+change their code based on   bad smells that are ``bad''; i.e. 
+ irrelevant or misleading
+for some particular project.
+
+It turns out that it is hard to learn ``good'' bad smell detectors. Such a detector
+is a combination of some code metric (e.g. lines of code in a method)
+and a threshold measurement (e.g. {\em loc} $>$ 100). 
+Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
+{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
+discuss what OO design metrics might defect bad smells (e.g. a excessive
+depth of inheritance tree) but they never comment on what thresholds
+means that some section of code starts to smell bad. After
+an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
+of  studies  conducted to formulate bad smell guidelines,
+represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.
+
+Why are there so few papers offer bad smell metrics and thresholds? For the few that
+do exist, why do developers disagree on the relevance of those bad smells?
+As discussed on the next page, one reason for this is that 
+   cognitive biases result in different
+developers declaring that
+bad smell X or Y is most important.
+
+When developers cannot fully assess the value of bad smells,
+it is possible to use automatic means. We assume that
+developers will change code until a bad smell detector stops triggering.
+That is, if the bad smell detector is {\em loc} $>$ 100, they will
+redistribute the functionality of a long method into several smaller ones.
+Recalling the italics in the above quote, Fowler's criteria for successful bad smell removal is the extent
+to which the reorganized code has fewer errors/defects.
+Data miners can be used to predict the defects in the reorganized code
+so, in this paper, we will use static code defect predictors to assess
+bad smells.
+
+After making that assessment, this paper finds
+that   state-of-the-art methods
+(Shantnawi~\cite{Shatnawi10}  in TSE 2010;
+and our own instance-based systems~\cite{me12c} from 2012) 
+are less effective for bad smell assessment than a new
+contrast set learning method called XTREE.  
+Standard learners, such as the regression methods used by Shantnawi~\cite{Shatnawi10}
+only comment on current distributions.  
+  Contrast set learners   recommend {\em what to change} in order to move from
+     one classification to another (e.g. from defective=yes to
+     defective=no).
+     When applied to defect predictors from static code measures,
+that delta is a change to static code measures;
+i.e. a code reorganization. Hence, constrast set learners
+are a better way to learn when code needs to be reorganized
+than, say, the methods of Shantnawi~\cite{Shatnawi10}.
+
+We recommend using XTREE as a judge to rank the value of a set
+of bad smell detectors:
+\begin{itemize}
+    \item Consider $N$ developers each proposing a software reorganization
+based on a set of bad smells $S$.
+\item XTREE will (a)~learn contrast sets
+from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
+ on any member of $S_{\mathit{bad}} \in S$
+that XTREE could {\em not} be found in the historical record.
+\end{itemize}
+Note that we    recommend  XTREE  is used to critize, but  not to create,
+bad smells. 
+XTREE uses frequency counts of code features seen in the historical record
+to generate
+predictions about the effects of changing code
+attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
+where spurious correlations can suggest effects that are not
+truly causal.
+That said, if someone else's proposed bad smell detector $S_{\mathit{bad}} $ falls {\em outside} of the space of
+effects found  using XTREE's analysis, then we can say that XTREE found 
+{\em neither} correlation {\em or}
+causal evidence that $S_{\mathit{bad}} $ can be found in the historical record.
+
+XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
+by developers. XTREE does not perform those refactorings.
+Secondly, XTREE assumes   Fowler's bad smells
+should be assessed via their contribution to software defects.   Some researchers 
+ comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
+debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
+about interactions within the code. In that view, 
+the concept of refactoring can be useful
+even if it does not address issues of defects. 
+That said, using
+defect predictors,
+it is possible to operationalize XTREE
+while it might be harder to otherwise operationalize the assessment of bad smells.
+Hence, 
+this paper assumes that {\em better} bad smells and
+bettering at finding code changes that  {\em reducing} the chances
+that the new code will contain defects.
+
+The contributions of this paper are
+\begin{itemize}
+
+    \item A new kind of contrast set learner called XTREE
+    that can critique plans of what to change within a data set in order
+    to improve outcomes.
+    \item An experimental rig for assessing the value of tools like
+    XTREE that propose changes to software. That rig uses data mining
+    to build an oracle for the proposed changes.
+    \item Using that learner and that oracle, we offer a method for avoid code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
+    \item A resolution between proponents of bad smells (e.g.
+Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}).
+We say
+    that bad smells can guide code change, but only if those bad smells are
+    first carefully assessed. 
+\end{itemize}
+
+The rest of this paper discusses the cognitive
+biases which make developers disagree on the relevance of those bad smells to
+their particular projects.  
+ That discussion is our motivation for developing the XTREE  tool
+that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.
 
@@
 
-\input{tex/jur_char.tex}
-
-The thesis of this paper is that the XTREE tool can stop developers
-wasting time on useless software refactorings. Before presenting experiments
-that supports that claim, this section motivates the need for 
-such tools.
-
-In many fields,  practitioners hang on to persistent beliefs, even
-if the evidence for those beliefs is not strong.
+Why build tools like XTREE to  critique proposed developer actions?
+Are not those developers the best guide to what is important, or irrelevant
+to their code?
+
+It turns our that developer {\em cognitive biases} can mislead them into
+asserting that some things are important and relevant when they are not.
+This section discusses those biases-- firstly in the general sense
+across all SE beliefs; then secondly in the   context of code smells.
+
+A widespread malaise in software engineering is that
+beliefs about software are rarely revised and hence may be
+  inaccurate and 
+misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
+strong views on many issues, including bad smells, but those views may be wrong for the current
+project.
+It should be noted that  software engineering is not the only field with this problem.
   For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10).
@@
 
-To support the last paragraph, we offer two
-specific examples
-were  widely-held views in SE that are   now questionable, given recent results. 
-Firstly, Nagappan et al. analyzed~\cite{mei15} 
-how the {\tt goto} statement was used  in 11,000 Github repositories, 
-they found that developers
-rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his famous 1968
+To support the last paragraph, we offer several
+specific studies
+showing that   widely-held views in SE that are   now questionable, given recent results. 
+The first few come from the general SE literature while the rest
+relate specifically to bad smells:
+\input{tex/refactor}
+\begin{enumerate} 
+\item
+Nagappan et al. analyzed~\cite{mei15} 
+how   {\tt goto}  was used  in 11,000 Github repositories. 
+They found that developers
+rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
 memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
-
-Secondly, in other work with the Software Engineering Institute (SEI), we have revisited
+\item
+In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
 the ``phase delay'' truism that {\em fixing a requirements error  during
 coding can be thousands of times more expensive than recognizing and fixing that error
-as soon as it arises}. Boehm and Basili comment that this phase delay effect
-``..has been a major driver in focusing industrial software practice on thorough requirements analysis and design, on early verification and validation, and on up-front prototyping and simulation to avoid costly downstream fixes''~\cite{boehm01}.
-After reading the literature, and surveying academics
-and practitioners, we can assert that this is a very widely-held belief. Nevertheless,
-this effect seems now heavily mitigated by $21^{st}$ software development languages,
-tools and development practices.  When we mined the 171 software projects shepherded
-by the SEI in the period 2006-2014, we found that phase delay results
-in only minor changes in the 
-time to resolve issues (usually less than a 30\% increase)~\cite{me16phase}.
-
-These two studies are symptomatic
-of a widespread malaise in software engineering.
+as soon as it arises}. 
+We found no evidence for such large   phase delay effect in modern
+software (in 171 software projects shepherded
+by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
+Whatever the reason, the main point  is that phase delay is a widely
+held belief, despite little recent evidence to support it.
+\item
 Devanbu et al.  examined responses from 564 Microsoft software developers from around
-the world, They found that  `` (a)~programmers do indeed have very
+the world, They found that  ``(a)~programmers do indeed have very
 strong beliefs on certain topics; (b)~their beliefs are primarily formed
@@
 
-One area where poorly-justifed beliefs can waste valuable project resources is
-refactoring in order to remove code bad smells.
+If the above remarks hold true for bad smells, then we would expect
+to see much disagreement on what bad smells are important and relevant
+to  a particular project. This is indeed the case.
+The first column of \fig{smells} 
+lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{Fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
+The other
+columns show other work that comments on which bad smells matter most.
+The columns marked as Lanza'06, Yamashita'13, and Atwood'06 come from
+peer reviewed literature. The column maked SonarQube is an open source
+code assessment tool that includes dectectors for six of the bad smells
+in column one. 
+The {\em developer survey}   (in the right-hand-side column),
+   this shows the results of a hour-long whiteboard sessions with a group of a
+ dozen developers from one Washington
+    D.C. web-tools development organization. In that study, participants
+    worked in a round robin manner to rank the bad smells they thought were
+    important (and any disagreements were discussed with the whole group).
+     Amongst those group, there was there was  some
+    consensus on  the priority of fixing which  bad smells   
+    (see the annotations VH=very high,
+    H=high, L=low, VL=very low, and ``?''= no consensus).
+
+A  blank cell in \fig{smells}
+indicates where   other work has chosen to ignore
+one the   bad smells in column one. 
+Note that most of the cells are blank.
+While the Atwood'06~\cite{Atwood06} study  comments on   
+all the   bad smells,
+all the  other  studies ignore the majority of the Fowler bad smells.
+For exampel SonarQube has no detectors for many of the column one bad smells.
+Also, nearly half the Yamashita list of bad smells
+    does not appear in the Fowler and Kerievsky list
+    (the eight numbers
+    in the  Yamashita'13 column show the rankings for the bad smells 
+    that overlap with Fowler and Kerievsky).
+
+Two of the studies in \fig{smells} offers some comments on the relative importance
+of the different bad smells the   Yamashita'13 study and the developer study:
+    \begin{itemize}
+        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
+        and {\em long method}.
+        \item
+ Note that this agreement also means that the
+  Yamashita'13 study and the {\em developer survey}   concluded
+        that most of the  listed code smells are {\em not} high priority issues
+        requiring code reorganization. 
+ \end{itemize} 
+In summary, just because a developer strongly believes something
+about bad smells, does
+not necessary mean that that belief is useful for the current project.
+Human developers can be clever, but their thinking can also be distorted
+by congitive biases.
+Hence it is important to double check their beliefs in, for example,
+bad smells. Tools like XTREE perform that check, using models learned
+from the historical record.
+
+\section{Oracles}
+\label{sect:eval}
+\input{tex/prediction_stats.tex}
+
+As highlighted previously, there are several detection techniques for anti-patterns, yet these fail to address the issue of checking the proposed changes for their efficacy.
+Much of the refactoring needs to evaluated for their usefulness.
+To this end, there needs to be an oracle, a procedure that can be employed to distinguish between good and bad options.
+The dearth oracles is not unique to this domain, for instance a recent survey by Barr et al \cite{barr15}
+
+Our strategy (discussed in greater detail below) divides this data into a training and a testing set.
+The purpose of training set is twofold: (1) We apply a data miner to construct an oracle that acts as a quality predictor; (2) We build the planners to reason across the data.
+
+Next, we employ the planners to generate plans to improve the quality of defective classes in the test set.
+In addition to the plans obtained from our planners we may have some common refactoring options that the developers tend to use.
+The oracle built previously can be used to assess the the changed examples and this can be used to compare refactoring operations.
+
+It's worth noting that this approach depends on having effective oracles for assessing the results.
+It proved to be more complicated to commission the Jureczko data sets for this study.
+For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could be salvaged using the techniques discussed in this section.
+
+\fig{j} shows our preliminary studies with the Jureczko data.
+Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}, this means that we are training on hundreds to thousands of classes and testing on smaller test suites).
+Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
+In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
+
+In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
+For such data, quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
+``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
+``pf''): the percent of non-fault classes that are {\em predicted} to be defective.
+
+As a preliminary study, we split the Jureczko  data into train and test groups.
+Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was built from the training data, then applied to the test data.
+The ``untuned'' columns of \fig{j} shows those results.
+We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
+For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
+With this threshold, however, none of our data sets were be suitable for this study.
+
+Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples and an under-sampling of majority-class examples.
+Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
+
+In summary, while we cannot trust predictors from some of our Jureczko data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+Accordingly, when this study explores the Jureczko data, we will use these five data sets.
+
+(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+
+\input{tex/jur_char}
 
@@
 This set of metrics is not restricted and can be easily extended with other metrics.
-
-\section{Oracles}
-\label{sect:eval}
-\input{tex/prediction_stats.tex}
-
-As highlighted previously, there are several detection techniques for anti-patterns, yet these fail to address the issue of checking the proposed changes for their efficacy.
-Much of the refactoring needs to evaluated for their usefulness.
-To this end, there needs to be an oracle, a procedure that can be employed to distinguish between good and bad options.
-The dearth oracles is not unique to this domain, for instance a recent survey by Barr et al \cite{barr15}
-
-Our strategy (discussed in greater detail below) divides this data into a training and a testing set.
-The purpose of training set is twofold: (1) We apply a data miner to construct an oracle that acts as a quality predictor; (2) We build the planners to reason across the data.
-
-Next, we employ the planners to generate plans to improve the quality of defective classes in the test set.
-In addition to the plans obtained from our planners we may have some common refactoring options that the developers tend to use.
-The oracle built previously can be used to assess the the changed examples and this can be used to compare refactoring operations.
-
-It's worth noting that this approach depends on having effective oracles for assessing the results.
-It proved to be more complicated to commission the Jureczko data sets for this study.
-For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could be salvaged using the techniques discussed in this section.
-
-\fig{j} shows our preliminary studies with the Jureczko data.
-Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}, this means that we are training on hundreds to thousands of classes and testing on smaller test suites).
-Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
-In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
-
-In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
-For such data, quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
-``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
-``pf''): the percent of non-fault classes that are {\em predicted} to be defective.
-
-As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was built from the training data, then applied to the test data.
-The ``untuned'' columns of \fig{j} shows those results.
-We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
-For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
-With this threshold, however, none of our data sets were be suitable for this study.
-
-Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
-Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples and an under-sampling of majority-class examples.
-Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
-The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
-For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
-However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
-
-In summary, while we cannot trust predictors from some of our Jureczko data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-Accordingly, when this study explores the Jureczko data, we will use these five data sets.
-
-(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
 

### 2016-03-08 tex/refactor.tex
Updates from ShareLaTeX

@@
 \centering
-\begin{tabular}{r|l|c|c|c|c}
-Mostly from Fowler'99& & & & & Developer\\
- (with a few additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey\\\hline
-  Alt. Classes with Diff. Interfaces & & & & + & \\
-  Combinatorial Explosion~\cite{Kerievsky2004} & & & & + & \\
-  Comments & & & 11 & + & VL\\
-  Conditional Complexity~\cite{Kerievsky2004} & & & 14 & + & ?\\
-  Data Class & Data Class & & & + &\\
-  Data Clumps &  &  & & + &\\
-  Divergent Change & & & & + & \\
-  Duplicated Code & Significant Duplication\textsuperscript{*} & + & 1 & + & VH\\
-  Feature Envy & Feature Envy & & 8 & + &\\
+\begin{tabular}{r|c|c|c|c|c}
+Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
+ (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
+  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
+  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
+  Comments & & & 11 & \checkmark & VL\\
+  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
+  Data Class & \checkmark & & & \checkmark &\\
+  Data Clumps &  &  & & \checkmark &\\
+  Divergent Change & & & & \checkmark & \\
+  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
+  Feature Envy & \checkmark & & 8 & \checkmark &\\
 
-  Inappropriate Intimacy & & - & & + & L\\
-  Indecent Exposure~\cite{Kerievsky2004} & & & & + & ?\\
-  Incomplete Library Class & & & & + &\\
-  Large Class & God Class/Brain Class & ++ & 4 & + & VH\\
-  Lazy Class/Freeloader & & - & 7 & + &\\
-  Long Method & Brain Method & ++ & 2 & + & VH\\
-  Long Parameter List &  & + & 9 & + & L \\
+  Inappropriate Intimacy & & \checkmark & & \checkmark & L\\
+  Indecent Exposure~\cite{Kerievsky2005} & & & & \checkmark & ?\\
+  Incomplete Library Class & & & & \checkmark &\\
+  Large Class & \checkmark & \checkmark & 4 & \checkmark & VH\\
+  Lazy Class/Freeloader & & \checkmark & 7 & \checkmark &\\
+  Long Method & \checkmark& \checkmark & 2 & \checkmark & VH\\
+  Long Parameter List &  & \checkmark & 9 & \checkmark & L \\
 
-  Message Chains & & & & + & H\\
-  Middle Man & &  & & + &\\
-  Oddball Solution~\cite{Kerievsky2004} & & & & + & \\
-  Parallel Inheritance Hierarchies & & & & + &\\
-  Primitive Obsession &  & & & + &\\
+  Message Chains & & & & \checkmark & H\\
+  Middle Man & &  & & \checkmark &\\
+  Oddball Solution~\cite{Kerievsky2005} & & & & \checkmark & \\
+  Parallel Inheritance Hierarchies & & & & \checkmark &\\
+  Primitive Obsession &  & & & \checkmark &\\
 
-  Refused Bequest & Refused Parent Bequest & - & & + & \\ 
-  Shotgun Surgery & Shotgun Surgery & & & + & \\
-  Solution Sprawl~\cite{Kerievsky2004} & & & & + &\\
-  Speculative Generality & & & & + & L\\
-  Switch Statements &  & - & & & L\\
-  Temporary Field & & - & & + & ?\\
+  Refused Bequest & \checkmark & \checkmark & & \checkmark & \\ 
+  Shotgun Surgery & \checkmark& & & \checkmark & \\
+  Solution Sprawl~\cite{Kerievsky2005} & & & & \checkmark &\\
+  Speculative Generality & & & & \checkmark & L\\
+  Switch Statements &  & \checkmark & & & L\\
+  Temporary Field & & \checkmark & & \checkmark & ?\\
   \end{tabular}
@@
 
-\caption{List of bad code smells from different sources. Empty cells
-denote some bad smell proposed by Fowler'99 that was not found relevant
-in subsequent studies. Note that there are many blank cells.}
+\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
+Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
+``?'' indicates lack of consensus on some bad smell. Empty cells
+denote some bad smell listed in column one that was not found relevant
+in subsequent studies.
+Note that there are many blank cells.}
 \label{fig:smells}

### 2016-03-09 Manuscript.tex
Updates from ShareLaTeX

@@
 When developers cannot fully assess the value of bad smells,
-it is possible to use automatic means. We assume that
-developers will change code until a bad smell detector stops triggering.
-That is, if the bad smell detector is {\em loc} $>$ 100, they will
-redistribute the functionality of a long method into several smaller ones.
-Recalling the italics in the above quote, Fowler's criteria for successful bad smell removal is the extent
-to which the reorganized code has fewer errors/defects.
-Data miners can be used to predict the defects in the reorganized code
-so, in this paper, we will use static code defect predictors to assess
-bad smells.
-
-After making that assessment, this paper finds
-that   state-of-the-art methods
-(Shantnawi~\cite{Shatnawi10}  in TSE 2010;
-and our own instance-based systems~\cite{me12c} from 2012) 
-are less effective for bad smell assessment than a new
-contrast set learning method called XTREE.  
-Standard learners, such as the regression methods used by Shantnawi~\cite{Shatnawi10}
-only comment on current distributions.  
+it is possible to use automatic means.
+
+This paper proposes the XTREE contrast set learner for recognizing
+``bad'' bad smells. 
 Contrast set learners   recommend {\em what to change} in order to move from
@@
 i.e. a code reorganization. Hence, constrast set learners
-are a better way to learn when code needs to be reorganized
-than, say, the methods of Shantnawi~\cite{Shatnawi10}.
-
-We recommend using XTREE as a judge to rank the value of a set
-of bad smell detectors:
+like XTREE can be used to  recognize ``bad''
+bad smells:
 \begin{itemize}
@@
 truly causal.
-That said, if someone else's proposed bad smell detector $S_{\mathit{bad}} $ falls {\em outside} of the space of
-effects found  using XTREE's analysis, then we can say that XTREE found 
-{\em neither} correlation {\em or}
-causal evidence that $S_{\mathit{bad}} $ can be found in the historical record.
+
+XXX
+
+Nevertheless, XTREE is a still a useful tool to critique and prune
+irrelevant bad smells since, 
+as show below, many of the bad smell detectors offered
+in the literature (and supported by readily available tools)
+fall outside of  space of
+effects found  using XTREE's analysis.
+That is,   XTREE can find 
+{\em neither} correlation {\em nor}
+causal evidence that those bad smells
+have any support   in the historical record.
+
+XXXX
+
+The same cannot be said for other methods that try to learn
+bad smell thresholds. \tion{outs}
+discusses {\em outlier} methods and {\em centroid delta} methods
+for learning bad smell thresholds.
+The {\em  outlier methods}  used by
+Shatnawi, Alves and Herman et al.~\cite{Shatnawi10,Alves2010,hermans15}
+propose thresholds on all code metrics with outlier values.
+Similarly, the older {\em centroid delta} methods  used in our 
+prior work~\cite{me12c} also returns thresholds on most of the code
+metrics. That is, outlier methods and centroid deltas methods
+seem to be saying that many bad smell thresholds are possible.
+
+we proposed CD, short for
+``Centroid Deltas'', as a way to find bad smell thresholds:
+\begin{itemize}
+    \item Clustered
+data and generated centroids for each cluster;
+\item Computed contrast sets (i.e. the bad smells)
+as the delta between attributes in adjacent centroids $C_1 - C_2$
+where examples in $C_1$ had more defects than $C_2$.
+\end{itemize}
+As shown below, Centroid Deltas
+generates change
+recommendations that are far less effective than XTREE.
+Worse still Centroid Deltas reported bad smells
+for far more attributes than XTREE. Hence, that prior work
+is less useful than XTREE at judging the  bad smells proposed by developers.
 
@@
 Secondly, XTREE assumes   Fowler's bad smells
-should be assessed via their contribution to software defects.   Some researchers 
+should be assessed via their contribution to software defects
+(evidence: see the section in italics of the quote
+that starts this introduction). Some researchers 
  comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
@@
 
+\newpage
 The contributions of this paper are
@@
     \item A new kind of contrast set learner called XTREE
-    that can critique plans of what to change within a data set in order
+    that   critique plans of what to change   in order
     to improve outcomes.
@@
     to build an oracle for the proposed changes.
-    \item Using that learner and that oracle, we offer a method for avoid code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
-    \item A resolution between proponents of bad smells (e.g.
+    \item Using that learner and that oracle,
+    a demonstration that XTREE is better than our prior approach~\cite{me12c}
+    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
+    at avoiding code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
+    \end{itemize}
+    Based on the above, we can offer
+    a resolution between proponents of bad smells (e.g.
 Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}).
@@
     first carefully assessed.
-\end{itemize}
 
@@
 Human developers can be clever, but their thinking can also be distorted
-by congitive biases.
-Hence it is important to double check their beliefs in, for example,
-bad smells. Tools like XTREE perform that check, using models learned
-from the historical record.
+by cognitive biases.
+Hence, as shown in \fig{smells}, developers and text books and tools 
+can disagree on which bad smells are important.
+Hence it is important to have some automated tool
+that can assess their beliefs in, for example,
+bad smells.
+
+\section{Learning Bad Smell Thresholds}
+
+After a review of the literature, we reached the same conclusion
+as Shatnawi~\cite{Shatnawi10}; i.e. there is a dearth of papers recommending specific thresholds
+for bad smells. In the recent literature (since 2010), we could
+only find the handful   papers discussed below.
+
+In the recent literature, there are two approaches for
+learning bad smell thresholds:
+\begin{itemize}
+    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
+    as used by Shatnawi~\cite{shatnawi10}, Alvenes et al.~\cite{Alves2010}
+    and Hermans et al.~\cite{hermans15};
+    \item  Approaches based on {\em cluster deltas} that we developed
+    for   Centroid Deltas~\cite{me12c} and XTREE.
+\end{itemize}
+The outlier approach assume that unusually large measurements are risk-prone.
+Hence, they generate one bad smell threshold for any metric
+with such an unusually large measurements.
+
+The advantage of outlier
+methods is that they are simple to implement. However,
+their disadvantage are two-fold. 
+Firstly, the outlier approach is {\em verbose}; i.e.  it can
+propose thresholds on most of the code metrics.
+Such verbosity makes it hard   to critique and prune away proposed code reorganizations since any bad smell mentioned by a developer has some support
+somewhere amongst all those proposed thresholds.
+
+Secondly, they do not reflect
+on sets of code metrics that change together. For example,
+one way to remove unusually large methods is to reorganize that method's
+functionality into several smaller ones. If that is done,
+this means that {\em reducing} lines of code in one method
+will {\em increase} the coupling of the methods of that class to
+either itself or to other classes (depending on whether or not
+the new smaller methods are placed in the original class, or in other classes).
+Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
+large lines of code {\em and} seemingly excessive coupling.
+
+Borges and Menzies first proposed the centroid delta approach to address
+this second problem. Centroid deltas assume...
+
+One is {\em succinct}; i.e. comments
+on too many code metrics. If the
+
+class' coupling mea
+\begin{itemize}
+    \item 
+\end{itemize}that they do not consider the associations
+between variables.
+
+Cluster deltas methods assume that at
+Prior work on deriving metric thresholds from benchmark data
+used {\em outlier} statistics to identify abnormally large ranges. 
+Another approach, used by
+
+For example,
+one of the first methods of finding thresholds was proposed Erni and Lewerentz~\cite{erni96}.
+Their technique to identify thresholds was based on the data distribution, specifically the mean and the standard deviation of the metric values.
+They propose the use of values that lie beyond one standard deviation from the mean as a threshold.
+The minimum value $T_{min}$ is given by $T_{min}=\mu-\sigma$, and this is used when metric definition considers very small values as an indicator of problems.
+Otherwise, $T_{max}=\mu+\sigma$ is used, when large metric values are considered problematic.
+
+Several researchers~\cite{shatnawi10}~cite{alves10} have pointed out that this method is subject to a few problems.
+Firstly, it doesn't consider the fault-proneness of classes when the thresholds are computed.
+Secondly, there is a lack of empirical validation of this methodology, which impedes reasonable comparisons.
 
@@
 
-\begin{figure*}[tbp!]
+\begin{figure*}[!y]
 \centering
-\includegraphics[width=0.8\linewidth]{figs/changes0.png}
+\includegraphics[width=0.90\linewidth]{figs/changes0.png}
 \caption{Suggested Changes}

### 2016-03-09 Manuscript.tex
Updates from ShareLaTeX

@@
 
-\title{  How to Recognize  a Bad  ``Bad Smell'' (using Contrast Sets)}
+\title{  How to Recognize  a Bad  ``Bad Smell'' using Contrast Sets}
 \numberofauthors{2} 
@@
        \affaddr{Fraunhofer CESE, College Park, USA}\\ 
-       {llayman@fc-md.umd.edu}
+       {llayman@cese.fraunhofer.org}
 \setlength{\columnsep}{7mm}}
@@
 It is hard to assess
-if a bad smell is ``bad''; i.e. if it is irrelevant or misleading
+if a ``bad smell'' is irrelevant or misleading
 for a specific project.
@@
 Specifically, it is unclear at what threshold values
-for which metrics should trigger code reorganization since
+for which metrics should trigger code reorganization\footnote{From LL: The term reorganization is a bit weird.
+Bad code smells are ``refactored''.
+Do you have a reason that you are using reorganization instead of refactoring throughout the paper?
+Recall that Fowler's code smell book is called ``Refactoring''} since
 most studies do not offer
-metrics {\em and} a threshold for their proposed bad smells.
+both metrics {\em and} a threshold for their proposed bad smells.
 This paper proposes the XTREE contrast set
@@
 like XTREE to stop work on useless code reorganizations.
-Further, using XTREE, it is possible to show that many
-of the bad smells endorsed by standard texts or tools
+Further, XTREE shows that many
+of the bad smells defined in standard texts and tools
 are not useful for particular projects. 
 Based on these results, we endorse  
-guiding code reorganization via bad smells. However,  
-it is important to XTREE the bad smells defined for a 
-particular project.
+guiding code reorganization via bad smells, but one must empirically validate the importance of specific bad smells within the context of each project using an approach like XTREE.
 
@@
 bad smells. 
-XTREE uses frequency counts of code features seen in the historical record
+XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record.
+A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
 to generate
@@
 causal evidence that those bad smells
-have any support   in the historical record.
+have any support   in the historical record.\footnote{From LL: You need to say what ``support from the historical record'' is early in the paper.
+Are you mapping code smells in source files to defects?
+Code smells to performance regressions?
+Whatever it is, give an example.}
 
@@
 
-\newpage
 The contributions of this paper are
@@
 
+%\input{tex/refactor}
+
 \section{Motivation}\label{sect:prelim}
@@
 relate specifically to bad smells:
-\input{tex/refactor}
+
 \begin{enumerate} 
@@
 
+\input{tex/jur_char}
+
 A  blank cell in \fig{smells}
@@
 \section{Learning Bad Smell Thresholds}
-
-After a review of the literature, we reached the same conclusion
-as Shatnawi~\cite{Shatnawi10}; i.e. there is a dearth of papers recommending specific thresholds
-for bad smells. In the recent literature (since 2010), we could
-only find the handful   papers discussed below.
-
-In the recent literature, there are two approaches for
+The SE literature offers  two ways to 
 learning bad smell thresholds:
@@
     \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
-    as used by Shatnawi~\cite{shatnawi10}, Alvenes et al.~\cite{Alves2010}
+    as used by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
     and Hermans et al.~\cite{hermans15};
@@
 \end{itemize}
+\subsection{  Outliers}
+
 The outlier approach assume that unusually large measurements are risk-prone.
 Hence, they generate one bad smell threshold for any metric
-with such an unusually large measurements.
-
-The advantage of outlier
-methods is that they are simple to implement. However,
-their disadvantage are two-fold. 
-Firstly, the outlier approach is {\em verbose}; i.e.  it can
+with such an ``unusually large'' measurements. 
+The literature lists several ways to define ``unusually large''.
+
+\subsubsection{Enri \& Lewerentz}
+One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
+Given classes described with the  code metrics of \fig{ck},
+Enri and Lewerentz finds the   mean $\mu$ and the standard deviation $\sigma$
+of each
+code metrics. Their definition of problematic outlier was any code
+metric with a measurement greater than $\mu+\sigma$.
+
+Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
+using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, there is a lack of empirical validation of this method.
+
+\subsubsection{ Shatnami}
+Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
+is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
+thresholds which, following Shatnami's guidance, we set to
+$p_0=p_1=0.05$. VARL encodes the defect count
+for each class to 0 (for no defects known) to 1 (for defects known).
+Univariate binary logistic regression is applied to learn three coefficients: $\alpha, \beta,p_0$ being
+the  constant intercept ($\alpha$), the coefficient for maximizing log-likelihood function ($\beta$),
+and a measure of how well this regression model predicts for the defects ($p_0$).
+Any code metric with $p_0>0.05$ is then rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
+the equation proposed by Bender:
+\begin{equation}
+ \mathit{bad\; smell\; if\;} M_c > \frac{1}{\beta }\left( {\log \left( {\frac{{{p_1}}}{{1 - {p_1}}}} \right) - \alpha } \right) 
+\end{equation}
+\subsubsection{ Alves et al.}
+Alves et al.~\cite{Alves2010} propose another approach
+to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
+Evey metric value is weighted according to the source lines of code (LOC) of the class.
+All the weighted metrics are then normalized; i.e.
+they are divided by the sum of all weights of the same system.
+As with Shatnawi, univariant regression
+is then applied to reject code metrics that are poor predictors of defects (i.e. all those  with $p > 0.05$).
+Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
+Alves et al. then compute their 
+thresholds are then derived by choosing the percentage of the overall code that needs to be represented.
+For instance, Alves et al.
+suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric.
+The intuition here is that the worst
+code are the outliers beyond 90\% of the normal code measurements.
+
+Our reading of the literature is that the Alves et al.  
+approach is preferred to that of Shatnami. For example, Herman et al.~\cite{hermans15} used the
+Alves et al. method in their  2015 paper on
+exploring bad smells.
+
+\subsubsection{Discussion of the Outlier Approach}
+The advantage of the outlier
+approach is that it is simple to implement. 
+For example, in less than a day we could
+code up and apply all the approaches of Shatnami and 
+Alves et al.
+
+On the other hand, these outlier approaches have major two disadvantages. 
+Firstly, they are {\em verbose}; i.e.  they can
 propose thresholds on most of the code metrics.
-Such verbosity makes it hard   to critique and prune away proposed code reorganizations since any bad smell mentioned by a developer has some support
+Such verbosity makes it hard   to critique and prune bad smells
+since any bad smell mentioned by a developer has some support
 somewhere amongst all those proposed thresholds.
 
-Secondly, they do not reflect
-on sets of code metrics that change together. For example,
-one way to remove unusually large methods is to reorganize that method's
-functionality into several smaller ones. If that is done,
+Secondly, the outlier approaches described above are all 
+{\em univariant} which means
+they do  not understand
+that sets of code metrics must
+be changed at the same time.
+For example:
+\begin{itemize} 
+\item One way to remove unusually large methods is to reorganize that method's
+functionality into several smaller ones. 
+\item If that is done,
 this means that {\em reducing} lines of code in one method
@@
 the new smaller methods are placed in the original class, or in other classes).
+\item
 Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
-large lines of code {\em and} seemingly excessive coupling.
-
-Borges and Menzies first proposed the centroid delta approach to address
-this second problem. Centroid deltas assume...
-
-One is {\em succinct}; i.e. comments
-on too many code metrics. If the
-
-class' coupling mea
+large lines of code {\em and} seemingly excessive coupling since
+one cannot be reduced without increasing the other.
+\end{itemize}
+\subsection{Cluster Deltas}
+Borges and Menzies first proposed the centroid delta approach to avoid the drawbacks of univariant reasoning.
+Centroid deltas are a {\em multivariant} method which
+work  as follows:
 \begin{itemize}
-    \item 
-\end{itemize}that they do not consider the associations
-between variables.
-
-Cluster deltas methods assume that at
-Prior work on deriving metric thresholds from benchmark data
-used {\em outlier} statistics to identify abnormally large ranges. 
-Another approach, used by
-
-For example,
-one of the first methods of finding thresholds was proposed Erni and Lewerentz~\cite{erni96}.
-Their technique to identify thresholds was based on the data distribution, specifically the mean and the standard deviation of the metric values.
-They propose the use of values that lie beyond one standard deviation from the mean as a threshold.
-The minimum value $T_{min}$ is given by $T_{min}=\mu-\sigma$, and this is used when metric definition considers very small values as an indicator of problems.
-Otherwise, $T_{max}=\mu+\sigma$ is used, when large metric values are considered problematic.
-
-Several researchers~\cite{shatnawi10}~cite{alves10} have pointed out that this method is subject to a few problems.
-Firstly, it doesn't consider the fault-proneness of classes when the thresholds are computed.
-Secondly, there is a lack of empirical validation of this methodology, which impedes reasonable comparisons.
+    \item Cluster data about code modules;
+    \item Find
+neighboring clusters $C_+,C_-$ where $C_+$ has more examples of defective
+modules than $C_-$;
+\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
+\item $\Delta$ are changes needed in defective modules of $C_+$ to
+      make them more like the less-defective modules of $C_-$
+\end{itemize}
+Note that $\Delta$ is a multivariant recommendation
+that proposes multiple changes in order to reduce defects.
+Since it is computed
+from neighboring clusters, it   reasons about
+examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
+repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
+without also increasing a coupling measure.
+
+The drawback with with CD is that it is very {\em verbose}
+since
+CD   recommended changes to all code
+metrics with different values in those two centroids. 
+This makes it hard to use CD to   critique and prune away bad smells.
+\subsubsection{XTREE}
+
+  XTREE  is cluster delta algorithm
+  that avoids the problem of verbose $\Delta$s.
+  Instead of reasoning over cluster centroids,
+  XTREE utilizes a decision tree learning approach
+  to find the fewest differences groupings of examples.
+
+  Decision trees are built via
+  {\em iterative dichotomization} (ID).
+  ID   finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
+  and ID is called
+  recursively on the data in each split. In this way,
+  each split is the root of a sub-tree.
+
+  The CART iterative dichotomization algorithm
+  measures the variability of continuous attribites
+  using standard deviation 
+  $\sigma_x =\sqrt{ \sum^n_{i=1} \left(  \frac{(x_i - \bar{x})^2}{n-1}\right)$.
+  If a code metric is split into $r$ ranges and each range is of
+  size
+    $n_r$ is
+  associated with a set of defect counts  with standard deviation
+  $\sigma_r$ then the best split is the one that minimizes
+  $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).
+
+  \begin{figure}[t!!]
+\centering
+\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
+\caption{Sample plans from XTREE. This
+tree can be read like a nested if-then-else statement. For
+example, lines 3 and 8 show two branches for lines of code
+(denoted here as {\em \$loc}) below 698 and above 698.
+Any line with a colon ":" character  shows
+a leaf of this nesting. For example, if some new code module
+is passed down this tree and 
+falls to the  line marked in orange,
+the colon on that line indicates a 
+prediction that this module has a 100\% chance of being defective.
+XTREE looks for a nearby branch that has a lower chance of being
+defective.
+Finding the green branch, XTREE a bad smell threshold
+for that module that is the delta between 
+the orange branch and the green branch. In this case, that threshold
+relates to 
+lines of code and comments ({\em lcom}) and the cohesion between 
+classes ({\em cam}:   measures similarity of
+parameter list to assess the relatedness among the methods of class).
+\label{fig:xtree_samp}
+\end{figure}
+
+  While not usually labelled as such, ID can
+  be viewed as a clustering algorithm that groups together code modules
+  with similar defect counts and some shared ranges of code metrics.
+  XTREE is a post-processor to ID that reports the delta between 
+  nearby groups with different defect counts. \fig{xtree_samp} offers
+  a small example of that reflection. In this example:
+  \begin{itemize}
+      \item Some project has a table of data describing its  classes in the form of \fig{ck}.
+      \item After code inspections and running test cases or operational
+      tests, each such class is augmented with a defect count.
+      \item CART takes that table of data and, using iterative dichotomization,
+      it generated the tree of \fig{xtree_samp}.
+      \item Once the tree is built, some test case arrives and is passed
+      down the tree to a leaf note (see the orange line in \fig{xtree_samp}.
+      \item XTREE then looks for a nearby leaf node with a lower defect
+      count (see the green line in \fig{xtree_samp}. XTREE bad smell
+      thresholds for that method is then the difference between 
+      green and orange.
+  \end{itemize}
+
+\subsubsection{XTREE: the details}
+
+ {\em Step1:} Using the training data,  XTREE uses iterative dichotomization to
+  divide a table of $N$ code modules  into  groups of
+size $\alpha=\sqrt{N}$.
+
+{\em Step2:} For each test code module, 
+	  find the {\em current } leaf: take each test , run it down to a leaf in the decision tree.  
+After that,	  find the {\em desired} leaf:
+		\begin{itemize}[leftmargin=3mm]
+		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
+		\item Identify {\em sibling} leaves; i.e. leaf clusters that can be reached from level $lvl$ that are not same as {\em current }
+		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\gamma=0.5$ times the mean score of {\em current}. 
+		   If none found, then repeat for $lvl += 1$. Also,
+		    return no plan if the new $lvl$ is above the root. 
+		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
+		\ei
+	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta: 
+	 \begin{itemize}
+	     \item For discrete attributes, delta is the value from {\em desired};
+	     \item For  numerics, delta is the numeric difference;
+	     \item For numerics  discretized into ranges, delta is a random number selected from the low and high boundaries of the that range.
+	     \end{itemize}
+	     The bad smell threshold for this module is the delta that would move
+	     it from the current to desired lead.
 
@@
 the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
-
-\input{tex/jur_char}
 
@@
 \section{Threats to Validity}\label{sect:valid}
+
+XXXX only reorg due to the issues in the data sets.
+we are stuck with that. note hwever that if
+projects improved their data collection then
+our techniques woudl still apply, unmodifed, to the revised data sets. so here we are ding what we can with the data currently available (and also ensuring we are good to go on future data)
 

### 2016-03-09 tex/refactor.tex
Updates from ShareLaTeX

@@
 Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
- (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
+ (with a some additions~\cite{Kerievsky2005}) & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
   Alt. Classes with Diff. Interfaces & & & & \checkmark & \\

### 2016-03-10 Manuscript.tex
Updates from ShareLaTeX

@@
 \documentclass{sig-alternate}
+\usepackage{stfloats}
 \usepackage{tikz}
@@
 \newcommand{\fig}[1]{Figure~\ref{fig:#1}}
-\newcommand{\eq}[1]{Equation~\ref{ 0eq:#1}}
+\newcommand{\eq}[1]{Equation~\ref{eq:#1}}
 %\setlist{nolistsep,leftmargin=5mm}
@@
 Specifically, it is unclear at what threshold values
-for which metrics should trigger code reorganization\footnote{From LL: The term reorganization is a bit weird.
-Bad code smells are ``refactored''.
-Do you have a reason that you are using reorganization instead of refactoring throughout the paper?
-Recall that Fowler's code smell book is called ``Refactoring''} since
+for which metrics should trigger code reorganization  since
 most studies do not offer
@@
 In that approach, developers are always asking themselves ``should I reorganize this code''?
-Bad smells are a widely used  technique for deciding when to reorganize\footnote{Fowler textbook is widely read-- as of  early 2016, ACM Digital Portal reports 
+Bad smells are a widely used  technique for deciding when to reorganize\footnote{The Fowler text is widely read-- as of  early 2016, ACM Digital Portal reports 
 it has 6434 references. Also, bad smell-based research
@@
 
-%\input{tex/refactor}
+\input{tex/refactor}
 
@@
 they are divided by the sum of all weights of the same system.
-As with Shatnawi, univariant regression
-is then applied to reject code metrics that are poor predictors of defects (i.e. all those  with $p > 0.05$).
+For consistency
+with Shatnawi, we explore the correlation between
+the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
 Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
@@
 exploring bad smells.
+
+\begin{figure*}[!b]
+~\hrule~
+\begin{minipage}{.53\linewidth}
+\small
+ \begin{tabular}{p{0.95\linewidth}} 
+ On the right-hand-side is a tree
+ generated by iterative dichomization. 
+ This tree can be read like a nested if-then-else statement; e.g.
+ \begin{itemize}
+     \item Lines 3 and 8 show two branches for lines of code (denoted here as {\em `\$loc}) below 698 and above 698.
+     \item Any line with a colon ":" character shows  a  leaf  of  this  nesting.
+For  example,  if  some  new  code module is passed down this tree and falls to the line marked in \textcolor{orange}{{\bf orange}},  the colon on that line indicates a prediction that this module has a 100\% chance of being defective.
+     \end{itemize}
+Using this tree,
+ XTREE looks for a nearby branch that has a lower chance of being defective.
+Finding the \textcolor{green}{{\bf green}} desired  branch,  XTREE reports a bad smell threshold for that module that is the delta between the \textcolor{orange}{{\bf orange}}  current branch  and   \textcolor{green}{{\bf green}} designed branch.
+ In this case, that threshold relates to:
+ \begin{itemize}
+     \item Lines of code and comments ({\em lcom}) 
+     \item The cohesion between classes ({\em cam}) which measures similarity of parameter lists to assess the relatedness amongst class   methods.
+     \end{itemize}\\ 
+ \end{tabular}
+ \end{minipage}~~~~
+ \begin{minipage}{.45\linewidth}
+\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
+\end{minipage}
+~\hrule~
+ \caption{A brief tutorial on XTREE.} \label{fig:xtree_samp}
+\end{figure*}
 
@@
 \end{itemize}
+
 \subsection{Cluster Deltas}
-Borges and Menzies first proposed the centroid delta approach to avoid the drawbacks of univariant reasoning.
+
+\subsubsection{CD}
+Borges and Menzies first proposed the CD centroid delta approach to avoid the drawbacks of univariant reasoning~\cite{me12c}.
 Centroid deltas are a {\em multivariant} method which
@@
 
-  XTREE  is cluster delta algorithm
+  XTREE  is a cluster delta algorithm
   that avoids the problem of verbose $\Delta$s.
@@
   measures the variability of continuous attribites
-  using standard deviation 
-  $\sigma_x =\sqrt{ \sum^n_{i=1} \left(  \frac{(x_i - \bar{x})^2}{n-1}\right)$.
+  using the  standard deviation $\sigma$ of the defect counts.
   If a code metric is split into $r$ ranges and each range is of
@@
     $n_r$ is
-  associated with a set of defect counts  with standard deviation
+  associated with a set of defect counts $x$ with standard deviation
   $\sigma_r$ then the best split is the one that minimizes
   $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).
-
-  \begin{figure}[t!!]
-\centering
-\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
-\caption{Sample plans from XTREE. This
-tree can be read like a nested if-then-else statement. For
-example, lines 3 and 8 show two branches for lines of code
-(denoted here as {\em \$loc}) below 698 and above 698.
-Any line with a colon ":" character  shows
-a leaf of this nesting. For example, if some new code module
-is passed down this tree and 
-falls to the  line marked in orange,
-the colon on that line indicates a 
-prediction that this module has a 100\% chance of being defective.
-XTREE looks for a nearby branch that has a lower chance of being
-defective.
-Finding the green branch, XTREE a bad smell threshold
-for that module that is the delta between 
-the orange branch and the green branch. In this case, that threshold
-relates to 
-lines of code and comments ({\em lcom}) and the cohesion between 
-classes ({\em cam}:   measures similarity of
-parameter list to assess the relatedness among the methods of class).
-\label{fig:xtree_samp}
-\end{figure}
 
@@
       \item CART takes that table of data and, using iterative dichotomization,
-      it generated the tree of \fig{xtree_samp}.
+        generates the tree of \fig{xtree_samp}.
       \item Once the tree is built, some test case arrives and is passed
-      down the tree to a leaf note (see the orange line in \fig{xtree_samp}.
+      down the tree to a leaf note (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}.
       \item XTREE then looks for a nearby leaf node with a lower defect
-      count (see the green line in \fig{xtree_samp}. XTREE bad smell
-      thresholds for that method is then the difference between 
-      green and orange.
+      count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that test case, XTREE proposes a bad smell
+      thresholds that is  the difference between 
+      \textcolor{green}{{\bf green}} and \textcolor{orange}{{\bf orange}}.
   \end{itemize}
@@
 
- {\em Step1:} Using the training data,  XTREE uses iterative dichotomization to
+Firslty, using the training data,  XTREE uses iterative dichotomization to
   divide a table of $N$ code modules  into  groups of
@@
 
-{\em Step2:} For each test code module, 
+Next, for each test code module, 
 	  find the {\em current } leaf: take each test , run it down to a leaf in the decision tree.  
 After that,	  find the {\em desired} leaf:
-		\begin{itemize}[leftmargin=3mm]
+		\begin{itemize}
 		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
@@
 		   If none found, then repeat for $lvl += 1$. Also,
-		    return no plan if the new $lvl$ is above the root. 
+		    return nil  if the new $lvl$ is above the root. 
 		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
 		\ei
-	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta: 
+	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta,
+	 for discrete attributes, delta is the value from {\em desired};
+	   for  numerics, delta is the numeric difference;
+	   and  for numerics  discretized into ranges, delta is a random number selected from the low and high boundaries of the that range. 
+	     The bad smell threshold for   is then the delta that would move
+	     the test case from the current to desired leaf.
+
+\section{Evaluation}
+\subsection{Research Questions}
+
+Theoretically, XTREE is a good way to learn bad smell thresholds:
 \begin{itemize}
-	     \item For discrete attributes, delta is the value from {\em desired};
-	     \item For  numerics, delta is the numeric difference;
-	     \item For numerics  discretized into ranges, delta is a random number selected from the low and high boundaries of the that range.
+    \item Unlike the outlier approach, XTREE offers bad smell
+    thresholds that are specific to where test cases fall into
+    the leaves of a decision tree that predicts for defects.
+    That is, while the outlier approach offers one threshold for all
+    metrics, XTREE specializes those thresholds to particular regions of
+    the code.
+    \item
+    Unlike the outlier approach, XTREE is a multivariant approach
+    that  understands when several code
+metrics should  be changed in unison. 
+\item Unlike CD, it is less verbose. 
+XTREE does not report all the deltas between
+adjacent clusters. Rather, it just reports the deltas in the code
+metrics that a decision tree learner has determined are the minimun
+changes required to distinguish one grouping from another.
 \end{itemize}
-	     The bad smell threshold for this module is the delta that would move
-	     it from the current to desired lead.
-
-\section{Oracles}
+Both these claims need to be tested, which leads to the following research
+questions.
+
+{\bf RQ1: Effectiveness.} {\em In a comparison of  the outlier and cluster delta
+methods described above, which one reduces defects the most?}
+
+To find an answer to this question, we will assume that developers
+reorganize their code until the bad smell thresholds no longer apply.
+This reorganization will start with some {\em initial} code
+base that is changed to a {\em new} code base. 
+For example, if the bad smell is {\em loc $>100$} and a 
+code module has (e.g.) 500 lines of code, we conduct a ``what-if''
+query and create a {\em new} code base with that module having {\em loc=100}.
+Using a defect
+predictor learned from static code metrics, we then predict the
+number of defects in $d_0,d_1$ in {\em initial} and {\em new},
+respectively. In this experiment,
+the best bad smell threshold detector is the one that maximizes
+\begin{equation}\label{eq:diff}
+\mathit{improvement} = 100* \left(1 - \frac{ d_1 }{ d_0}\right)
+\end{equation}
+
+{\bf RQ2: Verbosity.}   {\em In a comparison of  the outlier and cluster delta
+methods described above, which one proposes the fewest thresholds?}
+
+Note that this is important since our goal is to critique and, possibly,
+reject irrelevant bad smell detectors. As mentioned above,
+  verbosity makes it hard   to critique and prune bad smells
+since any bad smell mentioned by a developer has some support
+somewhere amongst all those proposed thresholds.
+
+{\bf RQ3: Co-change.}   {\em Does XTREE respect code
+metrics that need to be changed together? }
+
+For example, in XTREE's thresholds,  if {\em loc} is
+reduced, does some coupling measure increase?
+
+\input{tex/prediction_stats.tex}
+
+\subsection{Test Data}\label{sect:tesd}
+
+To explore these research questions,
+we used data from
+the Jureczko et al.'s collection of object-oriented JAVA systems~\cite{jureczko10}. To access that data, goto   git.io/vGYxc.
+
+The Jureczko data records the number of known defects for each class using a post-release bug tracking system.
+The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc.
+For details on the Jureczko code
+metrics, see  \fig{ck}. For details on the rows and versions
+of that data, see the left-hand-side columns of \fig{j}.
+
+\subsection{Statistics}
+Three of our methods are deterministic (Shatnami, Alves et al.,
+and XTREES) while the CD uses some stochastic processing
+in its clustering (CD uses the  WHERE clusterer~\cite{Menzies2013:local}). Hence, to compare all
+four methods, we must run the analysis many times. 
+We use  40 repeated runs, each with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
+Each run collects the \fig{diff} values
+uses defect predictions make for the {\em new} and {\em initial} code.
+
+To rank these 40 numbers collected from CD, XTREE,Shatnami, and Alves et al.), we  use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}. 
+In accordance to that test, using the median values of each method, we
+sort a list of  $l=40$ values of \eq{diff} values found in  $ls=4$ different methods. 
+Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.
+E.g.
+for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$: \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
+We then apply a apply a statistical hypothesis test $H$ to check
+if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping).
+If so, Scott-Knott recurses on the splits.
+In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
+For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
+These researchers warn that even if a hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
+
+The Scott-Knott  results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
+The black dot shows the median \eq{diff} values and the horizontal likes stretch 
+from the 25th percentile to the 75th percentile (a region called the inter-quartile
+range, or IQR).
+\begin{center}
+{\small
+  Example 
+\begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
+\arrayrulecolor{lightgray}
+\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
+1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
+\hline 2 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
+2   & Alves        &    40  & 20  & \quart{30}{10}{50}{10}\\
+\end{tabular}}
+\end{center}
+In this example table, the rows are  sorted on the median values of each method.
+Note that all the methods have \eq{diff} $\gt 0\%$; i.e.
+all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 62\% from the original value).
+The above example table has a  left-hand-side  {\bf Rank} column, computed using the
+Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and Alves  are ranked the worst.
+
+\subsection{Orcales}
 \label{sect:eval}
-\input{tex/prediction_stats.tex}
-
-As highlighted previously, there are several detection techniques for anti-patterns, yet these fail to address the issue of checking the proposed changes for their efficacy.
-Much of the refactoring needs to evaluated for their usefulness.
-To this end, there needs to be an oracle, a procedure that can be employed to distinguish between good and bad options.
-The dearth oracles is not unique to this domain, for instance a recent survey by Barr et al \cite{barr15}
-
-Our strategy (discussed in greater detail below) divides this data into a training and a testing set.
-The purpose of training set is twofold: (1) We apply a data miner to construct an oracle that acts as a quality predictor; (2) We build the planners to reason across the data.
-
-Next, we employ the planners to generate plans to improve the quality of defective classes in the test set.
-In addition to the plans obtained from our planners we may have some common refactoring options that the developers tend to use.
-The oracle built previously can be used to assess the the changed examples and this can be used to compare refactoring operations.
-
-It's worth noting that this approach depends on having effective oracles for assessing the results.
-It proved to be more complicated to commission the Jureczko data sets for this study.
-For that data, we found that the quality predictors built from this data are far from perfect; However, for some data sets, the  predictors could be salvaged using the techniques discussed in this section.
-
-\fig{j} shows our preliminary studies with the Jureczko data.
+
+  An important part of this process is having an oracle
+  that can assess the defect potential of the {\em initial}
+  and {\em new code}. This section describes how we built
+  that oracle, using Random Forests~\cite{Breiman2001}
+  from the SciKit learn kit~\cite{Pedregosa2012}.
+  Random Forests are a decision tree learning method but
+  instead of building one tree, hundreds are built using
+  randomly selected subsets of the data. The final predictions
+  come from averaging the predictions over all the trees.
+  Recent students rank Random Forests as  the top-most
+  effective defect prediction learning algorithm~\cite{lessmann}.
+
+\fig{j} shows our preliminary studies with Random Forests and
+the Jureczko data.
 Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}, this means that we are training on hundreds to thousands of classes and testing on smaller test suites).
@@
 
-In order to identify the presence (or absence) of defects, we can consider using Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
-For such data, quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
+In order to identify the presence (or absence) of defects, we can   use Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
+For such data, the quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
 ``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
@@
 As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forest (again, from the SciKit learn kit~\cite{Pedregosa2012}) was built from the training data, then applied to the test data.
+Random Forests was built from the training data, then applied to the test data.
 The ``untuned'' columns of \fig{j} shows those results.
@@
 
-\subsection{Need for Planning}
-During lifecycle, a software system undergoes continuous changes in order to keep up with the business level demands \cite{lehman79}.
-Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{fowler09}.
-Code smells are symptomatic of design issues in codes like code understandability and code changeability.
-If left unaddressed, these can lead to a variety of maintenance problems and faults in software systems \cite{fowler09}\cite{marinescu06}.
-Fowler \cite{fowler09} and Brown et al.
-\cite{brown98} defined a catalogue of more than 30 anti-patterns.
-It is a common notion that the critiques motivated by anti-patterns and code smells would be more comprehensible to developers because many code smells are associated with conventional refactoring operations.
-
-The choice of refactoring operations to address the code smells arise from strongly held beliefs of developers on these topics, and their beliefs in turn are motivated by personal experiences.
-
-The distinction between developers' beliefs and empirical findings is particularly profound in the case of code smells.
-For instance, \cite{abbes11} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{abbes11}\cite{yama13}.
-While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern.
-In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{arcoverde11}\cite{chatzigeorgiou10}.
-Additionally, Yamashita and Moonen\cite{} observed that several developers did not take conscious actions to alleviate bad smells that were present in the code.
-
-Closer inspection revealed that this is not due to the lack of relevance of code smells, but a number of other factors, like: (1) lack of awareness; (2) More importantly, lack of developer friendly tools \cite{}.
-The study goes on to show that these developers are in need of better support tools that assist them during the software evolution cycle.
-Specifically, developers needed tools to (1) be developer friendly; (2) be operable in real-time; and (3) be capable of identifying problematic areas using error prediction.
-
-Therefore to assist developers in their refactoring efforts by providing them with reliable analytics it is useful to have automatic methods that can augment human expertise (or replace them if such experts are absent) in addition to offering a second opinion.
-Accordingly, the rest of this paper defines and evaluates automatic methods to address the above issues.
-
-\subsection{What is a ``Plan''?}
-To address the issues discussed in the previous section, we conjecture that the planners must generate valid and meaningful refactoring operations, we refer to these as \textit{plans}.
-These plans are synthesized by the planner and are thus unique in that they are agnostic to any prior notions of the developers in addition to them being local to the software project being refactored.
-
-To do this, our planners use tables of data with independent features and a dependent class feature.
-Classes have weights  that  indicate  what rows are ``bad'' or ``better''.
-Plans  change   a row such that it is more likely to be ``better''.
-Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to adjust   feature $Z_j$:
-
-{\small\[
-\forall \delta_j \in \Delta :  Z_j =  
-\begin{cases}
-     Z_j + \delta_j& \text{if $Z_j$ is numeric}\\
-    \delta_j              & \text{otherwise}
-\end{cases}
-\]}
-
-For example, to simplify a  large bug-prone  method, our planners might suggest to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across two simpler functions).
-
-Additionally, if a developer were to suggest a specific refactoring option, our planners can query the option and report on its possible usability.
-Note that we make no assumption that a plan mentions every feature (so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than plan2).
-
-\section{Test Data}\label{sect:tesd}
-
-To assess our planning methods, our data (see git.io/vGYxc) comes from Jureczko et al.'s object-oriented JAVA systems~\cite{jureczko10}.
-
-The Jureczko data records the number of known defects for each class using a post-release bug tracking system.
-The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc.
-For details on the Jureczko data, see  \fig{ck}.
-All the planning methods of this paper reflect on the numeric value of the raw defect counts to estimate the quality of the class.
-The predictor considers defects to be a Boolean class data set, where defects are \texttt{TRUE} if the numeric defect count is greater than zero and \texttt{FALSE} otherwise.
-The choice of this data set was is based on the taxonomy of the smells, in which we focus on the measurable properties needed to detect a given smell.
-This set of metrics is not restricted and can be easily extended with other metrics.
-
-\input{tex/algos}
-
-\section{Planning Methods}\label{sect:planners}
-
-In this we describe the planners used in our study.
-First, we reintroduce a planner called CD which was an outcome of our previous work~\cite{me12c} which generates plans using a top-down bi-clustering method described in \fig{where}.C to recursively partition the data along the dimension that captures the greatest variability in the data.
-Then, we present XTREE, a novel planner introduced in this work, which uses a decision tree learner \fig{where}.D for planning.
-
-% This section describes the new XTREE planner and a planner called CD which was an outcome of our previous work~\cite{me12c}.
-XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
-
-Both methods aim to address the issues highlighted in \tion{prelim}.
-To achieve this: (1)~they find plans, which can be use to generate reliable refactoring options, particular to each test case by reflecting on local data; and (2)~they offer a mechanism for critiquing bad refactoring options.
-
-% (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
-
-\subsection{  Methods}
-
-Our  description of the methods adopts the following convention.
-All variables set via  our engineering judgement  with Greek letters; e.g.
-$\alpha,\beta,\gamma,\omega$.
-In this paper, we show our current settings to these variables produces useful results.
-Elsewhere \cite{krall14,fu:ase15}, we are exploring tuning methods to find better settings but  we have nothing definitive yet to report on auto-tuning planners.
-
-\subsubsection{Centroid Deltas (CD)}
-
-% Method1 computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be (which we will call $C_j$).
-
-Large data sets can be adequately represented by a few dozen (or so) clusters, and meaningful inferences can be drawn from the centroids of these clusters.
-This method generates plans for improving faulty classes by looking at the difference between the cluster centroid that most closely resembles the defective class from the training data (denoted by $C_i$) and the cluster centroid that represents a better, non-defective, class (denoted by $C_j$).
-
-To do this, CD clusters training data based on the independent variables using the WHERE algorithm of \fig{where}.C and replaces all clusters with their centroids computed from the median value of each continuous/discrete feature.
-Then, for every cluster $C_i$ it finds the closest centroid $C_j$ that has a better performance score.
-For Jureczko data, ``better'' means fewer defective examples.
-CD then caches the delta between the independent features of $C_i$ and $C_j$.
-For continuous features, this delta is $C_j - C_i$.
-Finally, for every test case, CD uses the distance measure $d$ shown in \fig{where}.B to find the nearest centroid $C_i$.
-It then proposes a plan for improving that test case: the conjunction of all the deltas between $C_i$ and $C_j$.
-
-\input{tex/dtree.tex}
-
-\subsubsection{XTREE}
-
-A potential problem with CD is the {\em unsupervised} nature of the clustering algorithm (WHERE) that executes without knowledge of the target class.
-It would be beneficial for us to use {\em Supervised} methods that also reflect on the target class.
-
-XTREE was developed to address this problem.
-It uses a supervised decision tree algorithm of \fig{where}.D to categorize the data Next, XTREE devises plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
-That code asks three questions, the last of which returns the plan:
-
-\be
-\item What {\em current} branch does a test case fall in?
-\item What {\em desired} branch would the test case want to move to?
-\item What are the {\em deltas} between current and desired? 
-\ee
-
-In finding the answers to the above questions, XTREE can (a) Generate an wealth of plans (that can be potential refactoring operations); (b) Critique plans based on their location in the decision tree constructed.
-In addition to the above, XTREE also offers developers a visual medium to validate their proposals for refactoring, see ~\fig{xtree_samp}.
-
-\begin{figure}[btp!]
-\centering
-\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
-\caption{Sample plans from XTREE}\label{fig:xtree_samp}
-\end{figure}
-
-\section{Experiments}
-
-This section describes an experimental design (and results) for evaluating the above four methods. 
-\subsection{Experimental Design}
-
-\subsubsection{A Strategy for Evaluating Planners}
-
-Our experimental design is shown in \fig{design}. We divide the
-project data  into two disjoint sets {\em train} and {\em test}
-(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
-Next, from the train set, we build both a {\em planner} and
- a {\em  predictor}.
-
-Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
-\bi
-\item Our {\em planners} will be XTREE and CD;
-\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.
-We use these
-data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
-\ei
-As for the {\em test} data, this is passed to the { predictor}
-to measure performance statistics related to effectiveness.
-
-If our { predictors} fail to perform effectively on the test data,
-then we cannot trust them to comment on our plans. Accordingly,
-if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
-we should not use some of the  Jureczko data.
-
-Else, we (1)~apply the { planner} to alter the {\em test} data;
-then (2)~apply the { predictor} to the altered data $test'$;
-then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
-\bi
-\item If $R  = 0\%$, this means  ``no change from baseline''; 
-\item If $R \gt 0\%$, this indicates ``
-
-improvement over the baseline'';
-\item If $R \lt 0\%$, this indicates ``optimization failure''.
-\ei
-
-\begin{figure}[!t]
-{\small 
-\[
-\begin{array}{r} 
-\mathrm{project}\\
-\mathrm{data}
-\end{array} 
-\left\{\begin{array}{l}\mathit{train}
-        \left\{\begin{array}{l}
-                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
-                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
-              \end{array}\right.
-       \\
-      ~\\
-\mathit{test}  
-    \left\{\begin{array}{l@{~}l}
-           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
-           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
-           \mathrm{\bf then} &
-           \left\{
-            \begin{array}{l}
-                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
-                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
-                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
-            \end{array}
-          \right.
-   \end{array}\right.
-\end{array} \right. 
-\]}
- \caption{Experimental design .}\label{fig:design}
- \end{figure}
-
-\subsubsection{Statistical Methods}
-Our methods use some stochastic algorithms; e.g.
-WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan (see \fig{xtrees_bare}).
-Hence, we report the $R$ values seen in 40 repeated runs, with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
-
-To rank our methods using the results from these 40 repeats, we use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}.
-
-In accordance to that test, using the median values of each method, we
-sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
-Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.
-E.g.
-for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$: \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
-
-We then apply a apply a statistical hypothesis test $H$ to check
-if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping).
-If so, Scott-Knott recurses on the splits.
-In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
-
-For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
-For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
-These researchers warn that even if a hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
-Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
-
-\subsubsection{Report Format}
-
-Our results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
-The black dot shows the median $R$ value and the horizontal likes stretch 
-from the 25th percentile to the 75th percentile (a region called the inter-quartile
-range, or IQR).
-
-\begin{center}
-
-{\small Example \begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
-\arrayrulecolor{lightgray}
-\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
-1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
-\hline 2 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
-\hline \end{tabular}}
-\end{center}
-
-In this example table, the rows are  sorted on the median values of each method.
-Note that all the methods have $R\gt0\%$; i.e.
-all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 62\% from the original value).
-
-The above example table has a  left-hand-side  {\bf Rank} column, computed using the
-Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and CD+FS together are ranked the worst.
-
-\subsubsection{Other Details}
-
-\fig{jur} shows the effectiveness of our methods seen in 40 repeats with each data set.
-In these experiments,   the dependent variables of Jureczko data set is discrete. Hence, while choosing the predictor, we used Random Forest as a classifier for Jureczko data.
-
-The Jureczko data, being temporal in nature, allows us to implement a validation procedure that ensures only past data is ever used to predict future values.
-Hence, in that data, we used the train/test sets shown in \fig{j}.
-(Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).
-
-\input{tex/results_jur.tex}
-
-\subsection{Experimental Results}
-
-\begin{figure}[tb!]
-\centering
-\includegraphics[width=\linewidth]{figs/thresh.png}
-\caption{Thresholds used by DECOR to detect Spaghetti Code~\cite{moha10}.}\label{fig:thresh}
-\end{figure}
-
-Recall from our introduction that we are assessing planners on three criteria:
-{\em effectiveness}, which is how much they reduce the expected value of the changed examples;
-{\em succinctness}, which is how many things we need to change to achieve a plan;
-and {\em insightfulness}, which is how different are the plans from standard truisms.
-
-\begin{figure}[tbp!]
-\centering
-\includegraphics[width=\linewidth]{figs/PercentChanges.png}
-\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
-\end{figure}
-
-\begin{figure*}[!y]
-\centering
-\includegraphics[width=0.90\linewidth]{figs/changes0.png}
-\caption{Suggested Changes}
-\label{figs:changes}
-\end{figure*}
-
-\subsubsection{RQ1: Can a planner suggest plans to improve the quality ofa defective class?}
-
-Measured in terms of effectiveness, some data sets were harder to optimize than others.
-However, in most data sets, large reductions were observed.
-We noticed an improvement of 85.54\%, as compared to the original baseline in Ant and 36.36\% in Lucene, \fig{jur}.
-
-Overall, XTREE was most the effective.
-It was the top-ranked method in 3 cases and performed just as well in the other two.
-Where it ranked better, it had significant improvements in the median performance values.
-In the Jureczko data sets, there was an improvement of at least 25\% compared to our previous method.
-
-\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
-The left-hand-side plot of that figure reports results from one of the Jureczko data sets ({\em lucene}) and the right-hand-side shows a Siegmund data set ({\em BDBJ}).
-
-In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e.
-the vertical bars in that plot are {\em lower}).
-For example, XTREE's plans were usually succinct --- in all  data sets, XTREE   changes around a fifth of the features (see \fig{types}).
-On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features (observe the change frequencies as high as 100\% for all features).
-Method1's policy of ``change everything''  might be acceptable if this approach lead to the most effective changes.
-However, in \fig{jur} there is no evidence for this.
-
-\input{tex/deltas_avg.tex}
-
-\subsubsection{RQ2: Can a planner be used to critique bad refactoring options?}\label{sect:surprise}
-
-If a planner only ever reported conclusions that were already known, then that planner offers
-little value over ``just use established wisdom''. Accordingly, we studied our results
-for plans that were somewhat counter-intuitive.
-
-Such a surprising plan can be  seen in {\em lucene}. Recall the standard advice for
-OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
-to specific projects by checking how often a planner changes the
- coupling-related features:
-\bi
-\item {\em ca}:   afferent couplings  =  \# classes using this
-				class;
-\item {\em  	ce}:  efferent couplings =  \# classes  used by this
-				  class. 
-\item {\em cbm}: coupling between methods =  \# new/redefined methods
-				to which all the inherited methods are coupled
-\item	{\em cbo}:  coupling between objects = a value that increases when the methods of one
-				class access services of another.
-
-\item {\em ic}:   inheritance coupling =  \# parent classes  which a given
-				class is coupled (including methods and variables inherited)
-\ei
-In many of our results with the Jureczko data, it was indeed true that the changes
-proposed by XTREE lead to lower coupling. However, the {\em lucene} results were quite
-different and rather surprising.
-
-In the {\em lucene} XTREE results  of  \fig{changed},
-the most frequent change was to alter the lines of code in a class (see the tallest
-red histogram in that figure on the {\em loc}, or lines of code). 
-Looking at the logs of our planner, we can see
-that XTREE's proposed change is to {\em reduce} the size of a class.
-The only way to do that,  while keeping the  same functionality, is to create a network
-of smaller classes that interact to produce that functionality.
-That is, we would need to {\em increase} the coupling of those classes to achieve XTREE's plan.
-
-In theory, increasing coupling between classes complicates and confuses a class design.
-But the {\em lucene} XTREE results  of  \fig{changed} rarely proposes changing   the
-coupling features {\em ca, ce, cbm, ic} (in fact, XTREE never proposes any change to {\em ic}).
-
-The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
-in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
-XTREE is warning against, say, some factory class generating a large
-community of agents, all of the same class, who
-co-ordinate on some task. This is a different issue to the class redesign issue that would
-be triggered by altering {\em loc}.
-
-In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
-At least for the {\em lucene} data set, we can see advice that recommends {\em increasing coupling}
-to reduce defects.
+\subsection{Resylts}
 
@@
 compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.
-
-\section{Related work}
-
-\subsection{Planning in AI}
-
-The XTREE planner is somewhat different to the logic-based planners explored by 
-classical AI. 
-Those kinds of planners employ a logical procedure~\cite{Fikes1971}
-that seeks an ordering on {\em operators} to take some domain
-{\em state} from a {\em start} state to a {\em  goal} state.
-This classical logical approach is known to suffer from
-computational bottlenecks~\cite{Bylander1994}. On the other hand, tools like XTREE will scale to any domain
-that can generate decision trees.
-
-\subsection{Evaluating Changes}
-
-Some organizations have the resources to 
-run repeated trials to assess  project changes.
-For example, in one   study, Bente et al. reported results
-where the same  specification was developed by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
-to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
-then check in 12 months time
- which teams have fewer defects than the others.
-
-Seldom do industrial or research groups have access
-to the kinds of resources needed for this kind study  (evidence: in the six years since the
-publication of that work, we know of only one   similar study to Bente et al.). Also, given the
-diversity of modern software projects, it might be unreasonable to demand that all
-proposed changes for all projects are always evaluated by something like the Bente et al. study.
-Hence, this paper has used data miners to build an oracle that can assess changed examples. The advantage
-of our approach is that it required far less resources to assess the effectiveness of proposed
-changes to a project.
-
-\subsection{Search-based SE}
-
-Another way way to propose changes to software artifacts
-is   via some search-based method~\cite{Harman2009,Harman2011}. Such SBSE methods are   evolutionary programs that 
-make
- extensive changes to  some initial sample of project data
- (perhaps 
-100s to 100,000s of mutations). Each of these mutations
-is reassessed using some domain model.
-Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
-deb14,Cui2005a,zhang07:TEC}.
-
-One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
-of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
-to prevent new examples from mutating away from the kinds of data used to commission the model (in which case, we would
-start doubting the model's output).
-
-Another issue with standard search-based SE methods is that they require ready access to 
-trustworthy domain model that can offer an assessment
-of newly generated examples. While some domains have such models (e.g. see the COCOMO effort estimation model
-used in the last section), our experience is that many others do not.  For example, 
-consider software defect prediction and all the intricate issues that may lead to defects in a product. A model that includes {\em all} those
-potential issues would be very large and complex. Further,
-the empirical data required to validate any/all parts
-of that model can be hard to find.
-
-What we would recommend is a two-pronged policy.
-In domains with ready access to trusted models, we recommend
-the kinds of tools that are widely used in the search-based
-software engineering community such as GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
-deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like XTREE.
 
@@
 
-It turns out that ranking codiiiiie smells, and ignoring the low priority
+It turns out that ranking code smells, and ignoring the low priority
 ones, is a pressing problem. As shown in our {\em Related Work} section,
@@
 certain results raise doubts about the value of   code smells for reducing the risk of defects in software.
+
+\subsection{Need for Planning}
+During lifecycle, a software system undergoes continuous changes in order to keep up with the business level demands \cite{lehman79}.
+Unfortunately, due to market/customers constraints, the source code quality is often neglected with the risk of introducing \textit{code bad smell} (or code-smells) \cite{fowler09}.
+Code smells are symptomatic of design issues in codes like code understandability and code changeability.
+If left unaddressed, these can lead to a variety of maintenance problems and faults in software systems \cite{fowler09}\cite{marinescu06}.
+Fowler \cite{fowler09} and Brown et al.
+\cite{brown98} defined a catalogue of more than 30 anti-patterns.
+It is a common notion that the critiques motivated by anti-patterns and code smells would be more comprehensible to developers because many code smells are associated with conventional refactoring operations.
+
+The choice of refactoring operations to address the code smells arise from strongly held beliefs of developers on these topics, and their beliefs in turn are motivated by personal experiences.
+
+The distinction between developers' beliefs and empirical findings is particularly profound in the case of code smells.
+For instance, \cite{abbes11} showed that the presence of an anti-pattern in the source code does not decrease the developers' performance, while a combination of anti-patterns results in a significant decrease of their performance \cite{abbes11}\cite{yama13}.
+While the results of this study indicates that single anti-patterns are not harmful, they also reveals that anti-patterns are quite diffuse in software systems and very often code components are affected by more than one anti-pattern.
+In addition, we have empirical evidence that the number of anti-patterns in software systems increases over time and only in few cases they are removed through refactoring operations \cite{arcoverde11}\cite{chatzigeorgiou10}.
+Additionally, Yamashita and Moonen\cite{} observed that several developers did not take conscious actions to alleviate bad smells that were present in the code.
+
+Closer inspection revealed that this is not due to the lack of relevance of code smells, but a number of other factors, like: (1) lack of awareness; (2) More importantly, lack of developer friendly tools \cite{}.
+The study goes on to show that these developers are in need of better support tools that assist them during the software evolution cycle.
+Specifically, developers needed tools to (1) be developer friendly; (2) be operable in real-time; and (3) be capable of identifying problematic areas using error prediction.
+
+Therefore to assist developers in their refactoring efforts by providing them with reliable analytics it is useful to have automatic methods that can augment human expertise (or replace them if such experts are absent) in addition to offering a second opinion.
+Accordingly, the rest of this paper defines and evaluates automatic methods to address the above issues.
+
+\subsection{What is a ``Plan''?}
+To address the issues discussed in the previous section, we conjecture that the planners must generate valid and meaningful refactoring operations, we refer to these as \textit{plans}.
+These plans are synthesized by the planner and are thus unique in that they are agnostic to any prior notions of the developers in addition to them being local to the software project being refactored.
+
+To do this, our planners use tables of data with independent features and a dependent class feature.
+Classes have weights  that  indicate  what rows are ``bad'' or ``better''.
+Plans  change   a row such that it is more likely to be ``better''.
+Specifically, for every test example $Z$,   planners proposes a  plan $\Delta$ to adjust   feature $Z_j$:
+
+{\small\[
+\forall \delta_j \in \Delta :  Z_j =  
+\begin{cases}
+     Z_j + \delta_j& \text{if $Z_j$ is numeric}\\
+    \delta_j              & \text{otherwise}
+\end{cases}
+\]}
+
+For example, to simplify a  large bug-prone  method, our planners might suggest to a developer to reduce its size (i.e.  refactor that code by, say, splitting it across two simpler functions).
+
+Additionally, if a developer were to suggest a specific refactoring option, our planners can query the option and report on its possible usability.
+Note that we make no assumption that a plan mentions every feature (so plan1 can be  more succinct than plan2 when plan1  mentions fewer features than plan2).
+
+\input{tex/algos}
+
+\section{Planning Methods}\label{sect:planners}
+
+In this we describe the planners used in our study.
+First, we reintroduce a planner called CD which was an outcome of our previous work~\cite{me12c} which generates plans using a top-down bi-clustering method described in \fig{where}.C to recursively partition the data along the dimension that captures the greatest variability in the data.
+Then, we present XTREE, a novel planner introduced in this work, which uses a decision tree learner \fig{where}.D for planning.
+
+% This section describes the new XTREE planner and a planner called CD which was an outcome of our previous work~\cite{me12c}.
+XTREE, a novel planner introduced in this work,  uses the decision tree learner of \fig{where}.D.
+
+Both methods aim to address the issues highlighted in \tion{prelim}.
+To achieve this: (1)~they find plans, which can be use to generate reliable refactoring options, particular to each test case by reflecting on local data; and (2)~they offer a mechanism for critiquing bad refactoring options.
+
+% (i) generating reliable refactoring options, and (ii) critiquing bad refactoring options.
+
+\subsection{  Methods}
+
+Our  description of the methods adopts the following convention.
+All variables set via  our engineering judgement  with Greek letters; e.g.
+$\alpha,\beta,\gamma,\omega$.
+In this paper, we show our current settings to these variables produces useful results.
+Elsewhere \cite{krall14,fu:ase15}, we are exploring tuning methods to find better settings but  we have nothing definitive yet to report on auto-tuning planners.
+
+\subsubsection{Centroid Deltas (CD)}
+
+% Method1 computes a plan from the difference between where you are  (which we will call $C_i$) and where you want to be (which we will call $C_j$).
+
+Large data sets can be adequately represented by a few dozen (or so) clusters, and meaningful inferences can be drawn from the centroids of these clusters.
+This method generates plans for improving faulty classes by looking at the difference between the cluster centroid that most closely resembles the defective class from the training data (denoted by $C_i$) and the cluster centroid that represents a better, non-defective, class (denoted by $C_j$).
+
+To do this, CD clusters training data based on the independent variables using the WHERE algorithm of \fig{where}.C and replaces all clusters with their centroids computed from the median value of each continuous/discrete feature.
+Then, for every cluster $C_i$ it finds the closest centroid $C_j$ that has a better performance score.
+For Jureczko data, ``better'' means fewer defective examples.
+CD then caches the delta between the independent features of $C_i$ and $C_j$.
+For continuous features, this delta is $C_j - C_i$.
+Finally, for every test case, CD uses the distance measure $d$ shown in \fig{where}.B to find the nearest centroid $C_i$.
+It then proposes a plan for improving that test case: the conjunction of all the deltas between $C_i$ and $C_j$.
+
+\input{tex/dtree.tex}
+
+\subsubsection{XTREE}
+
+A potential problem with CD is the {\em unsupervised} nature of the clustering algorithm (WHERE) that executes without knowledge of the target class.
+It would be beneficial for us to use {\em Supervised} methods that also reflect on the target class.
+
+XTREE was developed to address this problem.
+It uses a supervised decision tree algorithm of \fig{where}.D to categorize the data Next, XTREE devises plans from the branches of the decision trees using the code of \fig{xtrees_bare}.
+That code asks three questions, the last of which returns the plan:
+
+\be
+\item What {\em current} branch does a test case fall in?
+\item What {\em desired} branch would the test case want to move to?
+\item What are the {\em deltas} between current and desired? 
+\ee
+
+In finding the answers to the above questions, XTREE can (a) Generate an wealth of plans (that can be potential refactoring operations); (b) Critique plans based on their location in the decision tree constructed.
+In addition to the above, XTREE also offers developers a visual medium to validate their proposals for refactoring, see ~\fig{xtree_samp}.
+
+\begin{figure}[btp!]
+\centering
+\includegraphics[width=\linewidth]{figs/XTREE_samp.png}
+\caption{Sample plans from XTREE}\label{fig:xtree_samp}
+\end{figure}
+
+\section{Related work}
+
+\subsection{Planning in AI}
+
+The XTREE planner is somewhat different to the logic-based planners explored by 
+classical AI. 
+Those kinds of planners employ a logical procedure~\cite{Fikes1971}
+that seeks an ordering on {\em operators} to take some domain
+{\em state} from a {\em start} state to a {\em  goal} state.
+This classical logical approach is known to suffer from
+computational bottlenecks~\cite{Bylander1994}. On the other hand, tools like XTREE will scale to any domain
+that can generate decision trees.
+
+\subsection{Evaluating Changes}
+
+Some organizations have the resources to 
+run repeated trials to assess  project changes.
+For example, in one   study, Bente et al. reported results
+where the same  specification was developed by four different organizations~\cite{Anda2009}. Given those kind of resources, it would be possible
+to (say) take a code base, assign it to different teams, make these teams  adopt different polices,
+then check in 12 months time
+ which teams have fewer defects than the others.
+
+Seldom do industrial or research groups have access
+to the kinds of resources needed for this kind study  (evidence: in the six years since the
+publication of that work, we know of only one   similar study to Bente et al.). Also, given the
+diversity of modern software projects, it might be unreasonable to demand that all
+proposed changes for all projects are always evaluated by something like the Bente et al. study.
+Hence, this paper has used data miners to build an oracle that can assess changed examples. The advantage
+of our approach is that it required far less resources to assess the effectiveness of proposed
+changes to a project.
+
+\subsection{Search-based SE}
+
+Another way way to propose changes to software artifacts
+is   via some search-based method~\cite{Harman2009,Harman2011}. Such SBSE methods are   evolutionary programs that 
+make
+ extensive changes to  some initial sample of project data
+ (perhaps 
+100s to 100,000s of mutations). Each of these mutations
+is reassessed using some domain model.
+Examples of these algorithms include GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
+deb14,Cui2005a,zhang07:TEC}.
+
+One problem with these   SBSE methods is that they can  make extensive mutations to the data they are exploring. In the language
+of \tion{trust}, these methods may not be {\em trust-increasing} since those algorithms make no attempt
+to prevent new examples from mutating away from the kinds of data used to commission the model (in which case, we would
+start doubting the model's output).
+
+Another issue with standard search-based SE methods is that they require ready access to 
+trustworthy domain model that can offer an assessment
+of newly generated examples. While some domains have such models (e.g. see the COCOMO effort estimation model
+used in the last section), our experience is that many others do not.  For example, 
+consider software defect prediction and all the intricate issues that may lead to defects in a product. A model that includes {\em all} those
+potential issues would be very large and complex. Further,
+the empirical data required to validate any/all parts
+of that model can be hard to find.
+
+What we would recommend is a two-pronged policy.
+In domains with ready access to trusted models, we recommend
+the kinds of tools that are widely used in the search-based
+software engineering community such as GALE, NSGA-II, NSGA-III, SPEA2, IBEA, particle swarm optimization, MOEA/D, etc.~\cite{krall14,deb00a,zit02,zit04,%
+deb14,Cui2005a,zhang07:TEC}. Otherwise, we recommend tools like XTREE.
+
+This section describes an experimental design (and results) for evaluating the above four methods. 
+\subsection{Experimental Design}
+
+\subsubsection{A Strategy for Evaluating Planners}
+
+Our experimental design is shown in \fig{design}. We divide the
+project data  into two disjoint sets {\em train} and {\em test}
+(so \mbox{{\em train} $\cap ${\em test} $=\;\emptyset$}).
+Next, from the train set, we build both a {\em planner} and
+ a {\em  predictor}.
+
+Our general framework does not   commit to any particular choice of { planner} or { predictor} but, for the purposes of this paper:
+\bi
+\item Our {\em planners} will be XTREE and CD;
+\item Our  {\em predictor} will be the Random Forest Classifier~\cite{Breiman2001} (for discrete classes) and Random Forest Regressor (for continuous classes) taken from  SciKit Learn~\cite{Pedregosa2012}.
+We use these
+data miners since extensive studies have shown these to be amongst the better alternatives for mining software data~\cite{lessmann}.
+\ei
+As for the {\em test} data, this is passed to the { predictor}
+to measure performance statistics related to effectiveness.
+
+If our { predictors} fail to perform effectively on the test data,
+then we cannot trust them to comment on our plans. Accordingly,
+if that performance is unsatisfactory, we abort. Recall from \tion{tesd} that this step indicated
+we should not use some of the  Jureczko data.
+
+Else, we (1)~apply the { planner} to alter the {\em test} data;
+then (2)~apply the { predictor} to the altered data $test'$;
+then (3)~return data on the {\em before, after} predictions expressed as percent improvement, denoted by  $R=(1-\frac{\mathit{after}}{\mathit{before}})\times100\%$, with the following following properties:
+\bi
+\item If $R  = 0\%$, this means  ``no change from baseline''; 
+\item If $R \gt 0\%$, this indicates ``
+
+improvement over the baseline'';
+\item If $R \lt 0\%$, this indicates ``optimization failure''.
+\ei
+
+\begin{figure}[!t]
+{\small 
+\[
+\begin{array}{r} 
+\mathrm{project}\\
+\mathrm{data}
+\end{array} 
+\left\{\begin{array}{l}\mathit{train}
+        \left\{\begin{array}{l}
+                \mathrm{learn\;a\;}\mathrm{predictor\;}\mathrm{(e.g.\;via \;Random\;Forest)}\\
+                \mathrm{learn\;a\;}\mathrm{planner\;}\mathrm{(e.g.\;via \; XTREE)}
+              \end{array}\right.
+       \\
+      ~\\
+\mathit{test}  
+    \left\{\begin{array}{l@{~}l}
+           \mathit{before}& =\mathrm{Performance\; scores \hspace{2pt} in \hspace{2pt}}\mathit{test}\\
+           \mathrm{\bf if\;}\mathit{before} & >  \mathit{0}\\
+           \mathrm{\bf then} &
+           \left\{
+            \begin{array}{l}
+                \mathit{test'} = \mathrm{planner}(\mathit{test})\\
+                \mathit{after} =\mathrm{predictor}(\mathit{test'})\\ 
+                \mathrm{{\bf return}\;} R=(1-\frac{\mathit{after}}{\mathit{before}})\times 100 \%
+            \end{array}
+          \right.
+   \end{array}\right.
+\end{array} \right. 
+\]}
+ \caption{Experimental design .}\label{fig:design}
+ \end{figure}
+
+\subsubsection{Statistical Methods}
+Our methods use some stochastic algorithms; e.g.
+WHERE's selection of ``what example to explore first'' (see \fig{where}.C) and XTREE' occasional use of a random guess when deciding what part of a discretized range to include in the plan (see \fig{xtrees_bare}).
+Hence, we report the $R$ values seen in 40 repeated runs, with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
+
+To rank our methods using the results from these 40 repeats, we use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}.
+
+In accordance to that test, using the median values of each method, we
+sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
+Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.
+E.g.
+for lists $l,m,n$ of size $ls,ms,ns$ where $l=m\cup n$: \[E(\Delta)=\frac{ms}{ls}abs(m.\mu - l.\mu)^2 + \frac{ns}{ls}abs(n.\mu - l.\mu)^2\]
+
+We then apply a apply a statistical hypothesis test $H$ to check
+if $m,n$ are significantly different  (in our case, the conjunction of A12 and bootstrapping).
+If so, Scott-Knott recurses on the splits.
+In other words, we divide the data if \textit{both} bootstrap sampling and effect size test agree that a division is statistically significant (with a confidence of 99\%) and not a small effect ($A12 \ge 0.6$).
+
+For a justification of the use of non-parametric bootstrapping, see Efron \& Tibshirani~\cite[p220-223]{efron93}.
+For a justification of the use of effect size tests see Shepperd\&MacDonell~\cite{shepperd12a}; Kampenes~\cite{kampenes07}; and Kocaguenli et al.~\cite{Kocaguneli2013:ep}.
+These researchers warn that even if a hypothesis test declares two populations to be ``significantly'' different, then that result is misleading if the ``effect size'' is very small.
+Hence, to assess the performance differences we first must rule out small effects using A12, a test   recently endorsed by Arcuri and Briand at ICSE'11~\cite{arcuri11}.
+
+\subsubsection{Report Format}
+
+Our results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
+The black dot shows the median $R$ value and the horizontal likes stretch 
+from the 25th percentile to the 75th percentile (a region called the inter-quartile
+range, or IQR).
+
+\begin{center}
+
+{\small Example \begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
+\arrayrulecolor{lightgray}
+\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
+1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
+\hline 2 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
+\hline \end{tabular}}
+\end{center}
+
+In this example table, the rows are  sorted on the median values of each method.
+Note that all the methods have $R\gt0\%$; i.e.
+all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 62\% from the original value).
+
+The above example table has a  left-hand-side  {\bf Rank} column, computed using the
+Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and CD+FS together are ranked the worst.
+
+\subsubsection{Other Details}
+
+\fig{jur} shows the effectiveness of our methods seen in 40 repeats with each data set.
+In these experiments,   the dependent variables of Jureczko data set is discrete. Hence, while choosing the predictor, we used Random Forest as a classifier for Jureczko data.
+
+The Jureczko data, being temporal in nature, allows us to implement a validation procedure that ensures only past data is ever used to predict future values.
+Hence, in that data, we used the train/test sets shown in \fig{j}.
+(Aside: note  that all the SMOTE-ing and Random Forest tunings (discussed in \tion{tesd}) occurred in the {\em train} phase of \fig{design}).
+
+\input{tex/results_jur.tex}
+
+\subsection{Experimental Results}
+
+\begin{figure}[tb!]
+\centering
+\includegraphics[width=\linewidth]{figs/thresh.png}
+\caption{Thresholds used by DECOR to detect Spaghetti Code~\cite{moha10}.}\label{fig:thresh}
+\end{figure}
+
+Recall from our introduction that we are assessing planners on three criteria:
+{\em effectiveness}, which is how much they reduce the expected value of the changed examples;
+{\em succinctness}, which is how many things we need to change to achieve a plan;
+and {\em insightfulness}, which is how different are the plans from standard truisms.
+
+\begin{figure}[tbp!]
+\centering
+\includegraphics[width=\linewidth]{figs/PercentChanges.png}
+\caption{Percent frequency for how often certain feature was changed by a plan.}\label{fig:changed}
+\end{figure}
+
+\begin{figure*}[!y]
+\centering
+\includegraphics[width=0.90\linewidth]{figs/changes0.png}
+\caption{Suggested Changes}
+\label{figs:changes}
+\end{figure*}
+
+\subsubsection{RQ1: Can a planner suggest plans to improve the quality ofa defective class?}
+
+Measured in terms of effectiveness, some data sets were harder to optimize than others.
+However, in most data sets, large reductions were observed.
+We noticed an improvement of 85.54\%, as compared to the original baseline in Ant and 36.36\% in Lucene, \fig{jur}.
+
+Overall, XTREE was most the effective.
+It was the top-ranked method in 3 cases and performed just as well in the other two.
+Where it ranked better, it had significant improvements in the median performance values.
+In the Jureczko data sets, there was an improvement of at least 25\% compared to our previous method.
+
+\fig{changed} reports the percent of times in the 40 repeats that a method proposed changing a feature.
+The left-hand-side plot of that figure reports results from one of the Jureczko data sets ({\em lucene}) and the right-hand-side shows a Siegmund data set ({\em BDBJ}).
+
+In these plots, {\em more} succinct a planning method, {\em fewer} the frequency (in percent) where it recommends changing a particular feature (i.e.
+the vertical bars in that plot are {\em lower}).
+For example, XTREE's plans were usually succinct --- in all  data sets, XTREE   changes around a fifth of the features (see \fig{types}).
+On other hand,  Method1 (CD) was the least succinct  since it  wanted to change all features (observe the change frequencies as high as 100\% for all features).
+Method1's policy of ``change everything''  might be acceptable if this approach lead to the most effective changes.
+However, in \fig{jur} there is no evidence for this.
+
+\input{tex/deltas_avg.tex}
+
+\subsubsection{RQ2: Can a planner be used to critique bad refactoring options?}\label{sect:surprise}
+
+If a planner only ever reported conclusions that were already known, then that planner offers
+little value over ``just use established wisdom''. Accordingly, we studied our results
+for plans that were somewhat counter-intuitive.
+
+Such a surprising plan can be  seen in {\em lucene}. Recall the standard advice for
+OO systems: build classes that are internally cohesive with low coupling to other parts of the system~\cite{Dhama199565}. We can assess the relevance of this advice
+to specific projects by checking how often a planner changes the
+ coupling-related features:
+\bi
+\item {\em ca}:   afferent couplings  =  \# classes using this
+				class;
+\item {\em  	ce}:  efferent couplings =  \# classes  used by this
+				  class. 
+\item {\em cbm}: coupling between methods =  \# new/redefined methods
+				to which all the inherited methods are coupled
+\item	{\em cbo}:  coupling between objects = a value that increases when the methods of one
+				class access services of another.
+
+\item {\em ic}:   inheritance coupling =  \# parent classes  which a given
+				class is coupled (including methods and variables inherited)
+\ei
+In many of our results with the Jureczko data, it was indeed true that the changes
+proposed by XTREE lead to lower coupling. However, the {\em lucene} results were quite
+different and rather surprising.
+
+In the {\em lucene} XTREE results  of  \fig{changed},
+the most frequent change was to alter the lines of code in a class (see the tallest
+red histogram in that figure on the {\em loc}, or lines of code). 
+Looking at the logs of our planner, we can see
+that XTREE's proposed change is to {\em reduce} the size of a class.
+The only way to do that,  while keeping the  same functionality, is to create a network
+of smaller classes that interact to produce that functionality.
+That is, we would need to {\em increase} the coupling of those classes to achieve XTREE's plan.
+
+In theory, increasing coupling between classes complicates and confuses a class design.
+But the {\em lucene} XTREE results  of  \fig{changed} rarely proposes changing   the
+coupling features {\em ca, ce, cbm, ic} (in fact, XTREE never proposes any change to {\em ic}).
+
+The only coupling issue that XTREE   usually adds to its plans is {\em cbo} (which appears 55\% of the time
+in \fig{changed}). But note  that this is {\em object} coupling measure, not class coupling. So here
+XTREE is warning against, say, some factory class generating a large
+community of agents, all of the same class, who
+co-ordinate on some task. This is a different issue to the class redesign issue that would
+be triggered by altering {\em loc}.
+
+In summary, XTREE satisfies that criteria that, sometimes, it produces surprising plans.
+At least for the {\em lucene} data set, we can see advice that recommends {\em increasing coupling}
+to reduce defects.

### 2016-03-10 Manuscript.tex
Updates from ShareLaTeX

@@
 
+\usepackage{rotating}
 %\usepackage{stfloats}
@@
 
-\title{  How to Recognize  a Bad  ``Bad Smell'' using Contrast Sets}
+\title{  Which ``Bad Smells'' can be Ignored?}
 \numberofauthors{2} 
@@
 \begin{abstract} 
-It is hard to assess
-if a ``bad smell'' is irrelevant or misleading
-for a specific project.
+The literature lists dozens of bad smells which might
+be detrimental to a software system.
+But which are relevant to the current project
+and which can be ignored?
 Developers, text books, and tools
@@
 Specifically, it is unclear at what threshold values
-for which metrics should trigger code reorganization  since
+a measurement should trigger code refactorings  since
 most studies do not offer
-both metrics {\em and} a threshold for their proposed bad smells.
-This paper proposes the XTREE contrast set
-learner to assess bad smells
-and reject any that are not supported by
-evidence in the historical record.  
-Using XTREE, we   show that the range of useful
-``bad smells'' changes from project to project. Hence,
-before refactoring, it is important to run a critique tool
-like XTREE to stop work on useless code reorganizations.
-Further, XTREE shows that many
-of the bad smells defined in standard texts and tools
-are not useful for particular projects. 
-Based on these results, we endorse  
-guiding code reorganization via bad smells, but one must empirically validate the importance of specific bad smells within the context of each project using an approach like XTREE.
+{\em both} measurable metrics {\em and} a threshold for their proposed bad smells.
+
+This paper proposes   XTREE, a tool to  prioritize bad smells. Given an historical
+log of defects
+seen previously in the code, XTREE
+  deprioritizes   bad smells  that have little to no support
+  in that log.
+
+  This paper compares the bad smells found by XTREE to
+  those found by previous work by Shatnawi, Alves, and Borges. 
+  If code modules are  changed to avoid the bad
+  smells found by XTREE,  that changed system
+  has significantly fewer expected defects.
+  Also, XTREE   endorses
+  changes to far fewer code metrics.
+  Hence, it is 
+  a better tool for ruling out irrelvant bad smells.
+  For example, the XTREE results show that many of the bad smells learned 
+  by Shatnawi, Alves, and Borges can be ignored.
+
+Based on these results, we endorse using bad smells to  
+guiding refactoring. However, not all bad smells
+are   relevant to the current project.
+Hence, it is important
+to  prioritize the importance of specific bad smells within the context of each project.
 
@@
 as triggers for code change.
-In our own work, documented later in this paper, we  show
-that tools and text books and developers can disagree on what bad smells
+As we show in this paper,  
+ tools and text books and developers can disagree on what bad smells
 are important enough to hunt down and eliminate. Our findings
@@
 test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
-In that approach, developers are always asking themselves ``should I reorganize this code''?
-Bad smells are a widely used  technique for deciding when to reorganize\footnote{The Fowler text is widely read-- as of  early 2016, ACM Digital Portal reports 
-it has 6434 references. Also, bad smell-based research
-is widely reported-- a search  on {\em "bad smell"+programming} 
+In that approach, developers are always asking themselves ``should I refactor this code''?
+Bad smells are a widely used  technique for deciding when to refactor\footnote{Evidence: As of  early 2016, ACM Digital Portal reports 
+the Fowler text has 6434 references. Also, bad smell-based research
+is widely reported. A search  on {\em "bad smell"+programming} 
 where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
@@
 Valuable development resources can be wasted if developers
-change their code based on   bad smells that are ``bad''; i.e. 
+refactor their code based on   bad smells that are ``bad''; i.e. 
  irrelevant or misleading
@@
 
-It turns out that it is hard to learn ``good'' bad smell detectors. Such a detector
+It turns out that it is hard to learn ``good'' bad smell detectors.
+As discussed on the next page, programmer
+   cognitive biases result in different
+developers declaring that
+bad smell X or Y is most important.
+
+Also, a bad smell  detector
 is a combination of some code metric (e.g. lines of code in a method)
@@
 
+Further, it is hard to judge the  effects of removing bad smells
+for three reasons.
+Firstly, the bad smell learners currently found in the literature
+report bad smells for each code metric {\em in isolation}
+Refactored code cannot be assessed just by a rerun of the the test
+suite since
+code refactorings is meant to
+preserve the semantics of the code (i.e. the same tests
+should pass before and after the refactoring). 
+It is also problematic to assessing refactorings by making the
+refactorings,  then seeing what happens to that code for the rest
+of the project. In an test-driven agile project, as
+SCRUM meetings change what stories are implemented next,
+the operational profile that lead to defects in the old
+code may not repeat.
+
+To address the first problem of programmer cognitive biases,
+this paper proposes using the   XTREE contrast set learner to
+find bad smells. Contrast set learners   recommend {\em what to change} in order to move from
+     one classification to another (e.g. from defective=yes to
+     defective=no).
+     When applied to defect predictors from static code measures,
+that delta is a change to static code measures;
+i.e. a code refactoring. Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little to no support in that log.
+
+Note that XTREE also addresses the second problem since it reports
+code metrics and useful thresholds that
+
+To resolve this third problem, this paper adopts
+a technique used by Harman et al.
+
 Why are there so few papers offer bad smell metrics and thresholds? For the few that
 do exist, why do developers disagree on the relevance of those bad smells?
-As discussed on the next page, one reason for this is that 
-   cognitive biases result in different
-developers declaring that
-bad smell X or Y is most important.
 
@@
 that delta is a change to static code measures;
-i.e. a code reorganization. Hence, constrast set learners
+i.e. a code refactoring. Hence, constrast set learners
 like XTREE can be used to  recognize ``bad''
@@
 \begin{itemize}
-    \item Consider $N$ developers each proposing a software reorganization 
+    \item Consider $N$ developers each proposing a software refactoring 
 based on a set of bad smells $S$.
 \item XTREE will (a)~learn contrast sets
-from the historical record; then (b)~reject  any reorganization based on ``bad'' bad smells, i.e.
+from the historical record; then (b)~reject  any refactoring based on ``bad'' bad smells, i.e.
  on any member of $S_{\mathit{bad}} \in S$
@@
 
-XTREE   has certain limits. Firstly, it is an oracle that  might reject code reorganizations proposed
+XTREE   has certain limits. Firstly, it is an oracle that  might reject   refactorings proposed
 by developers. XTREE does not perform those refactorings.
@@
     as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
-    at avoiding code reorganizations that are irrelevant to the task of reducing the risk of introducing defects.
+    at avoiding refactorings that are irrelevant to the task of reducing the risk of introducing defects.
     \end{itemize}
@@
         that most of the  listed code smells are {\em not} high priority issues
-        requiring code reorganization. 
+        requiring refactoring. 
  \end{itemize} 
@@
 \begin{itemize} 
-\item One way to remove unusually large methods is to reorganize that method's
+\item One way to remove unusually large methods is to refactor that method's
 functionality into several smaller ones. 
@@
 To find an answer to this question, we will assume that developers
-reorganize their code until the bad smell thresholds no longer apply.
-This reorganization will start with some {\em initial} code
+refactor their code until the bad smell thresholds no longer apply.
+This refactor will start with some {\em initial} code
 base that is changed to a {\em new} code base. 
@@
 
-\subsection{Resylts}
+\subsection{Results}
+\input{tex/results_jur.tex}
+% \input{tex/results_shatnawi.tex}
+% \input{tex/results_alves.tex}
 

### 2016-03-10 tex/refactor.tex
Updates from ShareLaTeX

--- 2016-03-08 tex/refactor.tex
@@
-\begin{figure*}[!t] 
+\begin{figure}[!t] 
 \scriptsize
@@
 \begin{tabular}{r|c|c|c|c|c}
-Mostly from Fowler'99~\cite{fowler99}& & & & & Developer\\
- (with a some additions)~\cite{fowler99} & Lanza'06~\cite{Lanza2006} & SonarQube~\cite{sq15}  &  Yamashita'13\cite{Yamashita2013} & Atwood'06\cite{Atwood06}  & Survey 2015\\\hline
+
+\begin{turn}{75}Fowler'99~\cite{fowler99}, and~\cite{Kerievsky2005})\end{turn} &\begin{turn}{75} Lanza'06~\cite{Lanza2006}\end{turn} & \begin{turn}{75}SonarQube~\cite{sq15} \end{turn} &  \begin{turn}{75}Yamashita'13\cite{Yamashita2013} \end{turn}& \begin{turn}{75}Atwood'06\cite{Atwood06}\end{turn}  & \begin{turn}{75} Developer Survey 2015\end{turn}\\\hline
   Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
@@
 \label{fig:smells}
-\end{figure*}
+\end{figure}

### 2016-03-11 Manuscript.tex
Updates from ShareLaTeX

@@
 \documentclass{sig-alternate}
-\usepackage{stfloats}
+%\usepackage{stfloats}
 \usepackage{tikz}
@@
 \usepackage{framed,graphicx,xcolor}
-
+\usepackage{multirow}
 \usepackage{rotating}
@@
 
-This paper proposes   XTREE, a tool to  prioritize bad smells. Given an historical
-log of defects
-seen previously in the code, XTREE
-  deprioritizes   bad smells  that have little to no support
-  in that log.
-
-  This paper compares the bad smells found by XTREE to
-  those found by previous work by Shatnawi, Alves, and Borges. 
+This paper proposes   XTREE, a tool to   prioritize  bad  smells.
+Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little or no support in that log.
+
+When XTREE's recommendations are compared to 
+  those found by previous work (Shatnawi, Alves, and Borges.)
   If code modules are  changed to avoid the bad
@@
 are important enough to hunt down and eliminate. Our findings
-are consistent with several other papers.
-Mantyla et al.~\cite{Mantyla2004} found that within an organization,
- developers had different perceptions on the importance of different
- bad smells.    Yamashita  and Moonen~\cite{Yamashita2013}
- surveyed 85 professional developers to conclude that it remains
- open to debate if code smells are a meaningful conceptualization  of code quality issues (from a developer's perspective). 
- Sj\"oberg et al.~\cite{Sjoberg2013} studied six developers 
- performing maintenance tasks that took three to four weeks to complete. 
- After modifying nearly 300 Java files, it was observed
- that none of the dozen investigated code smells explored in 
- that study were mostly irrelevant to to that maintenance work.
+are consistent with other research suggesting that bad
+smells may not be a useful way to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.
 
@@
 
-It turns out that it is hard to learn ``good'' bad smell detectors.
-As discussed on the next page, programmer
+To address this problem, we propose the XTREE
+tool to judge refactorings
+proposed by developers.
+  XTREE requires:
+     \begin{itemize}
+         \item An historical log of defects seen previously in the code;
+         \item  Bad smells expressed in terms of the 
+          metrics   in that log;
+         \item Some proposed refactorings justified using those bad smells.
+     \end{itemize}
+Given that information,  XTREE can determine which
+  proposed refactorings are be ignored
+since they are not justifiedable  
+with respect to the history of this project.
+
+XTREE addresses four problems that  make it hard to  learn useful
+bad smells for a particular project.
+
+{\em PROBLEM \#1:} As discussed in the next section, programmer
    cognitive biases result in different
@@
 bad smell X or Y is most important. 
-
-Also, a bad smell  detector
-is a combination of some code metric (e.g. lines of code in a method)
-and a threshold measurement (e.g. {\em loc} $>$ 100). 
-Shantnawi~\cite{Shatnawi10} warn that  there are most  studies do not offer
+ XTREE offers a way to assess
+ competing claims about different refactorings.
+
+{\em PROBLEM \#2:} Not all bad smells can be fixed. 
+Consider
+a bad smell that is triggering on some very long method,
+yet that method must be passed as a single closure
+to a third-party GUI toolkit.  This is one
+example of a common problem-- when new code must
+be combined with code from other developers,
+sometimes the constraints imposed by the other code
+lead unavoidably to bad smells. In this case,
+a refactoring proposed to reduce that bad smell
+may be irrelevant since it cannot be applied to this code base.
+
+To address this second
+  problem, we let
+XTREE {\em reason optimistically} to find the set
+$R_0$ of refactorings that might be useful
+(for details on how that optimistic reasoning
+is implemented, see \tion{rq}).
+Since that set was generated optimistically,
+the space of actually usable refactorings is 
+the possibly smaller subset $R_1 \subseteq R_0$.
+If the developer's proposed refactoring $R_2$ is
+not in $R_0$, then it cannot be in the smaller
+set $R_1$.  As shown
+in this paper's experiments,  XTREE's
+recommendations typically include only a small subset of
+the code metrics. That is, we
+can still rule out some proposed refactorings
+since many of those proposals fall outside of the space
+of refactorings suggested by XTREE.
+
+{\em PROBLEM \#3:}  
+Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
 {\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
@@
 depth of inheritance tree) but they never comment on what thresholds
-means that some section of code starts to smell bad. After
-an extensive literature review on OO bad smells, Shantnawi remarks that there is a dearth
-of  studies  conducted to formulate bad smell guidelines,
-represented as metrics plus threshold values, that interpret the complexity of the software design using OO code metrics.
-
-Further, it is hard to judge the  effects of removing bad smells
-for three reasons.
-Firstly, the bad smell learners currently found in the literature
-report bad smells for each code metric {\em in isolation}
+means that some section of code starts to smell bad.  
+In all our  reading of the recent literature,
+the only work we can find from other researchers that
+offers metrics {\em and} thresholds come
+from Shatnami~\cite{Shatnawi10}
+and  Alves et al.~\cite{Alves2010} (for details on those
+methods, see \tion{methods}).
+One potential problem with their approaches is that 
+they do not explore the interconnections between sets
+of code metrics. This is important if sets of
+code metrics have to be changed at the same time.
+For example, it is not possible to refactor a
+long method it several smaller ones without increasing
+the coupling metrics of a class. 
+More generally, we say that   bad smell thresholds should be
+{\em conjunctions} of metric+thresholds pairs that need to be changed {\em togehter}.
+Yet Shatnami~\cite{Shatnawi10}
+and Alves et al.   report   {\em disjunctions} of multiple metric/thresholds~\cite{Shatnawi10,Alves2010}
+which could be separately and  inappropriately used
+to guide refactoring.
+
+To address this third problem, 
+XTREE   generates {\em conjunctions} of   thresholds
+than should  be considered at the same time.
+
+{\em PROBLEM \#4:} It is hard to judge the  effects of removing bad smells.
 Refactored code cannot be assessed just by a rerun of the the test
@@
 should pass before and after the refactoring). 
-It is also problematic to assessing refactorings by making the
+Also, it many not be practical
+ to assessing refactorings by making the
 refactorings,  then seeing what happens to that code for the rest
-of the project. In an test-driven agile project, as
+of the project. In a test-driven agile project, as
 SCRUM meetings change what stories are implemented next,
 the operational profile that lead to defects in the old
-code may not repeat.
-
-To address the first problem of programmer cognitive biases,
-this paper proposes using the   XTREE contrast set learner to
-find bad smells. Contrast set learners   recommend {\em what to change} in order to move from
-     one classification to another (e.g. from defective=yes to
-     defective=no).
-     When applied to defect predictors from static code measures,
-that delta is a change to static code measures;
-i.e. a code refactoring. Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little to no support in that log.
-
-Note that XTREE also addresses the second problem since it reports
-code metrics and useful thresholds that
-
-To resolve this third problem, this paper adopts
-a technique used by Harman et al.
-
-Why are there so few papers offer bad smell metrics and thresholds? For the few that
-do exist, why do developers disagree on the relevance of those bad smells?
-
-When developers cannot fully assess the value of bad smells,
-it is possible to use automatic means.
-
-This paper proposes the XTREE contrast set learner for recognizing
-``bad'' bad smells. 
-Contrast set learners   recommend {\em what to change} in order to move from
-     one classification to another (e.g. from defective=yes to
-     defective=no).
-     When applied to defect predictors from static code measures,
-that delta is a change to static code measures;
-i.e. a code refactoring. Hence, constrast set learners
-like XTREE can be used to  recognize ``bad''
-bad smells:
-\begin{itemize}
-    \item Consider $N$ developers each proposing a software refactoring 
-based on a set of bad smells $S$.
-\item XTREE will (a)~learn contrast sets
-from the historical record; then (b)~reject  any refactoring based on ``bad'' bad smells, i.e.
- on any member of $S_{\mathit{bad}} \in S$
-that XTREE could {\em not} be found in the historical record.
-\end{itemize}
-Note that we    recommend  XTREE  is used to critize, but  not to create,
-bad smells. 
-XTREE uses frequency counts of code features seen in the historical record\footnote{From LL: Note that ``new'' bad smells may not have a historical record.
-A limitation to be acknowledged and a reason why developer knowledge is still relevant.}
-to generate
-predictions about the effects of changing code
-attributes. Such predictions may be inaccurate due to the infamous correlation-vs-causation problem
-where spurious correlations can suggest effects that are not
-truly causal.
-
-XXX
-
-Nevertheless, XTREE is a still a useful tool to critique and prune
-irrelevant bad smells since, 
-as show below, many of the bad smell detectors offered
-in the literature (and supported by readily available tools)
-fall outside of  space of
-effects found  using XTREE's analysis.
-That is,   XTREE can find 
-{\em neither} correlation {\em nor}
-causal evidence that those bad smells
-have any support   in the historical record.\footnote{From LL: You need to say what ``support from the historical record'' is early in the paper.
-Are you mapping code smells in source files to defects?
-Code smells to performance regressions?
-Whatever it is, give an example.}
-
-XXXX
-
-The same cannot be said for other methods that try to learn
-bad smell thresholds. \tion{outs}
-discusses {\em outlier} methods and {\em centroid delta} methods
-for learning bad smell thresholds.
-The {\em  outlier methods}  used by
-Shatnawi, Alves and Herman et al.~\cite{Shatnawi10,Alves2010,hermans15}
-propose thresholds on all code metrics with outlier values.
-Similarly, the older {\em centroid delta} methods  used in our 
-prior work~\cite{me12c} also returns thresholds on most of the code
-metrics. That is, outlier methods and centroid deltas methods
-seem to be saying that many bad smell thresholds are possible.
-
-we proposed CD, short for
-``Centroid Deltas'', as a way to find bad smell thresholds:
-\begin{itemize}
-    \item Clustered
-data and generated centroids for each cluster;
-\item Computed contrast sets (i.e. the bad smells)
-as the delta between attributes in adjacent centroids $C_1 - C_2$
-where examples in $C_1$ had more defects than $C_2$.
-\end{itemize}
-As shown below, Centroid Deltas
-generates change
-recommendations that are far less effective than XTREE.
-Worse still Centroid Deltas reported bad smells
-for far more attributes than XTREE. Hence, that prior work
-is less useful than XTREE at judging the  bad smells proposed by developers.
-
-XTREE   has certain limits. Firstly, it is an oracle that  might reject   refactorings proposed
-by developers. XTREE does not perform those refactorings.
-Secondly, XTREE assumes   Fowler's bad smells
-should be assessed via their contribution to software defects
+code may not be repeated in future executions. 
+To resolve this  problem, SE researchers such as 
+Cheng et al.~\cite{Cheng10}, O'Keefe et al.~\cite{OKeeffe08,OKeeffe07},
+Moghadam~\cite{Moghadam2011} and Mkaouer et al.~\cite{Mkaouer14}
+use a pair of systems: 
+\begin{enumerate}
+    \item One system to propose changes;
+\item Another, completely separate, system to assess
+how defective is the code before and after some
+refactoring.
+\end{enumerate}
+For the second system,
+ Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
+quality model~\cite{Bansiya02}, constructed in 2002.
+Our problem with using tools like QMOOD
+is that   quality models learned from other projects,
+may perform poorly when applied to new projects~\cite{me12d}.
+
+To solve this fourth problem, this study  eschews
+older quality models like QMOOD.
+Instead, we use
+Random Forests~\cite{Breiman2001} to learn defect predictors
+that are specific to the systems   studied here.
+
+ As to XTREE's limitations, the tool 
+  can only comment
+on refactorings expressed in the   terminology
+as the data
+  in the historical log of a project.
+If developers want to justify their refactorings
+via bad smells expressed in other terminology,
+then a XTREE-style analysis must   wait till 
+data about those new
+terms has been collected.
+
+Another  aspect to note about XTREE  is that the tool assumes
+Fowler's bad smells
+should be assessed only via their contribution to software defects
 (evidence: see the section in italics of the quote
@@
 
-The contributions of this paper are
-\begin{itemize}
-
-    \item A new kind of contrast set learner called XTREE
-    that   critique plans of what to change   in order
-    to improve outcomes.
-    \item An experimental rig for assessing the value of tools like
-    XTREE that propose changes to software. That rig uses data mining
-    to build an oracle for the proposed changes.
-    \item Using that learner and that oracle,
-    a demonstration that XTREE is better than our prior approach~\cite{me12c}
-    as well as   state of the art methods~\cite{Shatnawi10,Alves2010,hermans15}
-    at avoiding refactorings that are irrelevant to the task of reducing the risk of introducing defects.
-    \end{itemize}
-    Based on the above, we can offer
-    a resolution between proponents of bad smells (e.g.
-Fowler, Shantnawi, Tufano~\cite{fowler99,Shatnawi10,Tufano2015}) and those who find them not useful or confusing   (e.g.Mantyla, Sj\"oberg,Yamahista et al.~\cite{Sjoberg2013,Mantyla2004,Yamashita2013}).
-We say
-    that bad smells can guide code change, but only if those bad smells are
-    first carefully assessed.
-
-The rest of this paper discusses the cognitive
+The rest of this paper is structured as follows. XXX.
+discusses the cognitive
 biases which make developers disagree on the relevance of those bad smells to
@@
 
-\input{tex/refactor}
+\subsection{Contributions of this Paper }
+The XTREE learner has not been reported
+previously.
+ The experimental rig
+ used to comparatively assess
+    XTREE is another novel contribution. Using the design
+    of that rig,
+    other researchers could confirm/improve/refute
+    the conclusions of this paper.
+    Another contribution here is
+    the comparative analysis of XTREE
+    versus previously published approaches to learning
+    bad smell detectors. that comparion shows
+    that XTREE   rarely perfoms
+      worse that the other approaches~\cite{Shatnawi10,Alves2010}
+      and often is it   much better at proposing
+      changes that reduce the expected
+      value of the defects. Further, the number
+    of code metrics mentioned by XTREE is far
+    fewer than other approaches This means that XTREE is a better
+    tool for ruling out   refactorings proposed by
+    developers.
+
+As to the relation to prior work, we
+show here that XTREE is a  superior 
+to the CD tool   developed previously with Borges~\cite{me12c}.
 
@@
 
+\input{tex/refactor}
+
 If the above remarks hold true for bad smells, then we would expect
@@
 Two of the studies in \fig{smells} offers some comments on the relative importance
-of the different bad smells the   Yamashita'13 study and the developer study:
-    \begin{itemize}
-        \item Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey. Those three were {\em duplicated code, large class}, 
+of the different bad smells the   Yamashita'13 study and the developer study.Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey.
+Those three were {\em duplicated code, large class},
         and {\em long method}. 
-        \item
  Note that this agreement also means that the
-  Yamashita'13 study and the {\em developer survey}   concluded
-        that most of the  listed code smells are {\em not} high priority issues
+  Yamashita'13 study and the {\em developer survey}   
+  believe
+        that very few  code smells are   high priority issues
         requiring refactoring.
- \end{itemize} 
+
 In summary, just because a developer strongly believes something
@@
 can disagree on which bad smells are important.
-Hence it is important to have some automated tool
-that can assess their beliefs in, for example,
+Special tools are needed to assess their beliefs in, for example,
 bad smells.
@@
 The SE literature offers  two ways to 
-learning bad smell thresholds:
-\begin{itemize}
-    \item Approaches based on {\em outlier statistics}~\cite{erni96,bender99}
-    as used by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
-    and Hermans et al.~\cite{hermans15};
-    \item  Approaches based on {\em cluster deltas} that we developed
-    for   Centroid Deltas~\cite{me12c} and XTREE.
-\end{itemize}
-\subsection{  Outliers}
+learning bad smell thresholds.
+One way relies on 
+{\em outlier statistics}~\cite{erni96,bender99}. This way
+has been used   by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
+    and Hermans et al.~\cite{hermans15}.
+    Another approach is 
+    based on {\em cluster deltas} that we developed
+    for   Centroid Deltas~\cite{me12c} and 
+    use here for XTREE.
+
+These two approaches are discussed below.
+\subsection{  Outliers}\label{sect:p}
 
@@
 Given classes described with the  code metrics of \fig{ck},
-Enri and Lewerentz finds the   mean $\mu$ and the standard deviation $\sigma$
+Enri and Lewerentz found the   mean $\mu$ and the standard deviation $\sigma$
 of each
@@
 Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
-using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, there is a lack of empirical validation of this method.
+using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.
 
@@
 thresholds which, following Shatnami's guidance, we set to
-$p_0=p_1=0.05$. VARL encodes the defect count
+$p_0=p_1=0.05$.
+
+VARL encodes the defect count
 for each class to 0 (for no defects known) to 1 (for defects known).
-Univariate binary logistic regression is applied to learn three coefficients: $\alpha, \beta,p_0$ being
-the  constant intercept ($\alpha$), the coefficient for maximizing log-likelihood function ($\beta$),
-and a measure of how well this regression model predicts for the defects ($p_0$).
-Any code metric with $p_0>0.05$ is then rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
-the equation proposed by Bender:
+Univariate binary logistic regression is applied to learn three coefficients:  
+ $\alpha$ is  constat intercept;
+    $\beta$ is the coefficient for maximizing log-likelihood;
+  and $p_0$  
+    measures   how well this   model predicts for   defects.
+Any code metric with $p_0>0.05$ is  rejected as being a poor defect predictor. Thresholds are then learned from the surviving metrics $M_c$ using
+an equation proposed by Bender:
 \begin{equation}
@@
 Alves et al. then compute their 
-thresholds are then derived by choosing the percentage of the overall code that needs to be represented.
+thresholds   by choosing the percentage of the overall code that needs to be represented.
 For instance, Alves et al.
@@
 The intuition here is that the worst
-code are the outliers beyond 90\% of the normal code measurements.
+code have outliers beyond 90\% of the normal code measurements.
 
@@
 
-\begin{figure*}[!b]
+\begin{figure*}[!t]
 ~\hrule~
@@
 
-\subsubsection{Discussion of the Outlier Approach}
+\subsubsection{Discussion }\label{sect:disc}
 The advantage of the outlier
@@
 For example, in less than a day we could
-code up and apply all the approaches of Shatnami and 
-Alves et al.
-
-On the other hand, these outlier approaches have major two disadvantages. 
+code up and apply  the Shatnami and 
+Alves et al. methods.
+
+The outlier approaches has   two  major disadvantages. 
 Firstly, they are {\em verbose}; i.e.  they can
@@
 
-Secondly, the outlier approaches described above are all 
-{\em univariant} which means
-they do  not understand
-that sets of code metrics must
-be changed at the same time.
-For example:
+Secondly, the outlier approaches suffers from
+the {\em disjunction} problem discussed in the introduction.
+That is, while
+ they propose thresholds for many code metrics
+individually, they make no comment on what
+metrics need to be changed at the same time.
+
+\subsection{Cluster Deltas}
+
+Cluster deltas are a general method
+for learning {\em conjunctions} of changes
+that need to be applied at the same time. 
+This approach works as follows:
 \begin{itemize}
-\item One way to remove unusually large methods is to refactor that method's
-functionality into several smaller ones. 
-\item If that is done,
-this means that {\em reducing} lines of code in one method
-will {\em increase} the coupling of the methods of that class to
-either itself or to other classes (depending on whether or not
-the new smaller methods are placed in the original class, or in other classes).
-\item
-Hence, it can be counter-productive to act on bad smell thresholds for {\em both}
-large lines of code {\em and} seemingly excessive coupling since
-one cannot be reduced without increasing the other.
-\end{itemize}
-
-\subsection{Cluster Deltas}
-
-\subsubsection{CD}
-Borges and Menzies first proposed the CD centroid delta approach to avoid the drawbacks of univariant reasoning~\cite{me12c}.
-Centroid deltas are a {\em multivariant} method which
-work  as follows:
-\begin{itemize}
-    \item Cluster data about code modules;
+    \item Cluster the data. 
     \item Find
@@
 \end{itemize}
-Note that $\Delta$ is a multivariant recommendation
-that proposes multiple changes in order to reduce defects.
+Note that $\Delta$ is a conjunction of  recommendations.
 Since it is computed
@@
 repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
-without also increasing a coupling measure.
-
-The drawback with with CD is that it is very {\em verbose}
+without also increasing a coupling measure
+
+Cluster deltas are used in CD~\cite{me12c} and XTREE.
+
+\subsubsection{CD}
+Borges and Menzies first proposed the CD centroid delta approach to
+generate {\em conjunctions} of code metrics
+that need to be changes at the same time
+in order to reduce defects~\cite{me12c}.
+CD used the WHERE clustering algorithm developed by the
+authors for a prior application~\cite{localvsglobal}.
+Each cluster was then replaced by its centroid
+and $\Delta$ was calculated directly from the difference
+between code metrics values between one centroid
+and its nearest neighbor.
+
+As shown below, one drawback with with CD is that it is very {\em verbose}
 since
@@
 This makes it hard to use CD to   critique and prune away bad smells.
-\subsubsection{XTREE}
+
+Another drawback with CD is revealed by the comparison experiments
+in this paper: CD never worked as well as XTREE in proposing
+changes to reduce defects.
+
+\subsubsection{XTREE: Overview}
 
@@
   Decision trees are built via
-  {\em iterative dichotomization} (ID).
-  ID   finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature. The data is then divided on that split
-  and ID is called
+  {\em iterative dichotomization} that  finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature.
+The data is then divided on that split
+  and iterative dichotomization is called
   recursively on the data in each split. In this way,
@@
 
-  The CART iterative dichotomization algorithm
-  measures the variability of continuous attribites
+  The CART iterative dichotomizer~\cite{breiman84}
+  measures   variability  
   using the  standard deviation $\sigma$ of the defect counts.
@@
 
-  While not usually labelled as such, ID can
+  While not usually labelled as such, CART can
   be viewed as a clustering algorithm that groups together code modules
   with similar defect counts and some shared ranges of code metrics.
-  XTREE is a post-processor to ID that reports the delta between 
+  XTREE is a post-processor to CART that reports the delta between 
   nearby groups with different defect counts. \fig{xtree_samp} offers
-  a small example of that reflection. In this example:
-  \begin{itemize}
-      \item Some project has a table of data describing its  classes in the form of \fig{ck}.
-      \item After code inspections and running test cases or operational
+  a small example of that reflection. In this example, some project has a table of data describing its  classes in the form of \fig{ck}. After code inspections and running test cases or operational
       tests, each such class is augmented with a defect count.
-      \item CART takes that table of data and, using iterative dichotomization,
+      CART takes that table of data and, 
         generates the tree of \fig{xtree_samp}.
-      \item Once the tree is built, some test case arrives and is passed
+
+     Once the tree is built, some test case arrives and is passed
       down the tree to a leaf note (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}.
-      \item XTREE then looks for a nearby leaf node with a lower defect
+       XTREE then looks for a nearby leaf node with a lower defect
       count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that test case, XTREE proposes a bad smell
@@
       \textcolor{green}{{\bf green}} and \textcolor{orange}{{\bf orange}}.
-  \end{itemize}
-
-\subsubsection{XTREE: the details}
-
-Firslty, using the training data,  XTREE uses iterative dichotomization to
-  divide a table of $N$ code modules  into  groups of
+
+\subsubsection{XTREE:   Details}
+
+Using the training data,  XTREE uses CART to
+  divide  $N$ code modules  into  clusters of
 size $\alpha=\sqrt{N}$.
@@
 Next, for each test code module, 
-	  find the {\em current } leaf: take each test , run it down to a leaf in the decision tree.  
-After that,	  find the {\em desired} leaf:
+	  find $C_+$ : take each test , run it down to a leaf in the decision tree.  
+After that,	  find $C_-$ as follows:
 		\begin{itemize}
-		\item Starting at {\em current}, ascend the tree $lvl\in \{0,1,2...\}$ levels;
+		\item Starting at the  $C_+$ leaf, ascend   $lvl\in \{0,1,2...\}$ tree levels;
 		\item Identify {\em sibling} leaves; i.e. leaf clusters that can be reached from level $lvl$ that are not same as {\em current }
@@
 		    return nil  if the new $lvl$ is above the root. 
-		\item  Return the {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
+		\item  Set $C_-$ to the  {\em closest} better sibling where distance is measured between the mean centroids of that sibling and {\em current}
 		\ei
-	 Also, find the {\em delta}; i.e. the set difference between  conditions in the decision tree branch to {\em desired} and {\em current}. To find that delta,
-	 for discrete attributes, delta is the value from {\em desired};
-	   for  numerics, delta is the numeric difference;
-	   and  for numerics  discretized into ranges, delta is a random number selected from the low and high boundaries of the that range. 
-	     The bad smell threshold for   is then the delta that would move
-	     the test case from the current to desired leaf.
+	 Now find $\Delta = C_+ - C_-$  by reflecting
+	 on the set difference between  conditions in the decision tree branch to $C_-$ and $C_+$. To find that delta,
+	 for discrete attributes, the delta is the value from {\em desired};
+	    for numerics  expressed as ranges, the delta is a random number selected from the low and high boundaries of the that range.
+
+\input{tex/prediction_stats.tex}
 
@@
 \subsection{Research Questions}
-
-Theoretically, XTREE is a good way to learn bad smell thresholds:
-\begin{itemize}
-    \item Unlike the outlier approach, XTREE offers bad smell
-    thresholds that are specific to where test cases fall into
-    the leaves of a decision tree that predicts for defects.
-    That is, while the outlier approach offers one threshold for all
-    metrics, XTREE specializes those thresholds to particular regions of
-    the code.
-    \item
-    Unlike the outlier approach, XTREE is a multivariant approach
-    that  understands when several code
-metrics should  be changed in unison. 
-\item Unlike CD, it is less verbose. 
-XTREE does not report all the deltas between
-adjacent clusters. Rather, it just reports the deltas in the code
-metrics that a decision tree learner has determined are the minimun
-changes required to distinguish one grouping from another.
-\end{itemize}
-Both these claims need to be tested, which leads to the following research
-questions.
-
-{\bf RQ1: Effectiveness.} {\em In a comparison of  the outlier and cluster delta
-methods described above, which one reduces defects the most?}
+To define our experiments, we  review some details
+about XTREE.
+
+ XTREE proposes
+conjunctions of changes that might perform better than
+the disjunctions proposed by Shatnami and Alves et al.
+Further,
+unlike CD, XTREE   should be less verbose since
+it just reports the deltas in the code
+metrics selected by a decision tree learner.
+
+In order to test the value of the above,
+we propose three research questions.
+
+\subsubsection{ RQ1: Effectiveness} \label{sect:rq}
+
+{\em In a comparison of   outlier and cluster delta
+methods, which one most reduces the expected number
+of defects?}
 
@@
 This refactor will start with some {\em initial} code
-base that is changed to a {\em new} code base. 
-For example, if the bad smell is {\em loc $>100$} and a 
-code module has (e.g.) 500 lines of code, we conduct a ``what-if''
-query and create a {\em new} code base with that module having {\em loc=100}.
+base which is changed to a {\em new} code base. 
+For example, if the bad smell is \mbox{{\em loc $>100$}} and a 
+code module has (e.g.) 500 lines of code, we reason
+optimistically that we can change that code metric
+to 100.  
 Using a defect
-predictor learned from static code metrics, we then predict the
-number of defects in $d_0,d_1$ in {\em initial} and {\em new},
-respectively. In this experiment,
-the best bad smell threshold detector is the one that maximizes
+predictor learned via Random Forest, we then predict the
+number of defects in $d_0,d_1$ in {\em initial} and {\em new}.  
+The best bad smell threshold detector is the one that maximizes
 \begin{equation}\label{eq:diff}
@@
 
-{\bf RQ2: Verbosity.}   {\em In a comparison of  the outlier and cluster delta
-methods described above, which one proposes the fewest thresholds?}
+\subsubsection{RQ2: Verbosity}
+
+{\em In a comparison of    outlier and cluster delta
+methods , which one proposes the fewest thresholds?}
 
@@
 
-{\bf RQ3: Co-change.}   {\em Does XTREE respect code
-metrics that need to be changed together? }
+\subsubsection{RQ3: Co-change}
+
+{\em Does XTREE understand what
+metrics get changed together? }
 
@@
 reduced, does some coupling measure increase?
-
-\input{tex/prediction_stats.tex}
 
@@
 
+\newpage
+\subsection{Assessment Oracles}
+\label{sect:eval}
+Prior work~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} 
+used the QMOOD hierarchical quality model ~\cite{Bansiya02} to assess the results of their refactorings
+Since the QMOOD method is quite old (2002), we  used Random Forests~\cite{Breiman2001} to build
+the assessment methods.
+
+  Random Forests are a decision tree learning method but
+  instead of building one tree, hundreds are built using
+  randomly selected subsets of the data. The final predictions
+  come from averaging the predictions over all the trees.
+  Recent students rank Random Forests as  the top-most
+  effective defect prediction learning algorithm~\cite{lessmann}.
+
+\fig{j} shows   studies with Random Forests and
+the Jureczko data.
+Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}.
+Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
+In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
+
+In order to identify the presence (or absence) of defects, we can   use Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
+For such data, the quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
+``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
+``pf''): the percent of non-fault classes that are {\em predicted} to be defective.
+
+The ``untuned'' columns of \fig{j}
+show a preliminary study.
+Random Forests was built from the training data, then applied to the test data.
+We call a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
+For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% (determined from standard results in other publications~\cite{me07b}).
+No  data set meet
+that criteria.
+
+Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
+an over-sampling of minority-class examples and an under-sampling of majority-class examples.
+Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc.
+The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
+For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
+However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
+
+In summary, while we cannot trust predictors from some of our Jureczko data sets,
+we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
+Accordingly, when this study explores the Jureczko data, we will use these five data sets.
+
+(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
+parameter tunings, those techniques were applied to the training data and {\em not}
+the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+
 \subsection{Statistics}
-Three of our methods are deterministic (Shatnami, Alves et al.,
-and XTREES) while the CD uses some stochastic processing
-in its clustering (CD uses the  WHERE clusterer~\cite{Menzies2013:local}). Hence, to compare all
+Two of our  methods  (Shatnami, Alves et al.) are fully deterministic
+(which is to say, if we run them $N$ times, we will always get the same result).
+However, our other methods use some random choices: CD uses the  stochastic WHERE clustering algorithm ~\cite{Menzies2013:local}
+while XTREE non-deterministically picks thresholds randomly from
+the high and low boundary of a range. 
+ Hence, to compare all
 four methods, we must run the analysis many times.
+
 We use  40 repeated runs, each with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
-Each run collects the \fig{diff} values
-uses defect predictions make for the {\em new} and {\em initial} code.
-
-To rank these 40 numbers collected from CD, XTREE,Shatnami, and Alves et al.), we  use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}. 
+Each run collects the \eq{diff} values.
+
+To rank these 40 numbers collected from CD, XTREE,Shatnami, and Alves et al., we  use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}. 
 In accordance to that test, using the median values of each method, we
@@
 \rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
-1 &      XTREE &    62  &  6 & \quart{47.2}{4.8}{49.6}{115}  \\
-\hline 2 &      CD &    44  &  18 & \quart{28}{14.4}{35.2}{115} \\
-2   & Alves        &    40  & 20  & \quart{30}{10}{50}{10}\\
+1 &      XTREE &    62  &  6 & \quart{67.2}{7}{69.6}{115}  \\
+\hline 2   & Alves        &    40  & 20  & \quart{45}{10}{50}{10}\\
+\hline 3 &      CD &    24  &  18 & \quart{28}{14.4}{35.2}{115} \\
+\hline 4   & Shatnawi        &    15  & 10  & \quart{28}{5}{30}{10}\\\hline
 \end{tabular}}
@@
 The above example table has a  left-hand-side  {\bf Rank} column, computed using the
-Scott-Knott test described above. In this example table, XTREE is ranked the best, while CD and Alves  are ranked the worst.
-
-\subsection{Orcales}
-\label{sect:eval}
-
-  An important part of this process is having an oracle
-  that can assess the defect potential of the {\em initial}
-  and {\em new code}. This section describes how we built
-  that oracle, using Random Forests~\cite{Breiman2001}
-  from the SciKit learn kit~\cite{Pedregosa2012}.
-  Random Forests are a decision tree learning method but
-  instead of building one tree, hundreds are built using
-  randomly selected subsets of the data. The final predictions
-  come from averaging the predictions over all the trees.
-  Recent students rank Random Forests as  the top-most
-  effective defect prediction learning algorithm~\cite{lessmann}.
-
-\fig{j} shows our preliminary studies with Random Forests and
-the Jureczko data.
-Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}, this means that we are training on hundreds to thousands of classes and testing on smaller test suites).
-Note the   \colorbox{lavenderpink}{three bottom} \colorbox{lavenderpink}{rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
-In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
-
-In order to identify the presence (or absence) of defects, we can   use Boolean classes in the  Jureczko data ( \texttt{True} if defects $\gt$ 0; \texttt{False} if defects = 0).
-For such data, the quality of the predictor can be measured using (1) the  probability of detection (a.k.a.
-``pd'' or recall):  the percent of faulty classes in the test data detected by the {\em predictor}; and (2) the  probability of false alarm (a.k.a.
-``pf''): the percent of non-fault classes that are {\em predicted} to be defective.
-
-As a preliminary study, we split the Jureczko  data into train and test groups.
-Random Forests was built from the training data, then applied to the test data.
-The ``untuned'' columns of \fig{j} shows those results.
-We called a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
-For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% to select suitable data sets.
-With this threshold, however, none of our data sets were be suitable for this study.
-
-Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
-Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
-an over-sampling of minority-class examples and an under-sampling of majority-class examples.
-Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
-The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
-For example, in {\em poi}, the recall increased by 4\% while the false alarm rate dropped by 21\%.
-However,  as might have been expected, we could not salvage the data sets in the  three bottom rows.
-
-In summary, while we cannot trust predictors from some of our Jureczko data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-Accordingly, when this study explores the Jureczko data, we will use these five data sets.
-
-(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+Scott-Knott test described above. In this example table, XTREE is ranked the best, while Shatnami is  ranked   worst.
 
 \subsection{Results}
+
 \input{tex/results_jur.tex}
+
+\fig{jur} shows the comparison results.  The Alves and Shatnami
+entries show net effects of applying (seperately) all the disjuncts
+proposed by those methods.
+
+Three data sets are very responsive to defect reduction suggestions:
+Jedit, Ant, and Ivy (all of which show best case improvements over 50\%.
+ The Jedit results are  
+surprisingly uniform  suggesting that in that data set that
+our methods all find the same ways to reduce the expected number of
+defects.
+
+Two data sets are not very responsive to defect reduction:
+Poi and Lucene. The reason for this can be see in \fig{j}:
+both these data sets contain more than 50\% defective modules.
+In that space, all our  methods lack a large sample to
+defect-free examples.
+
+More important that the size of the improvement is the relative
+rank of the different approaches,
+ CD and Shatnawi usually  perform comparatively worse while  XTREE gets top ranked position the most
+number of times. That said, Alves sometimes beats XTREE (see Ivy)
+while sometimes it ties (see Jedit).
+
+\subsection{RQ2}
+\input{tex/rq2.tex}
+
+\subsection{RQ3}
+\begin{figure*} 
+\centering
+\includegraphics[width=\linewidth]{figs/changes0.png}
+\caption{XTREE results.
+While \fig{counts} are the {\em number} of times a code metric was changed,
+\fig{changes} shows {\em how far} each code metric was changed. Each vertical bar
+marks the 27,50,75th percentile change seen in 40 repeats. }
+\label{figs:changes}
+\end{figure*}
+
 % \input{tex/results_shatnawi.tex}
@@
 
-\input{tex/results_models.tex}
+% \input{tex/results_models.tex}
 

### 2016-03-11 tex/prediction_stats.tex
Updates from ShareLaTeX

\begin{figure*}[!t]
	\small
	\begin{center}
		\begin{minipage}{.46\linewidth}
			\begin{tabular}{r@{~}|l@{~}|r@{~}|l@{~}|r@{~}|r@{~}|} \cline{2-6}

				& \multicolumn{5}{c|}{ Data set  properties}\\

				& \multicolumn{2}{c|}{training}   & \multicolumn{3}{c|}{testing}      \\ \cline{2-6}
				data set      & versions           & cases & versions     & cases    & \% defective             \\ \hline
				jedit    & 3.2, 4.0, 4.1, 4.2 & 1257      & 4.3          & 492          & 2 \\
				ivy      & 1.1, 1.4           & 352       & 2.0          & 352          & 11 \\
				camel    & 1.0, 1.2, 1.4      & 1819      & 1.6          & 965          & 19 \\
				ant      & 1.3, 1.4, 1.5, 1.6 & 947       & 1.7          & 745          & 22 \\
				synapse  & 1.0, 1.1           & 379       & 1.2          & 256          & 34 \\
				velocity & 1.4, 1.5           & 410       & 1.6          & 229          & 34 \\
				lucene   & 2.0, 2.2           & 442       & 2.4          & 340          & 59 \\
				poi      & 1.5, 2, 2.5        & 936       & 3.0          & 442          & 64 \\
			 xerces   & 1.0, 1.2, 1.3      & 1055      & 1.4          & 588          & 74  \\ 
			 log4j    & 1.0, 1.1           & 244       & 1.2          & 205          & 92   \\
			 xalan    & 2.4, 2.5, 2.6      & 2411      & 2.7          & 909          & 99  \\\hline

			\end{tabular}\end{minipage}\begin{minipage}{.4\linewidth}
			\begin{tabular}{|rrr|rrr|rr|l} \cline{1-8}

				\multicolumn{8}{|c|}{  Results from learning}\\

				\multicolumn{3}{|c|}{untuned} & \multicolumn{3}{c|}{tuned} & \multicolumn{2}{c|}{change}\\
				\cline{1-8}

				pd & pf & good? & pd & pf & good? & pd & pf\\\cline{1-8}
				\rowcolor{celadon}55 & 29 &   & 64 & 29 & y & 9 & 0&$\star$\\
				\rowcolor{celadon}	65 & 35 & y & 65 & 28 & y & 0 & -7&$\star$\\
				49 & 31 &   & 56 & 37 &   & 5 & 6\\
				\rowcolor{celadon}	49 & 13 & y & 63 & 16 & y & 14 & 3&$\star$\\
				45 & 19 &   & 47 & 15 &   & 2 & -4\\
				78 & 60 &   & 76 & 60 &   & -2 & 0\\
				56 & 25 &   & 60 & 25 & y & 4 & 0\\
				\rowcolor{celadon}	56 & 31 &   & 60 & 10 & y & 4 & -21&$\star$\\
			\rowcolor{lavenderpink}	30 & 31 &   & 40 & 29 &   & 10 & -2&$\times$\\
				\rowcolor{lavenderpink}32 & 6 &   & 30 & 6 &   & -2 & 0&$\times$\\
				\rowcolor{lavenderpink}38 & 9 &   & 47 & 9 &   & 9 & 0&$\times$\\
				\hline 
			\end{tabular}

		\end{minipage}
	\end{center}

	\caption{Training and test {\em data set properties} for  Jureczko data ,
		sorted by \% defective examples.
		On the right-hand-side, we show the {\em results from learning}.
		Data is usable if it has a recall of 60\% or more and false alarm of 30\% or less (and note that, after tuning, there are more usable data sets than before). Results  	\colorbox{celadon}{ marked with ``$\star$''} show large improvements in performance, after tuning
		(lower {\em pf} or higher {\em pd}).
		Data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are  performing
		poorly-- that data so few non-defective examples  that it  is hard for
		our learners to distinguish between classes.
	}\label{fig:j}
\end{figure*}

### 2016-03-11 tex/refactor.tex
Updates from ShareLaTeX

@@
 
-\caption{List of bad code smells from different sources.  Check marks (\checkmark) denote that a bad smell was mentioned.
-Numbers of symbolic labels (e.g. "VH") denote  a priorization comment.
-``?'' indicates lack of consensus on some bad smell. Empty cells
+\caption{Bad   smells from different sources.  Check marks (\checkmark) denote   a bad smell was mentioned.
+Numbers or symbolic labels (e.g. "VH") denote  a priorization comment (and
+``?'' indicates lack of consensus). Empty cells
 denote some bad smell listed in column one that was not found relevant
-in subsequent studies.
-Note that there are many blank cells.}
+in other studies.
+Note:    there are many blank cells.}
 \label{fig:smells}

### 2016-03-11 tex/rq2.tex
Updates from ShareLaTeX

% Please add the following required packages to your document preamble:
% \usepackage{multirow}
% \usepackage[table,xcdraw]{xcolor}
% If you use beamer only pass "xcolor=table" option, i.e. \documentclass[xcolor=table]{beamer}
\newcommand{\ZZ}{.}
\begin{figure*}
\renewcommand{\baselinestretch}{0.8} 
\scriptsize  
\centering

\label{my-label}

\begin{tabular}{c|rrrr|rrrr|rrrr|rrrr|rrrr}
\multicolumn{1}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} }} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Ant}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Ivy}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Lucene}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Jedit}} & \multicolumn{4}{c}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Poi}} \\
\multicolumn{1}{c@{}}{\multirow{-2}{*}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Features}}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} XTREE}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} CD}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Alves}} & \multicolumn{1}{c@{}}{\cellcolor[HTML]{EFEFEF}{\color[HTML]{000000} Shatn}} \\
wmc &\ZZ& 92 & 100 & 100 &\ZZ& 95 & 100 & 100 & 14 & 95 & 100 &\ZZ&\ZZ& 63 &\ZZ&\ZZ&\ZZ& 100 & 100 &\ZZ\\
dit &\ZZ& 77 & 100 &\ZZ&\ZZ& 87 & 100 &\ZZ&\ZZ& 80 & 100 &\ZZ&\ZZ& 72 & 100 & 100 &\ZZ& 46 & 100 &\ZZ\\
noc &\ZZ& 20 & 100 &\ZZ&\ZZ&\ZZ& 100 &\ZZ&\ZZ& 26 &\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ&\ZZ\\
cbo & 83 & 99 & 100 & 100 & 65 & 100 & 100 & 100 & 13 & 94 & 100 & 100 &\ZZ& 100 & 100 &\ZZ&\ZZ& 74 & 100 &\ZZ\\
rfc & 100 & 100 & 100 &\ZZ& 35 & 95 & 100 &\ZZ& 100 & 83 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 & 95 & 100 &\ZZ\\
lcom &\ZZ& 98 & 100 & 100 &\ZZ& 100 & 100 & 100 &\ZZ& 94 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ&\ZZ& 100 & 100 & 100 \\
ca &\ZZ& 93 & 100 &\ZZ&\ZZ& 95 & 100 &\ZZ& 87 & 89 & 100 &\ZZ&\ZZ& 63 & 100 & 100 &\ZZ& 74 & 100 &\ZZ\\
ce &\ZZ& 100 & 100 &\ZZ&\ZZ& 97 & 100 &\ZZ&\ZZ& 90 & 100 &\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 64 & 100 &\ZZ\\
npm &\ZZ& 88 & 100 &\ZZ& 35 & 97 & 100 &\ZZ&\ZZ& 93 & 100 & 100 &\ZZ& 100 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ\\
lcom3 &\ZZ& 90 & 100 &\ZZ&\ZZ& 95 & 100 &\ZZ& 13 & 79 & 100 & 100 &\ZZ& 63 & 100 & 100 &\ZZ& 92 & 100 & 100 \\
loc & 100 & 99 & 100 & 100 & 65 & 97 & 100 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 &\ZZ& 100 & 100 & 100 & 100 & 100 \\
dam &\ZZ& 21 & 100 &\ZZ&\ZZ& 22 & 100 &\ZZ&\ZZ& 55 & 100 &\ZZ&\ZZ& 45 & 100 & 100 &\ZZ& 73 & 100 &\ZZ\\
moa &\ZZ& 67 & 100 &\ZZ&\ZZ& 82 & 100 &\ZZ&\ZZ& 60 & 100 & 100 &\ZZ& 54 & 100 & 100 &\ZZ& 58 & 100 &\ZZ\\
mfa & 17 & 93 & 100 &\ZZ&\ZZ& 90 & 100 &\ZZ&\ZZ& 80 & 100 &\ZZ&\ZZ& 72 & 100 &\ZZ&\ZZ& 72 & 100 &\ZZ\\
cam &\ZZ& 99 & 100 & 100 & 100 & 100 & 100 & 100 &\ZZ& 94 & 100 &\ZZ&\ZZ& 100 & 100 &\ZZ&\ZZ& 98 & 100 & 100 \\
ic &\ZZ& 52 & 100 & 100 &\ZZ& 70 &\ZZ&\ZZ&\ZZ& 68 & 100 &\ZZ&\ZZ& 36 & 100 &\ZZ&\ZZ& 43 & 100 & 100 \\
cbm &\ZZ& 59 & 100 &\ZZ&\ZZ& 85 &\ZZ&\ZZ&\ZZ& 71 & 100 &\ZZ&\ZZ& 36 & 100 & 100 &\ZZ& 67 & 100 &\ZZ\\
amc &\ZZ& 99 &\ZZ&\ZZ&\ZZ& 95 & 100 &\ZZ& 13 & 100 &\ZZ&\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 97 &\ZZ&\ZZ\\
max cc &\ZZ& 87 & 100 & 100 &\ZZ& 85 & 100 & 100 &\ZZ& 71 &\ZZ&\ZZ&\ZZ& 45 & 100 &\ZZ&\ZZ& 63 & 100 &\ZZ\\
avg cc & 17 & 99 & 100 & 100 &\ZZ& 95 & 100 & 100 & 13 & 98 &\ZZ&\ZZ&\ZZ& 100 & 100 & 100 &\ZZ& 92 & 100 & \ZZ\\\hline
\end{tabular}
\caption{Percentage counts of  how often an approach recommends changing a code metric
(in 40 runs). Cells marked with ``.'' indicate  0\%. For the Shatnami and Alves et al.
columns,  metrics are not recommended if they fail the $p \le 0.05$ test of \tion{p}.
For CD, cells are blank when two centroids have the same value for the same code
metrics. For XTREE, cells are blanks when they do not appear in the delta
between branches.  The most important feature of this data is that in the XTREE column, the changes occur to only a small number of code metrics. That is, all the other methods are far more {\em verbose} than XTREE (where ``verbose'' is defined in \tion{disc}).}\label{fig:counts}
\end{figure*}

### 2016-03-11 Manuscript.tex
Updates from ShareLaTeX

@@
 \pagenumbering{arabic}
+
+\usepackage[svgnames]{xcolor}
+\usepackage[framed]{ntheorem}
+\usepackage{framed}
+\usepackage{tikz}
+\usetikzlibrary{shadows}
+%\newtheorem{Lesson}{Lesson}
+\theoremclass{Lesson}
+\theoremstyle{break}
+
+% inner sep=10pt,
+\tikzstyle{thmbox} = [rectangle, rounded corners, draw=black,
+  fill=Gray!20,  drop shadow={fill=black, opacity=1}]
+\newcommand\thmbox[1]{%
+  \noindent\begin{tikzpicture}%
+  \node [thmbox] (box){%
+\begin{minipage}{.94\textwidth}%    
+      \vspace{-3mm}#1\vspace{-3mm}%
+    \end{minipage}%
+  };%
+  \end{tikzpicture}}
+
+\let\theoremframecommand\thmbox
+\newshadedtheorem{lesson}{Research answer}
+
 \begin{document}
@@
 \begin{abstract} 
-The literature lists dozens of bad smells which might
+The literature lists many   bad smells that may
 be detrimental to a software system.
@@
 disagree on which bad smells are important. 
-Specifically, it is unclear at what threshold values
-a measurement should trigger code refactorings  since
-most studies do not offer
+Further, it is unclear at what threshold values smell-related metrics
+should trigger code improvement or preventative maintenance. Most studies do not offer
 {\em both} measurable metrics {\em and} a threshold for their proposed bad smells.
@@
 This paper proposes XTREE, a tool to prioritize  bad  smells.
-Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have little or no support in that log.
-
-When XTREE's recommendations are compared to 
-  those found by previous work (Shatnawi, Alves, and Borges.)
-  If code modules are  changed to avoid the bad
-  smells found by XTREE,  that changed system
-  has significantly fewer expected defects.
-  Also, XTREE   endorses
-  changes to far fewer code metrics.
-  Hence, it is 
-  a better tool for ruling out irrelvant bad smells.
-  For example, the XTREE results show that many of the bad smells learned 
-  by Shatnawi, Alves, and Borges can be ignored.
+Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have   no prior  association with defects.
+
+We evaluate XTREE's recommendations for bad smell improvement against recommendations from previous work (Shatnawi, Alves, and Borges) using multiple data sets of code metrics and defect counts.
+Code modules that are changed in response to XTREE's recommendations contain significantly fewer defects than recommendations in previous studies.
+  Further, XTREE endorses
+  changes to far fewer code metrics, and that bad smell recommendations (learned from previous studies) are not universal to software projects.
 
 Based on these results, we endorse using bad smells to  
-guiding refactoring. However, not all bad smells
-are   relevant to the current project.
-Hence, it is important
-to  prioritize the importance of specific bad smells within the context of each project.
+guiding refactoring. However,
+not all bad smells
+are relevant to the current project. It is hence important
+to  prioritize the importance of specific bad smells in the context of each project.
 
@@
 The literature is divided on the merits of bad smells.
-Much research  endorses using them as a guide
-code change. For example,
-the recent literature review of Tufano et al. at ASE'15~\cite{Tufano2015}  
-lists dozens of  
-papers proposing bad smell detection and repair tools.
-
-On the other hand, other papers cast doubts on the value of use bad smalls
-as triggers for code change.
+Much research endorses them to guide
+code improvement (e.g., refactoring or preventative maintenance). A recent literature review by Tufano et al.~\cite{Tufano2015}  
+lists dozens of papers on smell detection and repair tools.
+
+Other papers cast doubt on the value of bad smells
+as triggers for code improvement~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}. 
 As we show in this paper,  
- tools and text books and developers can disagree on what bad smells
+ tools, text books, and developers disagree on what bad smells
 are important enough to hunt down and eliminate. Our findings
-are consistent with other research suggesting that bad
-smells may not be a useful way to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.
-
-Debates on when/what code should be restructed are interesting, especially
-in the current context of  widely-used SE practices such as
-test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
-In that approach, developers are always asking themselves ``should I refactor this code''?
-Bad smells are a widely used  technique for deciding when to refactor\footnote{Evidence: As of  early 2016, ACM Digital Portal reports 
-the Fowler text has 6434 references. Also, bad smell-based research
-is widely reported. A search  on {\em "bad smell"+programming} 
-where {\em proceedings=ICSE} finds 1,748 papers since 2006.}.
+are consistent with other research suggesting that universal bad
+smells that ignore project context are not useful to guide refactoring~\cite{Mantyla2004,Yamashita2013,Sjoberg2013}.
+
+% Debates on when and which code should be restructured are interesting, especially
+% in the current context of  widely-used SE practices such as
+% test-driven agile development~\cite{beck2003test,janzen05,williams2003test,george2003initial}.
+% In that approach, developers are always asking themselves ``should I refactor this code''?
+Bad smells are a widely used technique for evaluating code. As of  early 2016, ACM Digital Library reports 
+the Fowler text has 6434 references. A search  on {\em ``bad smell''programming} 
+where {\em proceedings=ICSE} finds 1,748 papers since 2006.
 Hence it is important to understand the value of
- bad smells.  
-Valuable development resources can be wasted if developers
-refactor their code based on   bad smells that are ``bad''; i.e. 
+this popular topic.  
+Valuable development resources may be wasted on bad smells that are ``bad''; i.e. 
  irrelevant or misleading
@@
   proposed refactorings are be ignored
-since they are not justifiedable  
-with respect to the history of this project.
+since they are not justifiable  
+with respect to the history of the project.
 
@@
 bad smell X or Y is most important. 
- XTREE offers a way to assess
- competing claims about different refactorings.
+ XTREE offers a way to assess the value of competing recommendations to address smell X vs. Y.
 
@@
 Consider
-a bad smell that is triggering on some very long method,
+a bad smell stemming from a long method,
 yet that method must be passed as a single closure
@@
 
-To address this second
-  problem, we let
-XTREE {\em reason optimistically} to find the set
-$R_0$ of refactorings that might be useful
-(for details on how that optimistic reasoning
-is implemented, see \tion{rq}).
-Since that set was generated optimistically,
-the space of actually usable refactorings is 
-the possibly smaller subset $R_1 \subseteq R_0$.
-If the developer's proposed refactoring $R_2$ is
-not in $R_0$, then it cannot be in the smaller
-set $R_1$.  As shown
-in this paper's experiments,  XTREE's
-recommendations typically include only a small subset of
-the code metrics. That is, we
-can still rule out some proposed refactorings
+To address this problem, XTREE {\em reasons optimistically} to find the set
+$R_0$ of refactorings that \textit{might} be useful (see \tion{rq} on optimistic reasoning).
+The space of \textit{actually usable} refactorings is 
+the subset $R_1 \subseteq R_0$.
+If a developer's proposed refactoring $R_2$ is
+not in $R_0$, then it cannot be in the usable
+set $R_1$.  We show experimentally that XTREE
+often recommends small subsets of
+code metrics as useful guides for refactoring. Thus, we
+can rule out some proposed refactorings
 since many of those proposals fall outside of the space
-of refactorings suggested by XTREE.
+useful refactorings ($R_0$) suggested by XTREE.
 
@@
 Shantnawi~\cite{Shatnawi10} warns most  studies do not offer
-{\em both} a metrics and a threshold for their proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
-discuss what OO design metrics might defect bad smells (e.g. a excessive
-depth of inheritance tree) but they never comment on what thresholds
-means that some section of code starts to smell bad.  
-In all our  reading of the recent literature,
-the only work we can find from other researchers that
-offers metrics {\em and} thresholds come
-from Shatnami~\cite{Shatnawi10}
-and  Alves et al.~\cite{Alves2010} (for details on those
-methods, see \tion{methods}).
-One potential problem with their approaches is that 
-they do not explore the interconnections between sets
-of code metrics. This is important if sets of
+{\em both} a metric and a threshold for proposed bad smells. For example,  Darcy and Kemerer~\cite{darcy05}
+discuss object-oriented (OO) design metrics that might defect bad smells (e.g. an excessive
+depth of inheritance tree), but they do not define thresholds
+for when the code starts to smell bad.  
+Shatnami~\cite{Shatnawi10}
+and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{methods}).
+However, their approaches do not explore the interconnections between sets of code metrics. This is important if sets of
 code metrics have to be changed at the same time.
@@
 More generally, we say that bad smell thresholds should be
-{\em conjunctions} of metric+thresholds pairs that need to be changed {\em togehter}.
-Yet Shatnami~\cite{Shatnawi10}
-and Alves et al.   report   {\em disjunctions} of multiple metric/thresholds~\cite{Shatnawi10,Alves2010}
+{\em conjunctions} of metric+threshold pairs that need to be changed {\em together}.
+Shatnami~\cite{Shatnawi10}
+and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010}
 which could be separately and  inappropriately used
-to guide refactoring.
+to guide refactoring.\footnote{From LL: I don't understand the conjunction/disjunction.
+Are you saying Metric1+Metric2 needs its own threshhold, or you need \{Metric1+Threshold1, Metric2+Threshold2\}.
+And what do Satnami and Alves do?}
 
@@
 Refactored code cannot be assessed just by a rerun of the the test
-suite since
-code refactorings is meant to
-preserve the semantics of the code (i.e. the same tests
-should pass before and after the refactoring). 
-Also, it many not be practical
- to assessing refactorings by making the
-refactorings,  then seeing what happens to that code for the rest
-of the project. In a test-driven agile project, as
-SCRUM meetings change what stories are implemented next,
-the operational profile that lead to defects in the old
-code may not be repeated in future executions. 
+suite since (refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
+% waiting to observe 
+% code refactorings is meant to
+% preserve the behavior of the system (i.e. the same tests
+% should pass before and after the refactoring). 
+% Also, it many not be practical
+%  to assessing refactorings by making the
+% refactorings,  then seeing what happens to that code for the rest
+% of the project. In a test-driven agile project, as
+% SCRUM meetings change what stories are implemented next,
+% the operational profile that lead to defects in the old
+% code may not be repeated in future executions. 
 To resolve this  problem, SE researchers such as 
@@
     \item One system to propose changes;
-\item Another, completely separate, system to assess
+\item A separate system to assess
 how defective is the code before and after some
@@
  Cheng, O'Keefe, Moghadam and  Mkaouer et al. use the QMOOD hierarchical
-quality model~\cite{Bansiya02}, constructed in 2002.
-Our problem with using tools like QMOOD
-is that   quality models learned from other projects,
+quality model~\cite{Bansiya02}.
+A shortcoming of QMOOD
+is that quality models learned from other projects
 may perform poorly when applied to new projects~\cite{me12d}.
@@
 To solve this fourth problem, this study  eschews
-older quality models like QMOOD.
-Instead, we use
+older quality models like QMOOD, we use
 Random Forests~\cite{Breiman2001} to learn defect predictors
-that are specific to the systems   studied here.
-
- As to XTREE's limitations, the tool 
-  can only comment
-on refactorings expressed in the   terminology
-as the data
-  in the historical log of a project.
-If developers want to justify their refactorings
-via bad smells expressed in other terminology,
-then a XTREE-style analysis must   wait till 
-data about those new
-terms has been collected.
-
-Another  aspect to note about XTREE  is that the tool assumes
-Fowler's bad smells
-should be assessed only via their contribution to software defects
-(evidence: see the section in italics of the quote
-that starts this introduction). Some researchers 
- comment that   debating bad smells is  useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
-debates about refactorings can   (a)~introducing new comers to the code base or (b)~share knowledge around a team
+from OO code metrics.
+Note that, unlilke QMOOD, such measurements 
+  are specific to the project, and the measurements can be reconstructed and the predictors rebuilt for other projects.
+
+The rest of this paper is structured as follows. XXX.
+discusses the cognitive
+biases that make developers disagree on the relevance of those bad smells to
+their particular projects.  
+ That discussion is our motivation for developing the XTREE  tool
+that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.
+
+XTREE   is focused on Fowler's bad smells
+as contributors to software defects. Some researchers 
+ suggest that discussing bad smells is useful for more just reducing defects. For example, Bosu and Carver~\cite{bosu13} report that
+debates about code smells and refactorings can (a)~introduce new comers to the code base or (b)~share knowledge on a team
 about interactions within the code. In that view, 
@@
 that the new code will contain defects.
-
-The rest of this paper is structured as follows. XXX.
-discusses the cognitive
-biases which make developers disagree on the relevance of those bad smells to
-their particular projects.  
- That discussion is our motivation for developing the XTREE  tool
-that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.
 
@@
 
-\newpage
 \subsection{Assessment Oracles}
@@
 
-The Scott-Knott  results are presented in the form of line diagrams like those shown on the right-hand-side of the following example table.
+The Scott-Knott  results are presented in the form of line diagrams like those shown on the right-hand-side of \fig{jur}
 The black dot shows the median \eq{diff} values and the horizontal likes stretch 
@@
 range, or IQR).
-\begin{center}
-{\small
-  Example 
-\begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
-\arrayrulecolor{lightgray}
-\rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
-1 &      XTREE &    62  &  6 & \quart{67.2}{7}{69.6}{115}  \\
-\hline 2   & Alves        &    40  & 20  & \quart{45}{10}{50}{10}\\
-\hline 3 &      CD &    24  &  18 & \quart{28}{14.4}{35.2}{115} \\
-\hline 4   & Shatnawi        &    15  & 10  & \quart{28}{5}{30}{10}\\\hline
-\end{tabular}}
-\end{center}
-In this example table, the rows are  sorted on the median values of each method.
+As an example of how to read this table, consider the {\em Ant}
+results.
+Those rows are  sorted on the median values of each method.
 Note that all the methods have \eq{diff} $\gt 0\%$; i.e.
-all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 62\% from the original value).
-The above example table has a  left-hand-side  {\bf Rank} column, computed using the
-Scott-Knott test described above. In this example table, XTREE is ranked the best, while Shatnami is  ranked   worst.
+all these methods reduced the expected value of the performance score in that experiment while XTREE achieved the greatest reduction (of 56\% from the original value).
+These results table has a  left-hand-side  {\bf Rank} column, computed using the
+Scott-Knott test described above. In the {\em Ant}
+results, XTREE is ranked the best, while CD is  ranked   worst.
+
+% \begin{center}
+% {\small
+%   Example 
+% \begin{tabular}{{l@{~~~}l@{~~~}r@{~~~~}r@{~~~}c@{}r}} 
+% \arrayrulecolor{lightgray}
+% \rowcolor{lightgray}\textbf{Rank} & \textbf{Treatment} & \textbf{Median} & \textbf{IQR} & \\
+% 1 &      XTREE &    62  &  6 & \quart{67.2}{7}{69.6}{115}  \\
+% \hline 2   & Alves        &    40  & 20  & \quart{45}{10}{50}{10}\\
+% \hline 3 &      CD &    24  &  18 & \quart{28}{14.4}{35.2}{115} \\
+% \hline 4   & Shatnawi        &    15  & 10  & \quart{28}{5}{30}{10}\\\hline
+% \end{tabular}}
+% \end{center}
+
+\input{tex/results_jur.tex}
+
+\input{tex/rq2.tex}
 
@@
 
-\input{tex/results_jur.tex}
-
-\fig{jur} shows the comparison results.  The Alves and Shatnami
-entries show net effects of applying (seperately) all the disjuncts
-proposed by those methods.
-
-Three data sets are very responsive to defect reduction suggestions:
-Jedit, Ant, and Ivy (all of which show best case improvements over 50\%.
- The Jedit results are  
-surprisingly uniform  suggesting that in that data set that
-our methods all find the same ways to reduce the expected number of
-defects.
-
-Two data sets are not very responsive to defect reduction:
-Poi and Lucene. The reason for this can be see in \fig{j}:
-both these data sets contain more than 50\% defective modules.
-In that space, all our  methods lack a large sample to
-defect-free examples.
-
-More important that the size of the improvement is the relative
-rank of the different approaches,
- CD and Shatnawi usually  perform comparatively worse while  XTREE gets top ranked position the most
-number of times. That said, Alves sometimes beats XTREE (see Ivy)
-while sometimes it ties (see Jedit).
-
-\subsection{RQ2}
-\input{tex/rq2.tex}
-
-\subsection{RQ3}
-\begin{figure*} 
+\begin{figure*}[t!]
 \centering
-\includegraphics[width=\linewidth]{figs/changes0.png}
+\includegraphics[width=\linewidth]{figs/changes01.png}
 \caption{XTREE results.
@@
 
+\subsubsection{RQ1: Effectiveness}
+
+{\em In a comparison of   outlier and cluster delta
+methods, which one most reduces the expected number
+of defects?}
+
+\fig{jur} shows the comparison results.  The Alves and Shatnami
+entries show the  net effects of applying (seperately) all the disjuncts
+proposed by those methods.
+
+Two data sets are very responsive to defect reduction suggestions:
+ Ant, and Ivy (both of which show best case improvements over 50\%).
+ The  expected value of defects  
+ is changed less in Jedit. This data sets' results
+ surprisingly uniform; i.e.   all methods
+ find the same ways to reduce the expected number of
+defects.
+
+Two data sets are not very responsive to defect reduction:
+Poi and Lucene. The reason for this can be see in \fig{j}:
+both these data sets contain more than 50\% defective modules.
+In that space, all our  methods lack a large sample to
+defect-free examples.
+
+More important that the size of the improvement is the relative
+rank of the different approaches,
+ CD and Shatnawi usually  perform comparatively worse while  XTREE gets top ranked position the most
+number of times. That said, Alves sometimes beats XTREE (see Ivy)
+while sometimes it ties (see Jedit).
+
+\begin{lesson}
+In terms of the expected number of defects
+  after refactoring,
+XTREE is marginally better than Alves while CD and Shatnawi
+are deprecated.
+\end{lesson}
+
+ \subsubsection{RQ2: Verbosity}
+
+{\em In a comparison of outlier and cluster delta methods, which one proposes the fewest thresholds?}
+
+\fig{counts} shows the frequency with which our different methods
+require changes to code methods.
+Note that XTREE proposes thresholds to
+few code metrics than the other approaches.
+
+\begin{lesson}
+In  terms of how often   a bad
+smell can be ignored, XTREE is best since it is  least verbose (i.e. least
+likely to  overlap   a bad smell mentioned by a developer).
+\end{lesson}
+
+When combining   \fig{jur} with \fig{counts}, we   see that
+even though XTREE proposes changes to far fewer code metrics, those few
+code metrics are usually just as effective (or
+more effective) than the multiple
+thresholds
+proposed by CD, Shatnawi or Alves.  That is, XTREE proposes
+{\em fewer} and better thresholds than the other approaches studied here.
+
+\subsubsection{RQ3: Co-Change}
+
+{\em Does XTREE understand what metrics get changed together?}
+xxxx
+
 % \input{tex/results_shatnawi.tex}
@@
 \section{Threats to Validity}\label{sect:valid}
-
-XXXX only reorg due to the issues in the data sets.
-we are stuck with that. note hwever that if
-projects improved their data collection then
-our techniques woudl still apply, unmodifed, to the revised data sets. so here we are ding what we can with the data currently available (and also ensuring we are good to go on future data)
 
@@
 mind.
+
+\subsection{External Validity}
+Based on the experimental results above,
+as well as the discussion in \tion{prelim},
+we believe that specific bad smell indicators (e.g. \mbox{{\em loc}$>$ 100})
+have very limited external validity. 
+The good news is that,
+while specific models are not externally valid,
+there may still be general methods like XTREE for finding the good local models.
+
+One   limitation to certifying bad smells
+is the terminology used to express the bad smell.
+XTREE, Shatnawi, Alves et al. can  only comment
+on refactorings expressed in the   terminology
+as the data
+  in the historical log of a project.
+If developers want to justify their refactorings
+via bad smells expressed in other terminology,
+then the  analysis of this paper must   wait till 
+data about those new
+terms has been collected.
+
+XTREE does not require a working defect predictor to generate
+its bad smell thresholds. However, to {\em certify}
+that those thresholds are satisfactory, some
+second quality predictor must be applied. Historically,
+SE researchers~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} have used the QMOOD hierarchical quality model~\cite{Bansiya02}.
+Since such quality models may be very project specific~\cite{localvsglobal}, we prefer building local 
+defect predictors using Random Forests. 
+Our results show that when a defect log contains too few
+non-defective examples then defector predictors
+perform poorly (see the data sets in the last few lines of \fig{j})
+as do their associated bad smell threshold generators
+(recall the performance of Lucene and Poi in~\fig{jur}).
+
+It should be noted that this is not a limitation of XTREE-- rather
+it seems fundamental to the problem of certifying bad smells.
+As evidence of this, recall that {\em all} the bad smell threshold generators of~\fig{jur}
+had issues with the defect data sets with  many defects
+(Lucene and Po).
 
@@
 \subsection{ Learner Bias}
-For building the defect predictors in this study, we elected
-to use  Random Forests.
-We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.
+This study used two kinds of learners: (1) methods for suggesting
+changes to code like Shatnawi, Alves et al., CD and XTREE;
+(2) methods for assessing the changes proposed by these learners.
+Prior studies implemented (2) using  the QMOOD hierarchical quality model
+which is so old, that we suspect it may no longer be current. Hence,
+we used Random Forests to build the assessment oracles that were specific
+to the data sets used in this study. .We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.
+
 Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.  
-
 That said, we have taken care to document in this paper the decisions made by engineering
-judgement that could effect our conclusions. The above code used a set of variables which future
-work should vary in order to test the internal validity of our conclusions:
-
-\bi
-\item The clustering scheme, WHERE, divides data into groups of size $\alpha=\sqrt{N}$;
-\item XTREE considered a sibling useful if it's score was less than $\gamma=0.5$ times the mean score of the current leaf.
- \ei
+judgement that could effect our conclusions.
 
@@
 Sampling bias threatens any data mining experiment; i.e., what matters
-there may not be true here. For example, the data sets used here comes from two sources
-(Seigmund et al. and Jureczko et al.) and any biases in their selection procedures
+there may not be true here. For example, the data sets used here comes from two  Jureczko et al. and any biases in their selection procedures
 threaten the validity of these results. 
@@
 \subsection{  Evaluation Bias}\label{sect:coc}
-Another threat to validity of this work is our use
-of predictors learned on the training data to assess the impact of our planners.
-This issue was discussed in detail in \tion{trust}.
-
-To those notes, we add a few more details. If possible, planners should be assessed via some external oracle that can accurately assess new examples. For example, in search-based software engineering,
-examples can be assigned objective scores via  some model. In this approach, a changed example can be assessed by
-generating actual objective scores from the model.
+
+The main evaluation bias in this study was the use of
+optimistic reasoning when generating the changed code modules.
+We assumed in \tion{rq}
+that any code module can be changed in order
+to stop triggering a bad smell detector. This is not necessarily
+true, as noted in {\em PROBLEM \#2} in this paper's introduction.
+To mitigate this issue, we do not propose using XTREE as a bad
+smell generator, but as a critique agent for bad smells
+proposed by some other oracle (e.g. a human developer).
+We showed about that XTREE's recommendations are not verbose
+so it will not often overlap with some random bad smell proposed
+by a developer.
 
@@
 
-The POM3 model~\cite{boehm2003using},~\cite{port2008} is a tool for exploring management challenges.
-POM3 implements the Boehm and Turner model of agile programming~\cite{boehm2003balancing} where teams select tasks as they appear in the scrum backlog.
-POM3 can study the implications of different ways to adjust task lists in the face of shifting priorities.
-The model outputs estimated task completion rates; programmer idle rates; and total overall cost.
-POM3 models requirements as a dependency tree.
-A single requirement in the tree of a prioritization value has a cost, along with a list of child-requirements and dependencies.
-Before any requirement can be satisfied, its children and dependencies must first be satisfied.
-POM3 simulates changing priorities by making teams aware of random items in the requirements tree at random intervals, thus forcing teams to constantly readjust their "to do" lists.
-For further details on this model see~\cite{boehm2003using},~\cite{port2008}, and~\cite{boehm2003balancing}.
-
-\fig{coc} shows results from using POM3 as an oracle to assess our planning methods and their ability to reduce the project cost.
-In this experiment, we generated a training and testing set with 1000 randomly generated instances, which we passed to our four methods.
-40 times, we let those methods propose changes to those projects. 
-For assessment purposes, the changed projects were then fed back to the POM3
-oracle.
-
-Using this approach, it is possible to assess the value of a plan by measuring its
-effectiveness with respect to some ground truth (in this case, the POM3 model).
-As shown in \fig{coc}, XTREE reduces the cost by 59\% thereby passing this assessment.
-These results from the POM3 model further endorse our previous conclusions; i.e.
-compared to three other methods,  XTREE's supervised methods are best for generating plans on how to change example projects.
-
 \section{Conclusions}
-
-The planner proposed in this paper proposes changes to software project details in order to improve the expected
-value of the performance scores of that part of the project.
-To evaluate these planners,
-data miners can be used to build oracles to assess planners.
-Such planners should be {\em trust-increasing}; i.e. they should propose changes that generate
-changed examples that are closer to the training data of the data miner.
-One caveat here is that the evaluations we can make on the planner are only as good as the predictive
-performance of the data miner. Hence, if domain data does not support satisfactory predictors, then
-planning in that domain cannot be evaluated.
-
-Four planners were assessed here for the tasks of reducing defects and runtimes. 
-Three of the methods come from our prior publications~\cite{me12c,krishna15}, and the conclusion of this
-paper is that a novel fourth method clearly out-performs the other three
-(measured in terms of {\em effectiveness, succinctness, and surprise}).
-We conjecture that XTREE worked better than the rest since:
-\bi
-\item
-It uses a supervised method to divide the data and;
-\item
-When planning how to move examples to better classes, it is best to reflect over 
-differences between those classes.
-\ei
-\section{Future Work}
-Future work in this area could explore numerous questions. For example, 
-XTREE, as used here, sought improvements in a one goal (the class variable). Does XTREE work for multi-goal reasoning?
-
-Also, the XTREE algorithm seems quite general to any data of table with rows containing
-weighted classes (so we can distinguish ``bad'' rows from ``better'' ones). 
-Does   XTREE works  on   domains (other than the defect/runtime data explored here)?
-
-As an example of a domain that might benefit from XTREE, recent results raise doubts about
-the value of changing code to remove ``bad smells''~\cite{Sjoberg13}. Can XTREE be used as a ``bad smell'' detector to select the subset of possible refactorings that have the most potential benefit?
-
-As to scalability, XTREE is a post-processor to a decision tree algorithm.
-Hence, in theory, XTREE   works on domain where data miners can generate decision trees.
-Given the current state of the art in Big Data, can XTREE  be applied to  very large data sets?
-
-We discussed above in \tion{p2p} the general  conclusions of Passos, J{\o}rgensen  et.al~\cite{passos11,jorgensen09};
-i.e. software developers are reluctant to surrender their old biases
-in the face of new data. 
-Accordingly, it must be asked if  the {\em mental resistance} of developers
-will prevent them applying XTREE's automatically generated recommendations of tools?
-Note this this issue is not just a concern for XTREE, but also for any 
-automatic tool proposing refactorings.
-
-There are many more methods for generating plans and
-no   one paper can survey them all. For example, this paper has not explored variations to the $\alpha,\beta$, and $\gamma$
-parameters that controlled XTREE. Would we get better results if we varied those parameters?
-
-That said, the goal of this paper was not to claim that (e.g.) XTREE is some absolute optimal algorithm. Rather, it is
-was to offer a baseline result (with XTREE) and an  evaluation strategy that  can assess  if alternate methods are better than XTREE.
-The authors of this paper would actively support other teams exploring this
-method (with or without using our current code base). So our last questions is this:
-will other  researchers try to repeat and/or improve
-(or even refute) our results?  Perhaps.
+When can bad smells be ignored? Our answer is as follows.
+When those
+bad smells are not supported by the historical log of data from
+the current project.
+
+When that data is not available (e.g. early
+in the project) then developers could use the general list of
+bad smells shown in \fig{smells}. However, 
+our results
+show that  bad smell detectors are most
+effective when they are based
+on a small handful of code metrics (as done by XTREE).
+Hence, using all the bad smells of \fig{smells} may not be optimal.
+
+XTREE improves on prior methods for generating bad smells:
+\begin{itemize}
+    \item As described in \tion{prelim}, bad smells generated by humans may not be apply to the current project. On the other hand, XTREE can learn specific thresholds for the current project.
+    \item As described in in \tion{}, XTREE generates conjunctions
+    \item Prior methods used an old quality predictor (QMOOD) which we replace with defect predictors learned from the current project.
+\end{itemize}
 

### 2016-03-12 Manuscript.tex
final save for the fse submission

@@
 \definecolor{shadecolor}{gray}{0.9}
-\conferenceinfo{ISCE}{'16 Austin, Texas}
-
-\title{  Which ``Bad Smells'' can be Ignored?}
+\conferenceinfo{FSE}{'16 Seattle, Washington}
+
+\title{  Which ``Bad Smells'' Can Be Ignored?}
 \numberofauthors{2} 
@@
 This paper proposes XTREE, a tool to prioritize  bad  smells.
-Given an historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have   no prior  association with defects.
+Given a historical log of defects seen previously in the code, XTREE deprioritizes bad smells that have   no prior  association with defects.
 
@@
 
-Based on these results, we endorse using bad smells to  
-guiding refactoring. However,
-not all bad smells
-are relevant to the current project. It is hence important
-to  prioritize the importance of specific bad smells in the context of each project.
+It is hence important to  prioritize the importance of specific bad smells in the context of each project since not all bad smells are relevant to the current project.
+Based on these results, we endorse using bad smells to guiding refactoring.
 
@@
      \begin{itemize}
-         \item An historical log of defects seen previously in the code;
+         \item A historical log of defects seen previously in the code;
          \item  Bad smells expressed in terms of the 
@@
 Given that information,  XTREE can determine which
-  proposed refactorings are be ignored
+  proposed refactorings are to be ignored
 since they are not justifiable  
@@
 for when the code starts to smell bad.  
-Shatnami~\cite{Shatnawi10}
-and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{methods}).
+Shatnawi~\cite{Shatnawi10}
+and  Alves et al.~\cite{Alves2010} are rare cases where both metrics and thresholds are defined (see \tion{bst}).
 However, their approaches do not explore the interconnections between sets of code metrics. This is important if sets of
@@
 For example, it is not possible to refactor a
-long method it several smaller ones without increasing
+long method if several smaller ones without increasing
 the coupling metrics of a class. 
@@
 {\em conjunctions} of metric+threshold pairs that need to be changed {\em together}.
-Shatnami~\cite{Shatnawi10}
-and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010}
-which could be separately and  inappropriately used
-to guide refactoring.\footnote{From LL: I don't understand the conjunction/disjunction.
-Are you saying Metric1+Metric2 needs its own threshhold, or you need \{Metric1+Threshold1, Metric2+Threshold2\}.
-And what do Satnami and Alves do?}
+Shatnawi~\cite{Shatnawi10}
+and Alves et al. report {\em disjunctions} of multiple metrics+thresholds~\cite{Shatnawi10,Alves2010} without indicating how changes to these metrics are related.
 
@@
 is that quality models learned from other projects
-may perform poorly when applied to new projects~\cite{me12d}.
+may perform poorly when applied to new projects~\cite{localvsglobal}.
 
@@
 
-The rest of this paper is structured as follows. XXX.
+The rest of this paper is structured as follows. 
+The next section
 discusses the cognitive
 biases that make developers disagree on the relevance of those bad smells to
-their particular projects.  
- That discussion is our motivation for developing the XTREE  tool
-that   carefully assesses the merit of old bad smells to new projects.  XTREE is presented in Section 3 and evalauted in the rest of the paper.
-
+their particular projects
+(that discussion is our motivation for developing 
+tools like XTREE). This is followed by a discussion
+on various methods to learn thresholds for bad smells.
+After evaluating those methods, we discuss various threats to
+validity that threaten our conclusions.
+
+Before  that, we offer one following note on a
+core assumption of XTREE.
 XTREE   is focused on Fowler's bad smells
@@
 even if it does not address issues of defects. 
-That said, using
-defect predictors,
-it is possible to operationalize XTREE
-while it might be harder to otherwise operationalize the assessment of bad smells.
-Hence, 
-this paper assumes that {\em better} bad smells and
-bettering at finding code changes that  {\em reducing} the chances
-that the new code will contain defects.
+% That said, using
+% defect predictors,
+% it is possible to operationalize XTREE
+% while it might be harder to otherwise operationalize the assessment of bad smells.
+% Hence, 
+% this paper assumes that {\em better} bad smells and
+% bettering at finding code changes that  {\em reducing} the chances
+% that the new code will contain defects.
 
@@
 previously. 
- The experimental rig
- used to comparatively assess
-    XTREE is another novel contribution. Using the design
-    of that rig,
-    other researchers could confirm/improve/refute
-    the conclusions of this paper.
-    Another contribution here is
-    the comparative analysis of XTREE
+The comparative analysis of XTREE
     versus previously published approaches to learning
-    bad smell detectors. that comparion shows
+    bad smell detectors shows
     that XTREE   rarely perfoms
@@
       changes that reduce the expected
-      value of the defects. Further, the number
+      number of of the defects. Further, the number
     of code metrics mentioned by XTREE is far
-    fewer than other approaches This means that XTREE is a better
-    tool for ruling out   refactorings proposed by
+    fewer than other approaches, making it easier to validate refactorings proposed by
     developers. 
-
-As to the relation to prior work, we
-show here that XTREE is a  superior 
-to the CD tool   developed previously with Borges~\cite{me12c}.
+We
+show here that XTREE is less verbose and offers better recommendations than the CD tool   developed previously with Borges~\cite{me12c}.
+ The experimental rig
+ used to comparatively assess
+    XTREE is another novel contribution\footnote{tiny.cc/XTREE}. Using the design
+    of that rig,
+    other researchers could confirm/improve/refute
+    the conclusions of this paper.
 
@@
 Why build tools like XTREE to  critique proposed developer actions?
-Are not those developers the best guide to what is important, or irrelevant
-to their code?
-
-It turns our that developer {\em cognitive biases} can mislead them into
-asserting that some things are important and relevant when they are not.
+Are not those developers the best guide to what is important or irrelevant
+to their code? Developer {\em cognitive biases} can mislead them to
+assert that some things are important and relevant when they are not.
 This section discusses those biases-- firstly in the general sense
@@
   inaccurate and 
-misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may hence have very
-strong views on many issues, including bad smells, but those views may be wrong for the current
+misleading~\cite{passos11,jorgensen09,mei15,me16phase,prem16}. Software developers may have strong views on many issues, including bad smells, but those views may be wrong for the current
 project.
-It should be noted that  software engineering is not the only field with this problem.
+Software engineering is not the only field with this problem.
   For example, the medical profession applies many practices based on studies that have been disproved (a recent article in the Mayo Clinic Proceedings~\cite{prasad13} found 146 medical practices based on studies in year i, but which were reversed by subsequent trials within years i + 10).
@@
 They report that the supposed software engineering ``gurus'' rarely use lessons from past projects to improve their future reasoning and that such poor past advice can be detrimental to new projects.~\cite{jorgensen09}.
-
-To support the last paragraph, we offer several
-specific studies
-showing that   widely-held views in SE that are   now questionable, given recent results. 
-The first few come from the general SE literature while the rest
-relate specifically to bad smells:
+Other studies have shown some widely-held views are   now questionable given new evidence:
 
 \begin{enumerate} 
-\item
-Nagappan et al. analyzed~\cite{mei15} 
-how   {\tt goto}  was used  in 11,000 Github repositories. 
-They found that developers
-rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
-memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
+% COMMENT FROM LL: This bullet is weak. Maybe goto isn't used in an unrestricted manner because of Djikstra's warning.
+% \item
+% Nagappan et al. analyzed~\cite{mei15} 
+% how   {\tt goto}  was used  in 11,000 Github repositories. 
+% They found that developers
+% rarely used {\tt goto} in the unrestricted manner that alarmed Dijkstra in his  1968
+% memo ``Go to statement considered harmful''~\cite{Dijkstra68}.
 \item
 In other work~\cite{me16phase} with the Software Engineering Institute (SEI), we have revisited
-the ``phase delay'' truism that {\em fixing a requirements error  during
-coding can be thousands of times more expensive than recognizing and fixing that error
-as soon as it arises}. 
+the ``phase delay'' truism that {\em the cost of fixing an error dramatically increases the longer it is in the system}. 
 We found no evidence for such large   phase delay effect in modern
 software (in 171 software projects shepherded
-by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ software development languages, tools and development practices. 
+by the SEI, 2006-2014) possibly since it has been  heavily mitigated by $21^{st}$ century software development languages, tools and development practices. 
 Whatever the reason, the main point  is that phase delay is a widely
@@
 Devanbu et al.  examined responses from 564 Microsoft software developers from around
-the world, They found that  ``(a)~programmers do indeed have very
+the world, they found that  ``(a)~programmers do indeed have very
 strong beliefs on certain topics; (b)~their beliefs are primarily formed
@@
 as the strongest influence in forming their opinions''. This is a troubling
-result, especially given the above comments from Passos,  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, then freeze, those opinions for the rest of their
-careers.
+result, especially given the above comments from Passos and  J{\o}rgensen et al.~\cite{passos11,jorgensen09} about how quickly practitioners form, freeze, and rarely revist those opinions.
 
@@
 If the above remarks hold true for bad smells, then we would expect
-to see much disagreement on what bad smells are important and relevant
+to see much disagreement on which bad smells are important and relevant
 to  a particular project. This is indeed the case.
 The first column of \fig{smells} 
-lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{Fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
+lists  commonly mentioned bad smells and comes from Fowler's 1999 text~\cite{fowler99} and a subsequent 2005 text by Kerievsky that is widely cited~\cite{Kerievsky2005}.
 The other
-columns show other work that comments on which bad smells matter most.
-The columns marked as Lanza'06, Yamashita'13, and Atwood'06 come from
-peer reviewed literature. The column maked SonarQube is an open source
-code assessment tool that includes dectectors for six of the bad smells
+columns show data from other studies on which bad smells matter most.
+The columns marked as Lanza'06 and Yamashita'13 are from peer reviewed literature. The column maked SonarQube is a popular open source
+code assessment tool that includes detectors for six of the bad smells
 in column one. 
 The {\em developer survey}   (in the right-hand-side column),
-   this shows the results of a hour-long whiteboard sessions with a group of a
- dozen developers from one Washington
-    D.C. web-tools development organization. In that study, participants
+   this shows the results of an hour-long whiteboard sessions with a group of 12 developers from a Washington
+    D.C. web tools development company. Participants
     worked in a round robin manner to rank the bad smells they thought were
     important (and any disagreements were discussed with the whole group).
-     Amongst those group, there was there was  some
-    consensus on  the priority of fixing which  bad smells   
+     Amongst the group, there was  some
+    consensus on  the priority of which bad smells to fix
     (see the annotations VH=very high,
@@
 one the   bad smells in column one. 
-Note that most of the cells are blank.
-While the Atwood'06~\cite{Atwood06} study  comments on   
-all the   bad smells,
-all the  other  studies ignore the majority of the Fowler bad smells.
-For exampel SonarQube has no detectors for many of the column one bad smells.
+Note that most of the cells are blank, and that the studies ignore the majority of the Fowler bad smells.
+SonarQube has no detectors for many of the column one bad smells.
 Also, nearly half the Yamashita list of bad smells
-    does not appear in the Fowler and Kerievsky list
-    (the eight numbers
+    does not appear in the Fowler and Kerievsky list. The eight numbers
     in the  Yamashita'13 column show the rankings for the bad smells 
-    that overlap with Fowler and Kerievsky).
+    that overlap with Fowler and Kerievsky; Yamashita also discussed other smells not covered in Fowler'99.
 
 Two of the studies in \fig{smells} offers some comments on the relative importance
-of the different bad smells the   Yamashita'13 study and the developer study.Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey.
+of the different bad smells in the Yamashita'13 study and the developer study.
+Three of the bad smells listed in the top half of the Yamashita'13 rankings also score very high in the developer survey.
 Those three were {\em duplicated code, large class},
@@
  Note that this agreement also means that the
-  Yamashita'13 study and the {\em developer survey}   
+  Yamashita'13 study and the developer survey   
   believe
@@
 
-In summary, just because a developer strongly believes something
-about bad smells, does
-not necessary mean that that belief is useful for the current project.
-Human developers can be clever, but their thinking can also be distorted
+In summary, just because one developer strongly believes in the importance of a bad smells does not mean that belief transfers to other developers or projects.
+Developers can be clever, but their thinking can also be distorted
 by cognitive biases.
-Hence, as shown in \fig{smells}, developers and text books and tools 
+Hence, as shown in \fig{smells}, developers, text books, and tools 
 can disagree on which bad smells are important.
-Special tools are needed to assess their beliefs in, for example,
+Special tools are needed to assess their beliefs, for example, their beliefs in
 bad smells.
 
-\section{Learning Bad Smell Thresholds}
+\section{Learning Bad Smell Thresholds}\label{sect:bst}
+Research studies and tools (e.g., SonarQube~\cite{sq15}) leverage OO metrics as representatives for code smells, as we do in this study.
+In this section, we focus on determining thresholds for interpreting these metrics as ``bad smells'' or not.
+
 The SE literature offers  two ways to 
 learning bad smell thresholds.
-One way relies on 
-{\em outlier statistics}~\cite{erni96,bender99}. This way
+One approach relies on 
+{\em outlier statistics}~\cite{erni96,bender99}. This approach
 has been used   by Shatnawi~\cite{Shatnawi10}, Alvenes et al.~\cite{Alves2010}
@@
     use here for XTREE. 
-
     These two approaches are discussed below.
-\subsection{  Outliers}\label{sect:p}
-
-The outlier approach assume that unusually large measurements are risk-prone.
+\subsection{  Outlier Statistics}\label{sect:p}
+
+The outlier approach assume that unusually large measurements indicate risk-prone code.
 Hence, they generate one bad smell threshold for any metric
@@
 \subsubsection{Enri \& Lewerentz}
-One of the first methods was proposed Erni and Lewerentz~\cite{erni96}. 
 Given classes described with the  code metrics of \fig{ck},
-Enri and Lewerentz found the   mean $\mu$ and the standard deviation $\sigma$
+Enri and Lewerentzz~\cite{erni96} found the   mean $\mu$ and the standard deviation $\sigma$
 of each
@@
 metric with a measurement greater than $\mu+\sigma$.
-
-Shatnami and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
-using $\mu+\sigma$. since it does not   consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.
-
-\subsubsection{ Shatnami}
-Shatnami~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
+Shatnawi and Alves et al.~\cite{Shatnawi10,Alves2010} depreciate
+using $\mu+\sigma$ since it does not consider the fault-proneness of classes when the thresholds are computed. Also, the method lacks  empirical validation.
+
+\subsubsection{ Shatnawi}
+Shatnawi~\cite{Shatnawi10}'s preferred alternative to $\mu+\sigma$
 is to use the VARL method (Value of Acceptable Risk Level) proposed by Bender~\cite{bender99} in his epidemiology studies.  This approach uses two
-thresholds which, following Shatnami's guidance, we set to
+thresholds which, following Shatnawi's guidance, we set to
 $p_0=p_1=0.05$.
@@
 VARL encodes the defect count
-for each class to 0 (for no defects known) to 1 (for defects known).
+for each class as 0 (no defects known in class) or 1 (defects known in class).
 Univariate binary logistic regression is applied to learn three coefficients:  
- $\alpha$ is  constat intercept;
+ $\alpha$ is the intercept constant;
     $\beta$ is the coefficient for maximizing log-likelihood;
@@
 \end{equation}
+For consistency
+with Shatnawi, we explore the correlation between
+the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
 \subsubsection{ Alves et al.}
@@
 to finding thresholds that  utilizes the underlying statistical distribution and scale of the metrics. 
-Evey metric value is weighted according to the source lines of code (LOC) of the class.
-All the weighted metrics are then normalized; i.e.
-they are divided by the sum of all weights of the same system.
-For consistency
-with Shatnawi, we explore the correlation between
-the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
-Following this, the normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
-Alves et al. then compute their 
-thresholds   by choosing the percentage of the overall code that needs to be represented.
-For instance, Alves et al.
-suggest the use 90\% quantile of the overall code to derive the threshold for a specific metric.
-The intuition here is that the worst
-code have outliers beyond 90\% of the normal code measurements.
-
-Our reading of the literature is that the Alves et al.  
-approach is preferred to that of Shatnami. For example, Herman et al.~\cite{hermans15} used the
+Metric values for each class are weighted according to the source lines of code (SLOC) of the class. All the weighted metrics are then normalized by the sum of all weights of for the system. 
+The normalized metric values are ordered in an ascending fashion (this is equivalent to computing a density function, in which the x-axis represents the weight ratio (0-100\%), and the y-axis the metric scale).
+Alves et al.
+then select a percentage value (they suggest 90\%) which represents the ``normal'' values for metrics.
+The metric threshold, then, is the metric value for which 90\% of the classes fall below.
+The intuition  is that the worst code has outliers beyond 90\% of the normal code measurements.
+Hermans et al.~\cite{hermans15} used the
 Alves et al. method in their  2015 paper on
@@
 \subsubsection{Discussion }\label{sect:disc}
-The advantage of the outlier
-approach is that it is simple to implement. 
-For example, in less than a day we could
-code up and apply  the Shatnami and 
-Alves et al. methods.
-
-The outlier approaches has   two  major disadvantages. 
-Firstly, they are {\em verbose}; i.e.  they can
-propose thresholds on most of the code metrics.
-Such verbosity makes it hard   to critique and prune bad smells
-since any bad smell mentioned by a developer has some support
-somewhere amongst all those proposed thresholds.
-
-Secondly, the outlier approaches suffers from
-the {\em disjunction} problem discussed in the introduction.
+The advantage of the outlier-based
+approaches is that they are simple to implement, but the approaches have   two  major disadvantages. 
+First, they are {\em verbose}.
+A threshold can be calculated for every metric -- so, which one should the developers focus on changing?
+Without a means for prioritizing the  thresholds and metrics against one another, developers may have numerous or conflicting recommendations on what to improve.
+Second, the outlier approaches suffers from
+the {\em disjunction} problem.
 That is, while
@@
 modules than $C_-$;
-\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$};
+\item Compute the  delta   in code metrics between the clusters using \mbox{$\Delta = C_+ - C_-$}, i.e.
+{\em towards} the cluster with lower defects;
 \item $\Delta$ are changes needed in defective modules of $C_+$ to
@@
 Since it is computed
-from neighboring clusters, it   reasons about
-examples with similar distributions.  Hence,   $\Delta$  respects the naturally occurring constraints in the data. For example,
-repeating the above example about large methods,   $\Delta$   will not  suggest lowering lines of code
-without also increasing a coupling measure
-
+from neighboring clusters, the examples contain similar distributions and $\Delta$ respects the naturally occurring constraints in the data. For example,
+given a bad smell pertaining to large methods,   $\Delta$   will not  suggest lowering lines of code
+without also increasing a coupling measure. 
 Cluster deltas are used in CD~\cite{me12c} and XTREE.
@@
 
-As shown below, one drawback with with CD is that it is very {\em verbose}
+As shown below, one drawback with with CD is that it is {\em verbose}
 since
@@
 metrics with different values in those two centroids. 
-This makes it hard to use CD to   critique and prune away bad smells.
-
-Another drawback with CD is revealed by the comparison experiments
-in this paper: CD never worked as well as XTREE in proposing
-changes to reduce defects.
+This makes it hard to use CD to   critique and prune away bad smells. Further, CD will be shown to be
+not as effective
+in proposing changes to reduce defects as XTREE.
 
@@
   XTREE utilizes a decision tree learning approach
-  to find the fewest differences groupings of examples.
-
-  Decision trees are built via
-  {\em iterative dichotomization} that  finds  a  split  in  the  values  of  independent  features  that  most  reduces the variability of the dependent feature.
-The data is then divided on that split
-  and iterative dichotomization is called
-  recursively on the data in each split. In this way,
-  each split is the root of a sub-tree.
-
-  The CART iterative dichotomizer~\cite{breiman84}
-  measures   variability  
-  using the  standard deviation $\sigma$ of the defect counts.
-  If a code metric is split into $r$ ranges and each range is of
+  to find the fewest differences between clusters of examples.
+
+ XTREE uses an iterative dichotomizer that
+first converts the values for each code
+metrics into a small number of ranges.
+Consider a code metric is split into $r$ ranges and each range is of
   size
@@
   associated with a set of defect counts $x$ with standard deviation
-  $\sigma_r$ then the best split is the one that minimizes
+  $\sigma_r$.
+  The best split for that range is
+  the one that minimizes the expected value of the
+  defect variance, after the split; i.e.
   $\sum_r\frac{n_r}{n}\sigma_x$ (where $n=\sum_r n_r$).
-
-  While not usually labelled as such, CART can
+  This discretizer then recurses on each part of the split
+  to find other splits.
+
+  When discretization finishes, each code metric $M$ has a 
+  final expected value $M_v$ for the defect standard deviation 
+  across all the discretized ranges of that metric.
+  Iterative dichomization sorts the metrics by $M_c$
+  to find the best ``splitter'' for the data
+  (the code metric with smallest $M_v$). The data
+  is then split on the ranges of the splitter and
+  iterative dichomization recurses on each range.  In this way,
+  each split is the root of a sub-tree.
+
+  While not usually labelled as such, iterative dichomization can
   be viewed as a clustering algorithm that groups together code modules
   with similar defect counts and some shared ranges of code metrics.
-  XTREE is a post-processor to CART that reports the delta between 
-  nearby groups with different defect counts. \fig{xtree_samp} offers
-  a small example of that reflection. In this example, some project has a table of data describing its  classes in the form of \fig{ck}. After code inspections and running test cases or operational
+  For our purposes, we score each cluster found in this way according
+  to the percent of classes with known defects. For example,
+  the last line of \fig{xtree_samp} shows a tree lead with 75\%
+  defective modules.
+
+ \fig{xtree_samp} offers
+  a small example of how XTREE builds
+    $\Delta$ by comparing branches that lead to leaf clusters
+  with different defect percentages.
+In this example, assume a project with a table of code metrics data describing its classes in the form of \fig{ck}.
+After code inspections and running test cases or operational
       tests, each such class is augmented with a defect count.
-      CART takes that table of data and, 
+       Iterative dichomization takes that table of data and, 
         generates the tree of \fig{xtree_samp}.
 
-     Once the tree is built, some test case arrives and is passed
-      down the tree to a leaf note (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}.
+     Once the tree is built, a class with code metric data is passed into the tree and evaluated down the tree to a leaf node (see the \textcolor{orange}{{\bf orange}} line in \fig{xtree_samp}).
        XTREE then looks for a nearby leaf node with a lower defect
-      count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that test case, XTREE proposes a bad smell
+      count (see the \textcolor{green}{{\bf green}} line in \fig{xtree_samp}). For that evaluated class, XTREE proposes a bad smell
       thresholds that is  the difference between 
@@
 
-Using the training data,  XTREE uses CART to
+Using the training data,  XTREE uses  
+iterative dichomization to
   divide  $N$ code modules  into  clusters of
-size $\alpha=\sqrt{N}$.
+size $\sqrt{N}$.
 
 Next, for each test code module, 
-	  find $C_+$ : take each test , run it down to a leaf in the decision tree.  
+	  find $C_+$ as follows take each test , run it down to a leaf in the decision tree.  
 After that,	  find $C_-$ as follows:
@@
 		\item Identify {\em sibling} leaves; i.e. leaf clusters that can be reached from level $lvl$ that are not same as {\em current }
-		\item Using the {\em score} defined above, find the {\em better} siblings; i.e. those with a {\em score} less than $\gamma=0.5$ times the mean score of {\em current}. 
+		\item Find the {\em better} siblings; i.e. those 50\% (or less)
+		fewer defects than $C_+$.
 		   If none found, then repeat for $lvl += 1$. Also,
@@
 \subsection{Research Questions}
-To define our experiments, we  review some details
-about XTREE.
+% To define our experiments, we  review some details
+% about XTREE.
 
@@
 conjunctions of changes that might perform better than
-the disjunctions proposed by Shatnami and Alves et al.
-Further,
-unlike CD, XTREE   should be less verbose since
-it just reports the deltas in the code
-metrics selected by a decision tree learner.
-
-In order to test the value of the above,
-we propose three research questions.
+the disjunctions proposed by Shatnawi and Alves et al.
+Unlike CD, XTREE should be less verbose since
+it only reports the deltas in the code
+metrics rather than values for all metrics.
+To evaluate XTREE,
+we propose two research questions.
 
@@
 To find an answer to this question, we will assume that developers
-refactor their code until the bad smell thresholds no longer apply.
-This refactor will start with some {\em initial} code
-base which is changed to a {\em new} code base. 
+refactor their code until the bad smell thresholds is not violated.
+This refactoring will start with some {\em initial} code
+base that is changed to a {\em new} code base. 
 For example, if the bad smell is \mbox{{\em loc $>100$}} and a 
-code module has (e.g.) 500 lines of code, we reason
+code module has 500 lines of code, we reason
 optimistically that we can change that code metric
@@
 predictor learned via Random Forest, we then predict the
-number of defects in $d_0,d_1$ in {\em initial} and {\em new}.  
-The best bad smell threshold detector is the one that maximizes
+number of defects in $d_+,d_-$ in {\em initial} and {\em new}. We evaluate the performance of XTREE, CD, Shatnawi, and Alves methods in setting bad smell thresholds.
+The best bad smell threshold method is the one that maximizes
 \begin{equation}\label{eq:diff}
-\mathit{improvement} = 100* \left(1 - \frac{ d_1 }{ d_0}\right)
+\mathit{improvement} = 100* \left(1 - \frac{ d_- }{ d_+}\right)
 \end{equation}
@@
 
-\subsubsection{RQ3: Co-change}
-
-{\em Does XTREE understand what
-metrics get changed together? }
-
-For example, in XTREE's thresholds,  if {\em loc} is
-reduced, does some coupling measure increase?
+To measure verbosity, we will compare the counts of how often
+different approaches assign thresholds to code metrics
+(and the one that makes the least assignments is the least verbose).
+
+% \subsubsection{RQ3: Co-change}
+
+% {\em Does XTREE understand what
+% metrics get changed together? }
+
+% How often does XTREE make recommendations that
+% use more than one code metric? 
+% For example, continuing the scenario described in the introduction,
+%   if {\em loc} is reduced, does XTREE also recommend that some coupling measure increase?
 
@@
 we used data from
-the Jureczko et al.'s collection of object-oriented JAVA systems~\cite{jureczko10}. To access that data, goto   git.io/vGYxc.
-
+Jureczko et al.'s collection of object-oriented Java systems~\cite{jureczko10}. To access that data, go to   git.io/vGYxc.
 The Jureczko data records the number of known defects for each class using a post-release bug tracking system.
-The classes are described in terms of nearly two dozen metrics, included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc.
+The classes are described in terms of nearly two dozen metrics included in the Chidamber and Kemerer metric suite, such as number of children (noc), lines of code (loc), etc.
 For details on the Jureczko code
@@
 \label{sect:eval}
+We must be able to assess how defective the code is before and after a proposed refactoring.
+Since we do not have actual  examples of defect counts post-refactoring, we   approximate   defect counts after the   changes   (as changes to the code metrics).
+
 Prior work~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} 
-used the QMOOD hierarchical quality model ~\cite{Bansiya02} to assess the results of their refactorings
-Since the QMOOD method is quite old (2002), we  used Random Forests~\cite{Breiman2001} to build
+used the QMOOD   quality model ~\cite{Bansiya02} to assess the results of their refactorings
+Since the QMOOD method based on old (2002) project data, we use Random Forests~\cite{Breiman2001} to build
 the assessment methods.
-
   Random Forests are a decision tree learning method but
@@
   come from averaging the predictions over all the trees.
-  Recent students rank Random Forests as  the top-most
-  effective defect prediction learning algorithm~\cite{lessmann}.
+  Recent studies endorsed the use
+  of  Random Forests for  defect prediction~\cite{lessmann}.
 
@@
 the Jureczko data.
-Given access to $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}.
+Given   $V$ released versions, we test on version $V$ and train on the available data from $V-1$ earlier releases (as shown in \fig{j}.
 Note the   \colorbox{lavenderpink}{three bottom rows}   marked with $\times$: these contain predominately defective classes (two-thirds, or more).
-In such data sets, it is hard to distinguish good from bad (since there are so many bad examples).
+In such data, it is hard to distinguish good examples (due to all the bad examples).
 
@@
 The ``untuned'' columns of \fig{j}
-show a preliminary study.
-Random Forests was built from the training data, then applied to the test data.
-We call a data set ``usable'' if Random Forest was able to classify majority of the instances correctly.
-For this purpose, we set a threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% (determined from standard results in other publications~\cite{me07b}).
-No  data set meet
+show a preliminary study. This study used
+Random Forest with its ``off-the-shelf'' tunings (i.e.  
+100 trees per forest).  
+The forests were built from training data and applied to test data
+not seen during training.  In this
+study, we called a data set ``usable'' if   Random Forest was able to classify the instances with a performance threshold of $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\% (determined from standard results in other publications~\cite{me07b}).
+Note that no  data set meet
 that criteria.
 
-Fortunately, the ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+The ``tuned'' columns of \fig{j} show that we can salvage some of the data sets.
+We applied both the SMOTE algorithm and differential evolution to improve the performance of the classifier.
 Pelayo and Dick~\cite{pelayo07} report that defect prediction is improved by SMOTE~\cite{Chawla2002}; i.e.
 an over-sampling of minority-class examples and an under-sampling of majority-class examples.
-Also, Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
-for tree generation, etc.
+Fu et al.~\cite{fu:ase15} report that parameter tuning with differential evolution~\cite{storn97} can quickly explore the tuning options of Random Forest to find better settings for the (e.g.) size of the forest, the termination criteria
+for tree generation, etc. We note that SMOTE-ing and
+parameter tunings were applied to the training data only and not to the test data.
+
 The rows \colorbox{celadon}{marked with a $\star$} in \fig{j} show data sets whose performance was improved remarkably by these techniques.
@@
 
-In summary, while we cannot trust predictors from some of our Jureczko data sets,
-we can plan ways to reduce defects in {\em jedit, ivy, ant, lucene} and {\em poi}.
-Accordingly, when this study explores the Jureczko data, we will use these five data sets.
-
-(Aside: One important detail to be stressed here is that, when we applied    SMOTE-ing and
-parameter tunings, those techniques were applied to the training data and {\em not}
-the test data; i.e. we took care that no clues from the test set were ever used in this tuning process.)
+We eliminate the data sets for which we could not build an adequately performing Random Forest classifier with $\mathit{pd}\ge 60 \wedge \mathit{pf} \le 30$\%.
+Thus, our analysis uses the {\em jedit, ivy, ant, lucene} and {\em poi} for evaluating recommended changes.
 
 \subsection{Statistics}
-Two of our  methods  (Shatnami, Alves et al.) are fully deterministic
-(which is to say, if we run them $N$ times, we will always get the same result).
-However, our other methods use some random choices: CD uses the  stochastic WHERE clustering algorithm ~\cite{Menzies2013:local}
+
+We use 40 repeated runs, each with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
+Each run collects the \eq{diff} values.
+We use multiple runs since two of our methods use some random choices: CD uses the  stochastic WHERE clustering algorithm~\cite{Menzies2013:local}
 while XTREE non-deterministically picks thresholds randomly from
@@
 
-We use  40 repeated runs, each with different random number seeds (we use 40 since that is  more than the 30 samples  needed to satisfy the central limit theorem).
-Each run collects the \eq{diff} values.
-
-To rank these 40 numbers collected from CD, XTREE,Shatnami, and Alves et al., we  use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}. 
-In accordance to that test, using the median values of each method, we
-sort a list of  $l=40$ values of \eq{diff} values found in  $ls=4$ different methods. 
+\input{tex/results_jur.tex}
+
+To rank these 40 numbers collected from CD, XTREE, Shatnawi, and Alves et al., we use the Scott-Knott test recommended by Mittas and Angelis~\cite{mittas13}. 
+Scott-Knott is a top-down clustering approach used to rank different
+treatments. If that clustering finds an interesting division of the data, then
+some statistical test is applied to the two divisions to check if they
+are statistically significant different. If so, Scott-Knott recurses
+into both halves.
+
+In implement Scott-Knott using our data,
+  we
+sorted a list of  $l=40$ values of \eq{diff} values found in  $ls=4$ different methods. 
 Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences in the observed performances before and after divisions.
@@
 
-The Scott-Knott  results are presented in the form of line diagrams like those shown on the right-hand-side of \fig{jur}
-The black dot shows the median \eq{diff} values and the horizontal likes stretch 
-from the 25th percentile to the 75th percentile (a region called the inter-quartile
-range, or IQR).
+The Scott-Knott  results are presented in the form of line diagrams like those shown on the right-hand-side of \fig{jur}.
+The black dot shows the median \eq{diff} values and the horizontal lines stretch 
+from the 25th percentile to the 75th percentile (the inter-quartile range, IQR).
 As an example of how to read this table, consider the {\em Ant}
@@
 
-\input{tex/results_jur.tex}
-
 \input{tex/rq2.tex}
-
+\newpage
 \subsection{Results}
 
-\begin{figure*}[t!]
+\begin{figure*}[!t]
 \centering
 \includegraphics[width=\linewidth]{figs/changes01.png}
-\caption{XTREE results.
+\caption{Results  from XTREE.
 While \fig{counts} are the {\em number} of times a code metric was changed,
-\fig{changes} shows {\em how far} each code metric was changed. Each vertical bar
-marks the 27,50,75th percentile change seen in 40 repeats. }
-\label{figs:changes}
+this  figure shows {\em how far} each code metric was changed. Each vertical bar
+marks the 27,50,75th percentile change seen in 40 repeats.
+All numbers are ratios of initial to final values.
+All bar regions marked in gray show {\em increases}.
+The interesting feature of these results are that many
+of the changes proposed by XTREE require {\em increases}
+(this puzzling observation is explained in the text).}
+\label{fig:changes}
 \end{figure*}
@@
 
-\fig{jur} shows the comparison results.  The Alves and Shatnami
-entries show the  net effects of applying (seperately) all the disjuncts
-proposed by those methods.
+\fig{jur} shows the comparison results.  The Alves and Shatnawi
+entries show the net effects of applying (separately) all the disjuncts
+proposed by those methods. In this part of the  analysis,
+we applied Scott-Knott to all the \eq{diff} values
+seen when code metrics were changed
+according to the  Alves and Shatnawi
 
@@
  find the same ways to reduce the expected number of
-defects.
+defects.   For an explanation of the Jedit uniformity, see \tion{inc}.
 
@@
 both these data sets contain more than 50\% defective modules.
-In that space, all our  methods lack a large sample to
+In that space, all our  methods lack a large sample of
 defect-free examples.
 
-More important that the size of the improvement is the relative
+Also consider the relative
 rank of the different approaches,
@@
   after refactoring,
-XTREE is marginally better than Alves while CD and Shatnawi
-are deprecated.
+XTREE is marginally better than Alves. CD and Shatnawi.
+on the other hand, usually perform worst.
 \end{lesson}
 
+Another interesting feature of the XTREE results of
+\fig{jur} was that the used code metrics vary
+from data set to data set. While changing lines of
+code {\em loc} was important in all data sets, 
+the other code metrics used were often different.
+
  \subsubsection{RQ2: Verbosity}
@@
 
-\fig{counts} shows the frequency with which our different methods
-require changes to code methods.
+\fig{counts} shows the frequency with which the methods
+recommend changes to specific code methods.
 Note that XTREE proposes thresholds to
-few code metrics than the other approaches.
-
-\begin{lesson}
-In  terms of how often   a bad
-smell can be ignored, XTREE is best since it is  least verbose (i.e. least
-likely to  overlap   a bad smell mentioned by a developer).
-\end{lesson}
+few code metrics compared to the other approaches.
+
+% COMMENT From LL: This is such a strange way of putting this lesson, and you haven't actually evaluated this.
+%
 
@@
 
-\subsubsection{RQ3: Co-Change}
-
-{\em Does XTREE understand what metrics get changed together?}
-xxxx
+\begin{lesson}
+XTREE finds finds far fewer recommended changes to code metrics, and yet performs as good as or better than other methods in recommending changes that results in fewer defects.
+\end{lesson}
+
+\subsubsection{Discussion}\label{sect:inc}
+
+In \fig{changes} we identify the magnitude of changes proposed by the plans offered by XTREE.
+At first glance, \fig{changes} seems strange since it 
+shows many metrics need to be {\em increased} in order
+to reduce defects. Note that such increases  to reduce
+defects would never be proposed by the outlier methods
+of Shatnawi or Alves since their core assumption is that bad
+smells arise from unusually large code metrics.
+
+The increases proposed by 
+\fig{changes} arise from the interdependancies between
+code metrics. For example, consider the {\em Poi} results
+from \fig{counts} that recommends decreasing {\em loc}
+but making large increases to {\em cbo} (coupling between
+objects). Here, XTREE is advising us to break up
+large classes class by into services
+in other classes. Note that such a refactoring will, by
+definition, increase the coupling between objects.
+
+ % sorry. ur still working. backing off
+ % is this next para your "lesson" bubble? note that the main
+ % thing here is to celebrate as expected how sometimes
+ %XTREE increases certain measures.
+
+% I see.. I've tried to write about that below
+
+It is worth noting that XTREE suggests that it is perhaps better to reduce only certain forms of coupling.
+Refactoring operations that reduce one form of coupling at the cost of increasing another can sometimes be beneficial.
+This is most noticeable in Ivy, where one of the better approaches is to reduce coupling between objects (cbo) (efferent coupling) while increasing the afferent coupling (CA), again implying a redistribution of responsibilities.
+Using thresholds may not always appreciate the existence of these co-changes between metrics, and as a result many refactoring operations are usually left unexplored.
+
+Note that we prefer XTREE to Shatnawi and the Alves
+method since those other methods
+do not reason
+about decreases {\em and} increases of multiple 
+code metrics, at the same time.
+
+As a final comment, we note
+that  \fig{changes} enables us to explain the uniformity
+of the results seen with Jedit in \fig{jur}.
+Observe how in \fig{changes} the only change ever
+found is a reduction to {\em rfc}. Clearly, in this
+data set, there is very little that can be usefully changed.
+
+% \fig{changes} shows the mag
 
@@
 
-As with any empirical study, biases can affect the final results. Therefore, any
-conclusions made from this work must be considered with the following issues in
-mind.
+As with any empirical study, several issues can affect the final results. Therefore, any
+conclusions made from this work must be offered with the following
+caveats.
 
@@
 as well as the discussion in \tion{prelim},
-we believe that specific bad smell indicators (e.g. \mbox{{\em loc}$>$ 100})
-have very limited external validity. 
-The good news is that,
-while specific models are not externally valid,
+we believe that bad smell indicators (e.g. \mbox{{\em loc}$>$ 100})
+have limited external validity beyond the projects from which they are derived. 
+While specific models are externally valid,
 there may still be general methods like XTREE for finding the good local models.
 
-One   limitation to certifying bad smells
-is the terminology used to express the bad smell.
+Our definition of bad smells is limited to those represented by OO code metrics (a premise often used in related work).   
 XTREE, Shatnawi, Alves et al. can  only comment
-on refactorings expressed in the   terminology
-as the data
+on bad smells   expressed as code metrics 
   in the historical log of a project.
+
 If developers want to justify their refactorings
 via bad smells expressed in other terminology,
-then the  analysis of this paper must   wait till 
+then the  analysis of this paper must:
+\begin{itemize}
+    \item Either wait till 
 data about those new
 terms has been collected. 
-
-XTREE does not require a working defect predictor to generate
-its bad smell thresholds. However, to {\em certify}
-that those thresholds are satisfactory, some
-second quality predictor must be applied. Historically,
-SE researchers~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} have used the QMOOD hierarchical quality model~\cite{Bansiya02}.
-Since such quality models may be very project specific~\cite{localvsglobal}, we prefer building local 
-defect predictors using Random Forests. 
-Our results show that when a defect log contains too few
-non-defective examples then defector predictors
-perform poorly (see the data sets in the last few lines of \fig{j})
-as do their associated bad smell threshold generators
-(recall the performance of Lucene and Poi in~\fig{jur}).
-
-It should be noted that this is not a limitation of XTREE-- rather
-it seems fundamental to the problem of certifying bad smells.
-As evidence of this, recall that {\em all} the bad smell threshold generators of~\fig{jur}
-had issues with the defect data sets with  many defects
-(Lucene and Po).
+\item Or, apply cutting edge transfer learning
+methods~\cite{Nam15,Jing15} to map data from other projects
+into the current one.
+\end{itemize}
+Note that the transfer learning approach would
+be highly experimental and require more study
+before it can be safely recommended.
+
+Sampling bias threatens any data mining experiment; i.e., what matters
+there may not be true here. For example, the data sets used here comes from Jureczko et al. and any biases in their selection procedures
+threaten the validity of these results. 
+That said,
+the best we can do is define our methods and publicize our data and code so that other researchers can
+try to repeat our results and, perhaps, point out a previously unknown bias
+in our analysis. Hopefully, other researchers will emulate our methods in
+order to repeat, refute, or improve our results.
+
+% Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.  
+% That said, we have taken care to document in this paper the decisions made by engineering
+% judgement that could effect our conclusions.
 
@@
 
-\subsection{ Learner Bias}
-This study used two kinds of learners: (1) methods for suggesting
-changes to code like Shatnawi, Alves et al., CD and XTREE;
-(2) methods for assessing the changes proposed by these learners.
-Prior studies implemented (2) using  the QMOOD hierarchical quality model
-which is so old, that we suspect it may no longer be current. Hence,
-we used Random Forests to build the assessment oracles that were specific
-to the data sets used in this study. .We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.
-
-Data mining is a large and active field and any single study can only use a small subset of the known classification algorithms.  
-That said, we have taken care to document in this paper the decisions made by engineering
-judgement that could effect our conclusions.
-
-\subsection{  Sampling bias} 
-Sampling bias threatens any data mining experiment; i.e., what matters
-there may not be true here. For example, the data sets used here comes from two  Jureczko et al. and any biases in their selection procedures
-threaten the validity of these results. 
-That said,
-the best we can do is define our methods and publicize our data and code so that other researchers can
-try to repeat our results and, perhaps, point out a previously unknown bias
-in our analysis. Hopefully, other researchers will emulate our methods in
-order to repeat, refute, or improve our results.
-
-\subsection{  Evaluation Bias}\label{sect:coc}
-
-The main evaluation bias in this study was the use of
-optimistic reasoning when generating the changed code modules.
+\subsection{  Internal Validity}\label{sect:coc}
+
+% The main internal validity threat   in this study was the use of
+% optimistic reasoning when generating the changed code modules.
 We assumed in \tion{rq}
@@
 true, as noted in {\em PROBLEM \#2} in this paper's introduction.
-To mitigate this issue, we do not propose using XTREE as a bad
-smell generator, but as a critique agent for bad smells
+To mitigate this issue:
+\begin{itemize}
+    \item We do not propose using XTREE as a bad
+smell generator;
+\item Rather, we offer it as  a critique agent for bad smells
 proposed by some other oracle (e.g. a human developer).
-We showed about that XTREE's recommendations are not verbose
-so it will not often overlap with some random bad smell proposed
-by a developer.
+\end{itemize}
+We showed about that XTREE's recommendations are not verbose.
+That is, even if XTREE recommends the superset of possibly important bad smells,
+that superset is not so large as to  overlap with many
+ bad smell proposed at random 
+by   developers.
+
+XTREE does not require a working defect predictor to generate
+its bad smell thresholds. However, to {\em certify}
+that those thresholds are satisfactory, some
+second quality predictor must be applied.
+A data set containing defect counts before and after refactoring activities would enable an accurate assessment of the performance of the threshold-setting and recommender methods, however, such a data set is not available at this time.
+
+ Historically,
+SE researchers~\cite{Cheng10,OKeeffe08,OKeeffe07,Moghadam2011,Mkaouer14} have used the QMOOD hierarchical quality model~\cite{Bansiya02}.
+Since such quality models may be very project specific~\cite{localvsglobal}, we prefer building local 
+defect predictors using Random Forests. We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.  
+Our results show that when a defect log contains too few
+non-defective examples then defector predictors
+perform poorly (see the data sets in the last few lines of \fig{j})
+as do their associated bad smell threshold generators
+(recall the performance of Lucene and Poi in~\fig{jur}). It should be noted that this is not a limitation of XTREE-- rather
+it seems fundamental to the problem of certifying bad smells.
+As evidence of this, recall that {\em all} the bad smell threshold generators of~\fig{jur}
+had issues with the defect data sets with  many defects
+(Lucene and Poi).
+
+% \subsection{ Learner Bias}
+% This study used two kinds of learners: (1) methods for suggesting
+% changes to code like Shatnawi, Alves et al., CD and XTREE;
+% (2) methods for assessing the changes proposed by these learners.
+% Prior studies implemented (2) using  the QMOOD hierarchical quality model
+% which is so old, that we suspect it may no longer be current. Hence,
+% we used Random Forests to build the assessment oracles that were specific
+% to the data sets used in this study. We chose this approach,  based on its reputation for having the better  performance of 21 other learners for defect prediction~\cite{lessmann}.
+
+% \subsection{  Sampling bias} 
+% Sampling bias threatens any data mining experiment; i.e., what matters
+% there may not be true here. For example, the data sets used here comes from two  Jureczko et al. and any biases in their selection procedures
+% threaten the validity of these results. 
+% That said,
+% the best we can do is define our methods and publicize our data and code so that other researchers can
+% try to repeat our results and, perhaps, point out a previously unknown bias
+% in our analysis. Hopefully, other researchers will emulate our methods in
+% order to repeat, refute, or improve our results.
 
@@
 
+\newpage
+
 \section{Conclusions}
-When can bad smells be ignored? Our answer is as follows.
-When those
-bad smells are not supported by the historical log of data from
+Which  bad smells be ignored? 
+We say: ignore those not supported by the historical log of data from
 the current project.
@@
 \begin{itemize}
-    \item As described in \tion{prelim}, bad smells generated by humans may not be apply to the current project. On the other hand, XTREE can learn specific thresholds for the current project.
-    \item As described in in \tion{}, XTREE generates conjunctions
-    \item Prior methods used an old quality predictor (QMOOD) which we replace with defect predictors learned from the current project.
+    \item As described in \tion{prelim}, bad smells generated by humans may not be applicable to the current project. On the other hand, XTREE can automatically learn specific thresholds
+    for bad smells for the current project.
+    \item Prior methods used an old quality predictor (QMOOD) which we replace with defect predictors learned via Random Forests
+    from   current project data.
+    \item As described in in \tion{bst}, 
+    XTREE generates conjunctions of code metric thresholds
+    that need to be satisfied at the same time.
+    Older methods, such as those proposed by  Shatnawi and Alves
+    offer only disjunctions that might be inappropriately applied
+    to yield impossible refactorings (e.g. reducing lines of code
+    and coupling at the same time).
+
+    \item XTREE's conjunctions proved to be arguably as effective
+    as those of Alves (see \fig{jur}) but far less verbose (see \fig{counts}) thus making XTREE a better tool for critiquing and rejecting
+    more bad smells proposed by developers.
 \end{itemize}
+In summary, this paper offers a baseline result
+that XTREE can be used as a {\em diagnostic tool}
+for identifying which bad smells can be ignored.
+But can it be used as a {\em prescriptive tool}
+to recommend new refactorings?
+
+As discussed in {\em PROBLEM \#2}
+(from the introduction), XTREE might propose
+changes to code metrics that are
+not feasible  (due to the local domain constraints
+of the code). To repair this, XTREE needs to be extended
+with architectural knowledge of the code it is studying.
+Elsewhere, we have had some   success
+with reasoning about such architectural knowledge
+(expressed as software product lines~\cite{sayyad13a,sayyad13b}). In future
+work, we will explore extending XTREE 
+with  
+architectural knowledge such that it can
+become a tool which perscribes
+feasible refactorings.
 
 \section*{Acknowledgements}
-The work has partially funded by NSF  award \#1506586.
+The work is partially funded by NSF  awards \#1506586 and \#1302169.
 \bibliographystyle{plain}
@@
 \balance
+\bibliographystyle{unsrt}
 \bibliography{References}
@@
 
-\input{tex/results_jur.tex}
-
 \subsection{Experimental Results}

### 2016-03-12 tex/prediction_stats.tex
final save for the fse submission

@@
 		Data in  the  \colorbox{lavenderpink}{three bottom rows}, marked with ``$\times$'', are  performing
-		poorly-- that data so few non-defective examples  that it  is hard for
+		poorly-- that data so many defective examples  that it  is hard for
 		our learners to distinguish between classes.

### 2016-03-12 tex/refactor.tex
final save for the fse submission

@@
 \centering
-\begin{tabular}{r|c|c|c|c|c}
+\begin{tabular}{r|c|c|c|c}
 
-\begin{turn}{75}Fowler'99~\cite{fowler99}, and~\cite{Kerievsky2005})\end{turn} &\begin{turn}{75} Lanza'06~\cite{Lanza2006}\end{turn} & \begin{turn}{75}SonarQube~\cite{sq15} \end{turn} &  \begin{turn}{75}Yamashita'13\cite{Yamashita2013} \end{turn}& \begin{turn}{75}Atwood'06\cite{Atwood06}\end{turn}  & \begin{turn}{75} Developer Survey 2015\end{turn}\\\hline
-  Alt. Classes with Diff. Interfaces & & & & \checkmark & \\
-  Combinatorial Explosion~\cite{Kerievsky2005} & & & & \checkmark & \\
-  Comments & & & 11 & \checkmark & VL\\
-  Conditional Complexity~\cite{Kerievsky2005} & & & 14 & \checkmark & ?\\
-  Data Class & \checkmark & & & \checkmark &\\
-  Data Clumps &  &  & & \checkmark &\\
-  Divergent Change & & & & \checkmark & \\
-  Duplicated Code & \checkmark & \checkmark & 1 & \checkmark & VH\\
-  Feature Envy & \checkmark & & 8 & \checkmark &\\
+\begin{turn}{75}Fowler'99~\cite{fowler99} and~\cite{Kerievsky2005}\end{turn} &\begin{turn}{75} Lanza'06~\cite{Lanza2006}\end{turn} & \begin{turn}{75}SonarQube~\cite{sq15} \end{turn} &  \begin{turn}{75}Yamashita'13\cite{Yamashita2013} \end{turn}& \begin{turn}{75} Developer Survey 2015\end{turn}\\\hline
+  Alt. Classes with Diff. Interfaces & & & & \\
+  Combinatorial Explosion~\cite{Kerievsky2005} & & &  & \\
+  Comments & & & 11 & VL\\
+  Conditional Complexity~\cite{Kerievsky2005} & & & 14  & ?\\
+  Data Class & \checkmark & &  &\\
+  Data Clumps &  &  &  &\\
+  Divergent Change & & &  & \\
+  Duplicated Code & \checkmark & \checkmark & 1  & VH\\
+  Feature Envy & \checkmark & & 8  &\\
 
-  Inappropriate Intimacy & & \checkmark & & \checkmark & L\\
-  Indecent Exposure~\cite{Kerievsky2005} & & & & \checkmark & ?\\
-  Incomplete Library Class & & & & \checkmark &\\
-  Large Class & \checkmark & \checkmark & 4 & \checkmark & VH\\
-  Lazy Class/Freeloader & & \checkmark & 7 & \checkmark &\\
-  Long Method & \checkmark& \checkmark & 2 & \checkmark & VH\\
-  Long Parameter List &  & \checkmark & 9 & \checkmark & L \\
+  Inappropriate Intimacy & & \checkmark &  & L\\
+  Indecent Exposure~\cite{Kerievsky2005} & & &  & ?\\
+  Incomplete Library Class & & &  &\\
+  Large Class & \checkmark & \checkmark & 4  & VH\\
+  Lazy Class/Freeloader & & \checkmark & 7  &\\
+  Long Method & \checkmark& \checkmark & 2  & VH\\
+  Long Parameter List &  & \checkmark & 9  & L \\
 
-  Message Chains & & & & \checkmark & H\\
-  Middle Man & &  & & \checkmark &\\
-  Oddball Solution~\cite{Kerievsky2005} & & & & \checkmark & \\
-  Parallel Inheritance Hierarchies & & & & \checkmark &\\
-  Primitive Obsession &  & & & \checkmark &\\
+  Message Chains & & &  & H\\
+  Middle Man & &  &  &\\
+  Oddball Solution~\cite{Kerievsky2005} & & &  & \\
+  Parallel Inheritance Hierarchies & & &  &\\
+  Primitive Obsession &  & &  &\\
 
-  Refused Bequest & \checkmark & \checkmark & & \checkmark & \\ 
-  Shotgun Surgery & \checkmark& & & \checkmark & \\
-  Solution Sprawl~\cite{Kerievsky2005} & & & & \checkmark &\\
-  Speculative Generality & & & & \checkmark & L\\
-  Switch Statements &  & \checkmark & & & L\\
-  Temporary Field & & \checkmark & & \checkmark & ?\\
+  Refused Bequest & \checkmark & \checkmark &  & \\ 
+  Shotgun Surgery & \checkmark& &  & \\
+  Solution Sprawl~\cite{Kerievsky2005} & & &  &\\
+  Speculative Generality & & &  & L\\
+  Switch Statements &  & &  & L\\
+  Temporary Field & & \checkmark &  & ?\\
   \end{tabular}
@@
 in other studies.
-Note:    there are many blank cells.}
+Note:\   there are many blank cells.}
 \label{fig:smells}

### 2016-03-13 Manuscript.tex
Updates from ShareLaTeX

@@
 Refactored code cannot be assessed just by a rerun of the the test
-suite since (refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
+suite (since refactorings are supposed to preserve behavior), and otherwise one must wait for new defect reports to evaluate the effect of refactoring changes on code.  
 % waiting to observe 
@@
 \end{equation}
-For consistency
-with Shatnawi, we explore the correlation between
-the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
+
+\newpage
 \subsubsection{ Alves et al.}
@@
 exploring bad smells.
+
+For consistency
+with Shatnawi, we explore the correlation between
+the code metrics and the defect counts and   reject code metrics that are poor predictors of defects (i.e.   those  with $p > 0.05$).
 
@@
 Each run collects the \eq{diff} values.
-We use multiple runs since two of our methods use some random choices: CD uses the  stochastic WHERE clustering algorithm~\cite{Menzies2013:local}
+We use multiple runs since two of our methods use some random choices: CD uses the  stochastic WHERE clustering algorithm~\cite{localvsglobal}
 while XTREE non-deterministically picks thresholds randomly from
@@
 
-In implement Scott-Knott using our data,
+To  apply Scott-Knott,
   we
@@
 \input{tex/rq2.tex}
-\newpage
+
 \subsection{Results}
@@
 \section{Conclusions}
-Which  bad smells be ignored? 
+Which  bad smells can  be ignored? 
 We say: ignore those not supported by the historical log of data from
@@
 
-In accordance to that test, using the median values of each method, we
+In accordance to that test,  we
 sort a list of  $l=40$ values of $R$ values found in  $ls=4$ different methods. 
-Then, we split $l$ into sub-lists $m,n$ in order to maximize the expected value of differences  in the observed performances before and after divisions.
+Then, we split $l$ into sub-lists $m,n$ as to maximize the expected value of differences  in   observed performances before and after divisions.
 E.g.

### 2016-03-13 Manuscript.tex
Updates from ShareLaTeX

@@
 To address this problem, we propose the XTREE
-tool to judge refactorings
-proposed by developers.
-  XTREE requires:
+tool that uses: 
      \begin{itemize}
@@
      \end{itemize}
-Given that information,  XTREE can determine which
+Given that,  XTREE can determine which
   proposed refactorings are to be ignored
 since they are not justifiable  
-with respect to the history of the project.
-
-XTREE addresses four problems that  make it hard to  learn useful
+with respect to the history of the project. XTREE addresses four problems that  make it hard to  learn useful
 bad smells for a particular project.
@@
 {\em PROBLEM \#2:} Not all bad smells can be fixed. 
-Consider
-a bad smell stemming from a long method,
-yet that method must be passed as a single closure
-to a third-party GUI toolkit.  This is one
-example of a common problem-- when new code must
-be combined with code from other developers,
-sometimes the constraints imposed by the other code
-lead unavoidably to bad smells. In this case,
+When new code must
+be combined with code from a third party, 
+sometimes the constraints of that code 
+lead unavoidably to bad smells violations (e.g.
+developers might need to passed as a single closure a  very long method
+to  a third-party GUI toolkit).   In this case,
 a refactoring proposed to reduce that bad smell
@@
 
-\newpage
 \subsubsection{ Alves et al.}
@@
 \input{tex/prediction_stats.tex}      
-
+	\newpage     
 \section{Evaluation}
@@
 \input{tex/rq2.tex}
-
+ \newpage
 \subsection{Results}
